---
title: "04.3_rf_SDM"
author: "Markus Min"
date: "`r Sys.Date()`"
output: html_document
---

## Description

In this document, we conduct the model selection process for YOY rockfishes from the PWCC/PRS survey data. We identify the best fit model, which is then used to model the distribution of CSSIF in the joint distribution models found in the folder `/05_stage1_SDM`.


### Model selection steps

1. Fit a "base" model, that includes all fixed effects but no random effects.

2. Use this base model to tune the structure of the spatial and spatiotemporal random effects using cAIC as the information criterion.

3. Keeping the random effects structure identified in step 2 fixed, identify the fixed effect structure for our covariates (year, temperature, distance from shore), again using cAIC as the information criterion. This includes exploring two alternative functional forms for temperature and distance from shore: linear and p-splines.

```{r chunk-options, message = FALSE, echo = FALSE, warning = FALSE}
# set chunk options
knitr::opts_chunk$set(message = FALSE, echo = FALSE, warning = FALSE)
```


## Data preparation for model selection process

```{r load_libraries_source_scripts, include = FALSE}
library(tidyverse)
library(here)
library(TMB)
library(ggforce)
library(sdmTMB)
library(kableExtra)
library(mgcv)

# source the function scripts
source("R/functions/cAIC.R")
source("R/functions/rmvnorm_prec.R")
source("R/functions/sample_var.R")
source("R/functions/plot_anisotropy_MM.R")
source("R/functions/make_anisotropy_spde.R")
source("R/functions/H_matrix_prior_predictive_check_for_ggplot.R")
source("R/functions/plot_distribution_PRS_PWCC.R")
source("R/functions/generate_prediction_maps.R")
source("R/functions/sdmTMB_smoothing_functions.R")

# source the script to load spatial data and create the mesh
source(here::here("R", "PRS_PWCC_make_mesh.R"))
```

```{r data_prep}
# make sure rf density is numeric not integer
rf$total <- as.numeric(rf$total)

#### prep model data ####
# generate the projection vector/matrix for covariates - this relies
# on the survey prediction grid from the PRS_PWCC_make_mesh.R script

distance_projection_vector <- as.numeric(filter(survey_predict_grid_jsoes_rf_years, year == 2016)$dist_shore_scaled)

# for SST - take prediction grid, drop salinity, add SST with locations as rows and years as column
survey_predict_grid_jsoes_rf_years %>% 
  # select only the years that we're keeping
  filter(!(year %in% c(1998, 2022:2025))) %>% 
  dplyr::select(-c(SST, dist_shore, dist_shore_scaled)) %>% 
  pivot_wider(names_from = year, values_from = SST_scaled) %>% 
  dplyr::select(-c(X, Y)) %>% 
  as.matrix() -> temperature_projection_matrix

# create a second temperature projection grid that contains the missing years
survey_predict_grid_jsoes_rf_allyears %>% 
  # select only the years that we're keeping
  filter(!(year %in% c(1998, 2022:2025))) %>% 
  dplyr::select(-c(SST, dist_shore, dist_shore_scaled)) %>% 
  pivot_wider(names_from = year, values_from = SST_scaled) %>% 
  dplyr::select(-c(X, Y)) %>% 
  as.matrix() -> temperature_projection_matrix_rf_allyears


#### Create SPDE objects ####

# there are multiple meshes that are created in JSOES_make_mesh.R
# here, we will use the cutoff 15 mesh, although a higher resolution is likely going
# to improve results (at the cost of run time)

PRS_PWCC_spde <- fm_fem( inla_mesh_cutoff15_jsoes_domain, 
                order = 2 )

# create projection matrix from vertices to samples
A_is_rf = fm_evaluator(inla_mesh_cutoff15_jsoes_domain, loc = as.matrix(data.frame(X = rf$X, Y = rf$Y)))$proj$A

# create projection matrix from vertices to prediction grid
# use the prediction grid for only the JSOES domain here

#### Create a survey prediction grid 
# Create a survey grid from scratch, using SST data

# only keep one year, since that's all we need to make a grid
SST_sf_proj %>% 
  filter(year == 2000) -> SST_sf_proj_2000

st_as_sf(SST_sf_proj_2000$geometry/1000, crs = UTM_zone_10_crs) -> SST_sf_proj_km_2000

SST_sf_proj_km_2000$dist_shore <- as.numeric(st_distance(SST_sf_proj_km_2000, US_west_coast_proj_km))


# KEY DECISION: To what area do you want to project the SDM to?
# Some observations of the data:
# The vast majority of stations are within 60 km of shore, but there are just a couple 
# that are further out in the Columbia River Plume up to 83 km from shore
# a total of 15 tows (out of over 2000) are >= 35 nmi (65 km) offshore; many more are 30 or 31 nmi (55.5 or 57.4 km) offshore.
# We also want to exclude values that are within 0.5 km of shore, because these
# don't have measureable SST values (they're basically on shore)
# this can also be seen in the following histogram:
# hist(jsoes_long$km_from_shore)

# For the N/S dimension, the most northerly transect is typically the FS transect and 
# the most southerly transect is the NH transect.
# this corresponds to 44.67 to 48.23 N
# Let's take these as our N/S boundaries and add 10 km as a buffer.
min_Y <- mean(subset(jsoes_long, grepl("NH", station))$Y)
max_Y <- mean(subset(jsoes_long, grepl("FS", station))$Y)

# Based on this, I'm going to propose that we generate a prediction that extends:
# from 0.5 to 65 km offshore
# from 44.67 to 48.23 N (+ 10 km on either side)

# trim longitudinally
SST_sf_proj_km_2000 %>% 
  filter(dist_shore <= 65 & dist_shore >= 0.5) -> grid_within_65km_shore

# trim latitudinally
survey_domain <- st_crop(grid_within_65km_shore, xmin = 0, xmax = 100000, ymin=min_Y-10, ymax=max_Y+10)

# let's inspect the grid that we created
# ggplot(survey_domain) +
#   geom_sf()

# We will need to manually trim out some of these areas

# manually remove sections for strait, hood canal (and other inland waters), Grays Harbor, Willapa Bay, Columbia River estuary
strait <- st_as_sfc(st_bbox(c(xmin=-124.65, xmax=-122, ymin=47.9375, ymax=49), crs = "WGS84"))
strait_proj <- sf::st_transform(strait, crs = UTM_zone_10_crs)
strait_proj_km <- st_as_sf(strait_proj/1000, crs = UTM_zone_10_crs)

inland_waters <- st_as_sfc(st_bbox(c(xmin=-123.5, xmax=-120, ymin=44, ymax=49.5), crs = "WGS84"))
inland_waters_proj <- sf::st_transform(inland_waters, crs = UTM_zone_10_crs)
inland_waters_proj_km <- st_as_sf(inland_waters_proj/1000, crs = UTM_zone_10_crs)

grays_harbor <- st_as_sfc(st_bbox(c(xmin=-124.15, xmax=-123.6, ymin=46.83, ymax=47.09), crs = "WGS84"))
grays_harbor_proj <- sf::st_transform(grays_harbor, crs = UTM_zone_10_crs)
grays_harbor_proj_km <- st_as_sf(grays_harbor_proj/1000, crs = UTM_zone_10_crs)

willapa_bay <- st_as_sfc(st_bbox(c(xmin=-124.06, xmax=-123.5, ymin=46.34, ymax=46.8), crs = "WGS84"))
willapa_bay_proj <- sf::st_transform(willapa_bay, crs = UTM_zone_10_crs)
willapa_bay_proj_km <- st_as_sf(willapa_bay_proj/1000, crs = UTM_zone_10_crs)

estuary <- st_as_sfc(st_bbox(c(xmin=-124.02, xmax=-123.5, ymin=46.12, ymax=46.35), crs = "WGS84"))
estuary_proj <- sf::st_transform(estuary, crs = UTM_zone_10_crs)
estuary_proj_km <- st_as_sf(estuary_proj/1000, crs = UTM_zone_10_crs)

survey_domain %>% 
  st_difference(., strait_proj_km) %>% 
  st_difference(., inland_waters_proj_km) %>% 
  st_difference(., grays_harbor_proj_km) %>% 
  st_difference(., willapa_bay_proj_km) %>% 
  st_difference(., estuary_proj_km) -> survey_domain

# Create a concave hull around points
st_concave_hull(st_union(survey_domain), ratio = 0.1) -> survey_domain_polygon

# ggplot(survey_domain_polygon) +
#   geom_sf()

# ggplot(survey_domain) +
#   geom_sf()

# let's inspect our revised survey domain
survey_area_basemap_km +
  geom_sf(data = survey_domain_polygon, fill = "blue", alpha = 0.2)

# Take SST grid and intersect with survey domain
SST_sf_proj %>% 
  mutate(geometry = geometry/1000) %>% 
  st_set_crs(UTM_zone_10_crs) %>% 
  st_intersection(survey_domain_polygon) %>% 
  dplyr::select(SST, year, geometry)-> SST_survey_domain

survey_area_basemap_km +
  geom_sf(data = SST_survey_domain, fill = "blue", alpha = 0.2)

# add in distance from shore - this is now your survey domain, with covariates
SST_survey_domain %>% 
  mutate(sf_dist_shore = as.numeric(st_distance(geometry, US_west_coast_proj_km))) -> survey_domain_cov
survey_domain_cov <- st_as_sf(survey_domain_cov, coords = c("X", "Y"), crs = UTM_zone_10_crs)



# because the survey_domain_polygon is constructed from a concave hull, some of the borders
# are a little coarse - and as a result some of the super nearshore data snuck back it.
# filter it out
survey_domain_cov %>% 
  filter(sf_dist_shore >= 0.5) -> survey_domain_cov


# extract the grid for the projection area - the grid dimensions are the same each year, so just select one year and drop fields besides geometry
survey_domain_cov %>% 
  filter(year == 2016) %>% 
  dplyr::select(geometry) -> survey_domain_cov_grid

A_gs_rf = fm_evaluator( inla_mesh_cutoff15_jsoes_domain, loc=st_coordinates(survey_domain_cov_grid))$proj$A
# A_gs_rf = fm_evaluator( inla_mesh_cutoff15_jsoes_domain, loc=st_coordinates(survey_domain_jsoes_cov_grid))$proj$A

# prep anisotropy input
# note that spatial list replaces the M0, M1, M2 objects 
spatial_list_rf <- make_anisotropy_spde( inla_mesh_cutoff15_jsoes_domain )
```

The trawl data for this taxon is visualized below:

```{r plot_distribtion, fig.cap = "Density of rockfishes caught in the PRS/PWCC surveys", fig.height = 8, fig.width = 12}
rockfishes_dist_plot <- plot_distribution_PRS_PWCC(data = rf, taxon_name = "Sebastes spp.")

rockfishes_dist_plot

ggsave(here::here("figures",  "rockfishes_dist_plot.png"), rockfishes_dist_plot,  
       height = 10, width = 16)
```

# Model selection process

## Step 1: Fit the base model with saturated fixed effects

We will first fit a model with all fixed effects included: temperature, distance from shore, and year.

- Year will be modeled as a factor
- Temperature and distance from shore will be modeled as linear predictors.

```{r fit_saturated_fixed_effects, cache = TRUE, include = FALSE}
# save.image("rf_model_selection.RData")
# load("rf_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "candidate_models", "saturated_fixed_effects_year_temp_dist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models", "saturated_fixed_effects_year_temp_dist")))

saturated_fixed_effects_year_temp_dist_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # "n_t" = max(rf$year - min(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
    # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, # "t_i" = rf$year - min(rf$year),
  "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  # "spatial_list" = spatial_list_rf
)


saturated_fixed_effects_year_temp_dist_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_t"=rep(0, max(rf$year - min(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  # "ln_tau_omega"=0, # "ln_tau_epsilon"=0,
  # "ln_kappa"=0, 
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0))
  # "epsilon_st_b"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=saturated_fixed_effects_year_temp_dist_Data$n_t)
  # 
)

saturated_fixed_effects_year_temp_dist_Obj = MakeADFun( data=saturated_fixed_effects_year_temp_dist_Data,
                                         parameters=saturated_fixed_effects_year_temp_dist_Params,
                                         # random= c("omega_s"),
                                         DLL = "saturated_fixed_effects_year_temp_dist")

# Optimize
saturated_fixed_effects_year_temp_dist_Opt = nlminb( start=saturated_fixed_effects_year_temp_dist_Obj$par, obj=saturated_fixed_effects_year_temp_dist_Obj$fn, grad=saturated_fixed_effects_year_temp_dist_Obj$gr )
# add getJointPrecision for index
# there are no random effects so turn off bias.correct and getJointPrecision
# saturated_fixed_effects_year_temp_dist_Opt$SD = sdreport( saturated_fixed_effects_year_temp_dist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
saturated_fixed_effects_year_temp_dist_Opt$SD = sdreport( saturated_fixed_effects_year_temp_dist_Obj, bias.correct=FALSE, getJointPrecision = FALSE )
saturated_fixed_effects_year_temp_dist_report = saturated_fixed_effects_year_temp_dist_Obj$report()

saturated_fixed_effects_year_temp_dist_output <- list(saturated_fixed_effects_year_temp_dist_Obj = saturated_fixed_effects_year_temp_dist_Obj,
                                       saturated_fixed_effects_year_temp_dist_Opt = saturated_fixed_effects_year_temp_dist_Opt,
                                       saturated_fixed_effects_year_temp_dist_report = saturated_fixed_effects_year_temp_dist_report)

save(saturated_fixed_effects_year_temp_dist_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "saturated_fixed_effects_year_temp_dist_output.rda"))
```


## Step 2: Tune the random effects, based on the model with saturated fixed effects

The primary random effects in this model are spatial and spatiotemporal random effects.

The options for the spatial random effects are as follows:

1. No spatial random fields
2. Spatial random fields (isotropic)
3. Spatial random fields (anisotropic)

The options for the spatiotemporal random effects are as follows:

1. No spatiotemporal random fields
2. Spatiotemporal random fields (isotropic, IID)
3. Spatiotemporal random fields (anisotropic, IID)
4. Spatiotemporal random fields (isotropic, AR1)
5. Spatiotemporal random fields (anisotropic, AR1)

However, it doesn't make ecological sense for spatial fields to be anisotropic and spatiotemporal fields to be isotropic, or vice versa (spatial fields isotropic and spatiotemporal fields anisotropic). So, the result of crossing these options with this constraint are the following models:

- 1-1 No spatial random fields, no spatiotemporal random fields
- 1-2 No spatial random fields, spatiotemporal random fields (isotropic, IID)
- 1-3 No spatial random fields, spatiotemporal random fields (anisotropic, IID)
- 1-4 No spatial random fields, spatiotemporal random fields (isotropic, AR1)
- 1-5 No spatial random fields, spatiotemporal random fields (anisotropic, AR1)

- 2-1 Spatial random fields (isotropic), no spatiotemporal random fields
- 2-2 Spatial random fields (isotropic), spatiotemporal random fields (isotropic, IID)
- 2-4 Spatial random fields (isotropic), spatiotemporal random fields (isotropic, AR1)

- 3-1 Spatial random fields (anisotropic), no spatiotemporal random fields
- 3-3 Spatial random fields (anisotropic), spatiotemporal random fields (anisotropic, IID)
- 3-5 Spatial random fields (anisotropic), spatiotemporal random fields (anisotropic, AR1)



The first cross (Model 1-1, no spatial random fields and no spatiotemporal random fields) is the saturated fixed effects model. Therefore, we will proceed with the ten other candidate random effect structures.

### Fit candidate models

#### Model 1-2: No spatial random fields, spatiotemporal random fields (isotropic, IID)

```{r model_1_2, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_2_spatial_off_spatiotemporal_IID_iso_missing_years.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_2_spatial_off_spatiotemporal_IID_iso_missing_years")))

model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Data = list(
  # number of years
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0, # "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0))
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Data$n_t)
  # 
)

model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Obj = MakeADFun( data=model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Data,
                                                             parameters=model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Params,
                                                             # random= c("omega_s"),
                                                             random= c("epsilon_st"),
                                                             DLL = "model1_2_spatial_off_spatiotemporal_IID_iso_missing_years")

# Optimize
model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Opt = nlminb( start=model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Obj$par, obj=model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Obj$fn, grad=model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Obj$gr )
# add getJointPrecision for index
model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Opt$SD = sdreport( model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_report = model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Obj$report()

model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_output <- list(model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Obj = model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Obj,
                                                           model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Opt = model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Opt,
                                                           model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_report = model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_report)

save(model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_output.rda"))
```


#### Model 1-3: No spatial random fields, spatiotemporal random fields (anisotropic, IID)

```{r model_1_3, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years")))

model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Data = list(
  # number of years
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0, # "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0))
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Data$n_t)
  # 
)

model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Obj = MakeADFun( data=model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Data,
                                                               parameters=model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Params,
                                                               # random= c("omega_s"),
                                                               random= c("epsilon_st"),
                                                               DLL = "model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years")

# Optimize
model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Opt = nlminb( start=model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Obj$par, obj=model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Obj$fn, grad=model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Obj$gr )
# add getJointPrecision for index
model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Opt$SD = sdreport( model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_report = model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Obj$report()

model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_output <- list(model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Obj = model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Obj,
                                                             model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Opt = model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Opt,
                                                             model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_report = model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_report)

save(model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_output.rda"))
```

#### Model 1-4: No spatial random fields, spatiotemporal random fields (isotropic, AR1)

```{r model_1_4, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years")))

model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Data = list(
  # number of years
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0, # "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0))
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Data$n_t)
  # 
)

model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Obj = MakeADFun( data=model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Data,
                                                             parameters=model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Params,
                                                             # random= c("omega_s"),
                                                             random= c("epsilon_st"),
                                                             DLL = "model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years")

# Optimize
model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Opt = nlminb( start=model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Obj$par, obj=model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Obj$fn, grad=model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Obj$gr )
# add getJointPrecision for index
model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Opt$SD = sdreport( model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_report = model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Obj$report()

model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_output <- list(model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Obj = model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Obj,
                                                           model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Opt = model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Opt,
                                                           model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_report = model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_report)

save(model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_output.rda"))
```

#### Model 1-5: No spatial random fields, spatiotemporal random fields (anisotropic, AR1)

```{r model_1_5, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years")))

model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Data = list(
  # number of years
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0, # "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0))
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Data$n_t)
  # 
)

model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Obj = MakeADFun( data=model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Data,
                                                               parameters=model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Params,
                                                               # random= c("omega_s"),
                                                               random= c("epsilon_st"),
                                                               DLL = "model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years")

# Optimize
model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Opt = nlminb( start=model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Obj$par, obj=model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Obj$fn, grad=model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Obj$gr )
# add getJointPrecision for index
model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Opt$SD = sdreport( model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_report = model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Obj$report()

model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_output <- list(model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Obj = model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Obj,
                                                             model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Opt = model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Opt,
                                                             model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_report = model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_report)

save(model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_output.rda"))
```


#### Model 2-1: Spatial random fields (isotropic), no spatiotemporal random fields


```{r model_2_1, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model2_1_spatial_on_spatiotemporal_off_missing_years.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model2_1_spatial_on_spatiotemporal_off_missing_years")))

model2_1_spatial_on_spatiotemporal_off_missing_years_Data = list(
  # number of years
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


model2_1_spatial_on_spatiotemporal_off_missing_years_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  # "ln_tau_epsilon"=0, 
  "ln_tau_omega"=0,
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0))
  # "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model2_1_spatial_on_spatiotemporal_off_missing_years_Data$n_t)
  # 
)

model2_1_spatial_on_spatiotemporal_off_missing_years_Obj = MakeADFun( data=model2_1_spatial_on_spatiotemporal_off_missing_years_Data,
                                                        parameters=model2_1_spatial_on_spatiotemporal_off_missing_years_Params,
                                                        random= c("omega_s"),
                                                        # random= c("epsilon_st"),
                                                        DLL = "model2_1_spatial_on_spatiotemporal_off_missing_years")

# Optimize
model2_1_spatial_on_spatiotemporal_off_missing_years_Opt = nlminb( start=model2_1_spatial_on_spatiotemporal_off_missing_years_Obj$par, obj=model2_1_spatial_on_spatiotemporal_off_missing_years_Obj$fn, grad=model2_1_spatial_on_spatiotemporal_off_missing_years_Obj$gr )
# add getJointPrecision for index
model2_1_spatial_on_spatiotemporal_off_missing_years_Opt$SD = sdreport( model2_1_spatial_on_spatiotemporal_off_missing_years_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model2_1_spatial_on_spatiotemporal_off_missing_years_report = model2_1_spatial_on_spatiotemporal_off_missing_years_Obj$report()

model2_1_spatial_on_spatiotemporal_off_missing_years_output <- list(model2_1_spatial_on_spatiotemporal_off_missing_years_Obj = model2_1_spatial_on_spatiotemporal_off_missing_years_Obj,
                                                      model2_1_spatial_on_spatiotemporal_off_missing_years_Opt = model2_1_spatial_on_spatiotemporal_off_missing_years_Opt,
                                                      model2_1_spatial_on_spatiotemporal_off_missing_years_report = model2_1_spatial_on_spatiotemporal_off_missing_years_report)

save(model2_1_spatial_on_spatiotemporal_off_missing_years_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model2_1_spatial_on_spatiotemporal_off_missing_years_output.rda"))
```


#### Model 2-2: Spatial random fields (isotropic), spatiotemporal random fields (isotropic, IID)


```{r model_2_2, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model2_2_spatial_on_spatiotemporal_IID_iso_missing_years.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model2_2_spatial_on_spatiotemporal_IID_iso_missing_years")))

model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Data = list(
  # number of years
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Data$n_t)
  # 
)

model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj = MakeADFun( data=model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Data,
                                                            parameters=model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Params,
                                                            random= c("omega_s", "epsilon_st"),
                                                            DLL = "model2_2_spatial_on_spatiotemporal_IID_iso_missing_years")

# Optimize
model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Opt = nlminb( start=model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj$par, obj=model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj$fn, grad=model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj$gr )
# add getJointPrecision for index
model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Opt$SD = sdreport( model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_report = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj$report()

model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_output <- list(model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj,
                                                          model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Opt = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Opt,
                                                          model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_report = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_report)

save(model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_output.rda"))
```


#### Model 2-4: Spatial random fields (isotropic), spatiotemporal random fields (isotropic, AR1)

```{r model_2_4, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years")))

model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Data = list(
  # number of years
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Data$n_t)
  # 
)

model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Obj = MakeADFun( data=model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Data,
                                                            parameters=model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Params,
                                                            random= c("omega_s", "epsilon_st"),
                                                            DLL = "model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years")

# Optimize
model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Opt = nlminb( start=model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Obj$par, obj=model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Obj$fn, grad=model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Obj$gr )
# add getJointPrecision for index
model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Opt$SD = sdreport( model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_report = model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Obj$report()

model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_output <- list(model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Obj = model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Obj,
                                                          model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Opt = model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_Opt,
                                                          model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_report = model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_report)

save(model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model2_4_spatial_on_spatiotemporal_AR1_iso_missing_years_output.rda"))
```

#### Model 3-1: Spatial random fields (anisotropic), no spatiotemporal random fields


```{r model_3_1, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model3_1_spatial_aniso_spatiotemporal_off_missing_years.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model3_1_spatial_aniso_spatiotemporal_off_missing_years")))

model3_1_spatial_aniso_spatiotemporal_off_missing_years_Data = list(
  # number of years
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


model3_1_spatial_aniso_spatiotemporal_off_missing_years_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  # "ln_tau_epsilon"=0, 
  "ln_tau_omega"=0,
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0))
  # "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model3_1_spatial_aniso_spatiotemporal_off_missing_years_Data$n_t)
  # 
)

model3_1_spatial_aniso_spatiotemporal_off_missing_years_Obj = MakeADFun( data=model3_1_spatial_aniso_spatiotemporal_off_missing_years_Data,
                                                           parameters=model3_1_spatial_aniso_spatiotemporal_off_missing_years_Params,
                                                           random= c("omega_s"),
                                                           # random= c("epsilon_st"),
                                                           DLL = "model3_1_spatial_aniso_spatiotemporal_off_missing_years")

# Optimize
model3_1_spatial_aniso_spatiotemporal_off_missing_years_Opt = nlminb( start=model3_1_spatial_aniso_spatiotemporal_off_missing_years_Obj$par, obj=model3_1_spatial_aniso_spatiotemporal_off_missing_years_Obj$fn, grad=model3_1_spatial_aniso_spatiotemporal_off_missing_years_Obj$gr )
# add getJointPrecision for index
model3_1_spatial_aniso_spatiotemporal_off_missing_years_Opt$SD = sdreport( model3_1_spatial_aniso_spatiotemporal_off_missing_years_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model3_1_spatial_aniso_spatiotemporal_off_missing_years_report = model3_1_spatial_aniso_spatiotemporal_off_missing_years_Obj$report()

model3_1_spatial_aniso_spatiotemporal_off_missing_years_output <- list(model3_1_spatial_aniso_spatiotemporal_off_missing_years_Obj = model3_1_spatial_aniso_spatiotemporal_off_missing_years_Obj,
                                                         model3_1_spatial_aniso_spatiotemporal_off_missing_years_Opt = model3_1_spatial_aniso_spatiotemporal_off_missing_years_Opt,
                                                         model3_1_spatial_aniso_spatiotemporal_off_missing_years_report = model3_1_spatial_aniso_spatiotemporal_off_missing_years_report)

save(model3_1_spatial_aniso_spatiotemporal_off_missing_years_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model3_1_spatial_aniso_spatiotemporal_off_missing_years_output.rda"))
```




#### Model 3-3: Spatial random fields (anisotropic), spatiotemporal random fields (anisotropic, IID)

```{r model_3_3, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years")))

model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Data = list(
  # number of years
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Data$n_t)
  # 
)

model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Obj = MakeADFun( data=model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Data,
                                                                 parameters=model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Params,
                                                                 random= c("omega_s", "epsilon_st"),
                                                                 DLL = "model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years")

# Optimize
model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Opt = nlminb( start=model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Obj$par, obj=model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Obj$fn, grad=model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Obj$gr )
# add getJointPrecision for index
model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Opt$SD = sdreport( model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_report = model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Obj$report()

model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_output <- list(model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Obj = model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Obj,
                                                               model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Opt = model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Opt,
                                                               model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_report = model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_report)

save(model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_output.rda"))
```



#### Model 3-5: Spatial random fields (anisotropic), spatiotemporal random fields (anisotropic, AR1)

```{r model_3_5, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1")))

model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Data = list(
  # number of years
  # "n_t" = length(unique(rf$year)),
  "n_t" = max(rf$year - min(rf$year))+1,
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix_rf_allyears,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, 
  # "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "t_i" = rf$year - min(rf$year),
  "year_factor_t_i" = c(0,1,2,3,4,5,6,7,8,-999,9,-999,10,11,12,13,14,15,16,-999,17),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  "spatial_list" = spatial_list_rf
)


model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Data$n_t)
  # 
)

model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Obj = MakeADFun( data=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Data,
                                                            parameters=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Params,
                                                            random= c("omega_s", "epsilon_st"),
                                                            DLL = "model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1")

# Optimize
model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Opt = nlminb( start=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Obj$par, obj=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Obj$fn, grad=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Obj$gr )
# add getJointPrecision for index
model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Opt$SD = sdreport( model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_report = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Obj$report()

model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_output <- list(model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Obj = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Obj,
                                                          model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Opt = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Opt,
                                                          model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_report = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_report)

save(model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_output.rda"))
```

### Compare candidate models

```{r random_effects_model_selection, cache = TRUE, include = FALSE}
# make sure that objects are loaded - need to do this if chunk outputs were cached
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models", "saturated_fixed_effects_year_temp_dist")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_2_spatial_off_spatiotemporal_IID_iso_missing_years")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model2_1_spatial_on_spatiotemporal_off_missing_years")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model2_2_spatial_on_spatiotemporal_IID_iso_missing_years")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model2_2_spatial_on_spatiotemporal_IID_iso_missing_years")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model3_1_spatial_aniso_spatiotemporal_off_missing_years")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "missing_years_candidate_models",  "model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1")))

# Calculate cAIC and AIC for each model

### Fixed effects
saturated_fixed_effects_year_temp_dist_cAIC <- cAIC(model_data = saturated_fixed_effects_year_temp_dist_Data,
                                                    model_obj = saturated_fixed_effects_year_temp_dist_Obj,  
                                                    model_parameters = saturated_fixed_effects_year_temp_dist_Params,
                                                    random = NULL,
                                                    DLL = "saturated_fixed_effects_year_temp_dist",
                                                    data_observation_names = "D_i")

saturated_fixed_effects_year_temp_dist_AIC <- 2*saturated_fixed_effects_year_temp_dist_Opt$objective + 2 * length(saturated_fixed_effects_year_temp_dist_Opt$par)

### Model 1-2
model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_cAIC <- cAIC(model_data = model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Data,
                                                         model_obj = model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Obj,  
                                                         model_parameters = model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Params,
                                                         random = "epsilon_st",
                                                         DLL = "model1_2_spatial_off_spatiotemporal_IID_iso_missing_years",
                                                         data_observation_names = "D_i")

model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_AIC <- 2*model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Opt$objective + 2 * length(model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_Opt$par)

### Model 1-3
model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_cAIC <- cAIC(model_data = model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Data,
                                                           model_obj = model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Obj,  
                                                           model_parameters = model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Params,
                                                           random = "epsilon_st",
                                                           DLL = "model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years",
                                                           data_observation_names = "D_i")

model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_AIC <- 2*model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Opt$objective + 2 * length(model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_Opt$par)

### Model 1-4
model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_cAIC <- cAIC(model_data = model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Data,
                                                         model_obj = model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Obj,  
                                                         model_parameters = model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Params,
                                                         random = "epsilon_st",
                                                         DLL = "model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years",
                                                         data_observation_names = "D_i")

model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_AIC <- 2*model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Opt$objective + 2 * length(model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_Opt$par)

### Model 1-5
model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_cAIC <- cAIC(model_data = model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Data,
                                                           model_obj = model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Obj,  
                                                           model_parameters = model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Params,
                                                           random = "epsilon_st",
                                                           DLL = "model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years",
                                                           data_observation_names = "D_i")

model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_AIC <- 2*model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Opt$objective + 2 * length(model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_Opt$par)

### Model 2-1
model2_1_spatial_on_spatiotemporal_off_missing_years_cAIC <- cAIC(model_data = model2_1_spatial_on_spatiotemporal_off_missing_years_Data,
                                                    model_obj = model2_1_spatial_on_spatiotemporal_off_missing_years_Obj,  
                                                    model_parameters = model2_1_spatial_on_spatiotemporal_off_missing_years_Params,
                                                    random = "omega_s",
                                                    DLL = "model2_1_spatial_on_spatiotemporal_off_missing_years",
                                                    data_observation_names = "D_i")

model2_1_spatial_on_spatiotemporal_off_missing_years_AIC <- 2*model2_1_spatial_on_spatiotemporal_off_missing_years_Opt$objective + 2 * length(model2_1_spatial_on_spatiotemporal_off_missing_years_Opt$par)

### Model 2-2
model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_cAIC <- cAIC(model_data = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Data,
                                                        model_obj = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj,  
                                                        model_parameters = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Params,
                                                        random = "omega_s",
                                                        DLL = "model2_2_spatial_on_spatiotemporal_IID_iso_missing_years",
                                                        data_observation_names = "D_i")

model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_AIC <- 2*model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Opt$objective + 2 * length(model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Opt$par)

### Model 2-4
model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_cAIC <- cAIC(model_data = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Data,
                                                        model_obj = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Obj,  
                                                        model_parameters = model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Params,
                                                        random = "omega_s",
                                                        DLL = "model2_2_spatial_on_spatiotemporal_IID_iso_missing_years",
                                                        data_observation_names = "D_i")

model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_AIC <- 2*model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Opt$objective + 2 * length(model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_Opt$par)

### Model 3-1
model3_1_spatial_aniso_spatiotemporal_off_missing_years_cAIC <- cAIC(model_data = model3_1_spatial_aniso_spatiotemporal_off_missing_years_Data,
                                                       model_obj = model3_1_spatial_aniso_spatiotemporal_off_missing_years_Obj,  
                                                       model_parameters = model3_1_spatial_aniso_spatiotemporal_off_missing_years_Params,
                                                       random = c("omega_s"),
                                                       DLL = "model3_1_spatial_aniso_spatiotemporal_off_missing_years",
                                                       data_observation_names = "D_i")

model3_1_spatial_aniso_spatiotemporal_off_missing_years_AIC <- 2*model3_1_spatial_aniso_spatiotemporal_off_missing_years_Opt$objective + 2 * length(model3_1_spatial_aniso_spatiotemporal_off_missing_years_Opt$par)

### Model 3-3
model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_cAIC <- cAIC(model_data = model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Data,
                                                             model_obj = model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Obj,  
                                                             model_parameters = model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Params,
                                                             random = c("omega_s", "epsilon_st"),
                                                             DLL = "model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years",
                                                             data_observation_names = "D_i")

model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_AIC <- 2*model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Opt$objective + 2 * length(model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_Opt$par)

### Model 3-5
model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_cAIC <- cAIC(model_data = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Data,
                                                             model_obj = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Obj,  
                                                             model_parameters = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Params,
                                                             random = c("omega_s", "epsilon_st"),
                                                             DLL = "model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1",
                                                             data_observation_names = "D_i")

model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_AIC <- 2*model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Opt$objective + 2 * length(model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_Opt$par)

# summarize this all in a table
random_effects_model_comparison_table <- data.frame(model = c("Spatial OFF + Spatiotemporal OFF",
                                                              "Spatial OFF + Spatiotemporal IID (isotropic)",
                                                              "Spatial OFF + Spatiotemporal IID (anisotropic)",
                                                              "Spatial OFF + Spatiotemporal AR1 (isotropic)",
                                                              "Spatial OFF + Spatiotemporal AR1 (anisotropic)",
                                                              "Spatial ON + Spatiotemporal OFF (isotropic)",
                                                              "Spatial ON + Spatiotemporal IID (isotropic)",
                                                              "Spatial ON + Spatiotemporal AR1 (isotropic)",
                                                              "Spatial ON + Spatiotemporal OFF (anisotropic)", 
                                                              "Spatial ON + Spatiotemporal IID (anisotropic)", 
                                                              "Spatial ON + Spatiotemporal AR1 (anisotropic)"),
                                                    AIC = c(saturated_fixed_effects_year_temp_dist_AIC,
                                                            model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_AIC,
                                                            model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_AIC,
                                                            model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_AIC,
                                                            model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_AIC,
                                                            model2_1_spatial_on_spatiotemporal_off_missing_years_AIC,
                                                            model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_AIC,
                                                            model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_AIC,
                                                            model3_1_spatial_aniso_spatiotemporal_off_missing_years_AIC,
                                                            model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_AIC,
                                                            model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_AIC),
                                                    cAIC = c(NA,
                                                             model1_2_spatial_off_spatiotemporal_IID_iso_missing_years_cAIC,
                                                             model1_3_spatial_off_spatiotemporal_IID_aniso_missing_years_cAIC,
                                                             model1_4_spatial_off_spatiotemporal_AR1_iso_missing_years_cAIC,
                                                             model1_5_spatial_off_spatiotemporal_AR1_aniso_missing_years_cAIC,
                                                             model2_1_spatial_on_spatiotemporal_off_missing_years_cAIC,
                                                             model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_cAIC,
                                                             model2_2_spatial_on_spatiotemporal_IID_iso_missing_years_cAIC,
                                                             model3_1_spatial_aniso_spatiotemporal_off_missing_years_cAIC,
                                                             model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_cAIC,
                                                             model3_5_spatial_aniso_spatiotemporal_AR1_aniso_missing_years_1_cAIC))
```

```{r random_effects_model_comparison_table}
random_effects_model_comparison_table %>% 
  arrange(cAIC) %>% 
  mutate(AIC = round(AIC, 2)) %>% 
  mutate(cAIC = round(cAIC, 2)) %>% 
  kbl(caption = "AIC and cAIC scores for the random effect structures.") %>% 
  kable_styling(latex_options = "HOLD_position")
```

**Based on both cAIC and AIC, the model with Spatial ON + Spatiotemporal AR1 (anisotropic) is the best structure for random effects. We will therefore use this model structure to explore different fixed effects structures.**

## Step 3: Tune fixed effects based on random effects selected in step 2

We have three covariates that we are exploring:

- Year
- Temperature
- Distance from shore

Year we will only model as a factor representing a year-wide intercept for the full survey region, without an autoregressive component. This is because we do not expect the abundance of salmon to be an autoregressive process (successive run years are not related in terms of abundance, given the salmon life history).

Temperature we can model as a linear fixed effect or a smooth spline (technically a random effect).

Distance from shore we can model as a linear fixed effect or a smooth spline (technically a random effect).

Therefore, our options for fixed effects are as follows:

- Year, as a factor (on/off)
- Temperature (off/linear/spline)
- Distance from shore (off/linear/spline)


Crossing these options (2 x 3 x 3) results in the following 18 candidate fixed effect structures:

1. Year (factor) + temperature (linear) + distance from shore (linear) - this is the base model that we used to test random effect structures.
2. Year (factor) + temperature (smooth spline) + distance from shore (smooth spline)
3. Year (factor) + temperature (linear) + distance from shore (smooth spline)
4. Year (factor) + temperature (smooth spline) + distance from shore (linear)
5. Year (factor) + temperature (linear)
6. Year (factor) + distance from shore (linear)
7. Year (factor) + temperature (smooth spline)
8. Year (factor) + distance from shore (smooth spline)
9. Year (factor)
10. Distance from shore (smooth spline) + temperature (smooth spline)
11. Distance from shore (smooth spline) + temperature (linear)
12. Distance from shore (linear) + temperature (smooth spline)
13. Distance from shore (linear) + temperature (linear)
14. Distance from shore (smooth spline)
15. Distance from shore (linear)
16. Temperature (smooth spline)
17. Temperature (linear)
18. No fixed effects


### Fit candidate models

#### Model 2: Year (factor) + temperature (smooth spline) + distance from shore (smooth spline)

```{r model_2, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(SST_scaled) + s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                             dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total ~ s(SST_scaled) + s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL,
      newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("rf_model_selection.RData")
# load("rf_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_stemp_sdist_year1.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_sdist_year1")))

rf_bestfitRE_stemp_sdist_year1_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
    # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_stemp_sdist_year1_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_stemp_sdist_year1_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_stemp_sdist_year1_Obj = MakeADFun( data=rf_bestfitRE_stemp_sdist_year1_Data,
                                                            parameters=rf_bestfitRE_stemp_sdist_year1_Params,
                                                            random= c("omega_s", "epsilon_st", "b_smooth"),
                                                            DLL = "rf_bestfitRE_stemp_sdist_year1")

# Optimize
rf_bestfitRE_stemp_sdist_year1_Opt = nlminb( start=rf_bestfitRE_stemp_sdist_year1_Obj$par, obj=rf_bestfitRE_stemp_sdist_year1_Obj$fn, grad=rf_bestfitRE_stemp_sdist_year1_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_stemp_sdist_year1_Opt$SD = sdreport( rf_bestfitRE_stemp_sdist_year1_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_stemp_sdist_year1_report = rf_bestfitRE_stemp_sdist_year1_Obj$report()

rf_bestfitRE_stemp_sdist_year1_output <- list(rf_bestfitRE_stemp_sdist_year1_Obj = rf_bestfitRE_stemp_sdist_year1_Obj,
                                                          rf_bestfitRE_stemp_sdist_year1_Opt = rf_bestfitRE_stemp_sdist_year1_Opt,
                                                          rf_bestfitRE_stemp_sdist_year1_report = rf_bestfitRE_stemp_sdist_year1_report)

save(rf_bestfitRE_stemp_sdist_year1_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_sdist_year1_output.rda"))
```

#### Model 3: Year (factor) + temperature (linear)

```{r model_3, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_year1.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_year1")))

rf_bestfitRE_temp_year1_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = rf$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


rf_bestfitRE_temp_year1_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_temp_year1_Data$n_t)
  # 
)

rf_bestfitRE_temp_year1_Obj = MakeADFun( data=rf_bestfitRE_temp_year1_Data,
                                                                 parameters=rf_bestfitRE_temp_year1_Params,
                                                                 random= c("omega_s", "epsilon_st"),
                                                                 DLL = "rf_bestfitRE_temp_year1")

# Optimize
rf_bestfitRE_temp_year1_Opt = nlminb( start=rf_bestfitRE_temp_year1_Obj$par, obj=rf_bestfitRE_temp_year1_Obj$fn, grad=rf_bestfitRE_temp_year1_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_temp_year1_Opt$SD = sdreport( rf_bestfitRE_temp_year1_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_temp_year1_report = rf_bestfitRE_temp_year1_Obj$report()

rf_bestfitRE_temp_year1_output <- list(rf_bestfitRE_temp_year1_Obj = rf_bestfitRE_temp_year1_Obj,
                                                               rf_bestfitRE_temp_year1_Opt = rf_bestfitRE_temp_year1_Opt,
                                                               rf_bestfitRE_temp_year1_report = rf_bestfitRE_temp_year1_report)

save(rf_bestfitRE_temp_year1_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "rf_bestfitRE_temp_year1_output.rda"))
```

#### Model 4: Year (factor) + distance from shore (linear)

```{r model_4, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_dist_year1.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_dist_year1")))

rf_bestfitRE_dist_year1_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  # "temp_i" = rf$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


rf_bestfitRE_dist_year1_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_dist_year1_Data$n_t)
  # 
)

rf_bestfitRE_dist_year1_Obj = MakeADFun( data=rf_bestfitRE_dist_year1_Data,
                                           parameters=rf_bestfitRE_dist_year1_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "rf_bestfitRE_dist_year1")

# Optimize
rf_bestfitRE_dist_year1_Opt = nlminb( start=rf_bestfitRE_dist_year1_Obj$par, obj=rf_bestfitRE_dist_year1_Obj$fn, grad=rf_bestfitRE_dist_year1_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_dist_year1_Opt$SD = sdreport( rf_bestfitRE_dist_year1_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_dist_year1_report = rf_bestfitRE_dist_year1_Obj$report()

rf_bestfitRE_dist_year1_output <- list(rf_bestfitRE_dist_year1_Obj = rf_bestfitRE_dist_year1_Obj,
                                         rf_bestfitRE_dist_year1_Opt = rf_bestfitRE_dist_year1_Opt,
                                         rf_bestfitRE_dist_year1_report = rf_bestfitRE_dist_year1_report)

save(rf_bestfitRE_dist_year1_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "rf_bestfitRE_dist_year1_output.rda"))
```

#### Model 5: Year (factor) + temperature (smooth spline)

```{r model_5, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(SST_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix))

sm_predict <- parse_smoothers(formula = "total ~ s(SST_scaled)", data = as.data.frame(rf), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_stemp_year1.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_year1")))

rf_bestfitRE_stemp_year1_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = rf$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_stemp_year1_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_stemp_year1_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_stemp_year1_Obj = MakeADFun( data=rf_bestfitRE_stemp_year1_Data,
                                                  parameters=rf_bestfitRE_stemp_year1_Params,
                                                  random= c("omega_s", "epsilon_st", "b_smooth"),
                                                  DLL = "rf_bestfitRE_stemp_year1")

# Optimize
rf_bestfitRE_stemp_year1_Opt = nlminb( start=rf_bestfitRE_stemp_year1_Obj$par, obj=rf_bestfitRE_stemp_year1_Obj$fn, grad=rf_bestfitRE_stemp_year1_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_stemp_year1_Opt$SD = sdreport( rf_bestfitRE_stemp_year1_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_stemp_year1_report = rf_bestfitRE_stemp_year1_Obj$report()

rf_bestfitRE_stemp_year1_output <- list(rf_bestfitRE_stemp_year1_Obj = rf_bestfitRE_stemp_year1_Obj,
                                                rf_bestfitRE_stemp_year1_Opt = rf_bestfitRE_stemp_year1_Opt,
                                                rf_bestfitRE_stemp_year1_report = rf_bestfitRE_stemp_year1_report)

save(rf_bestfitRE_stemp_year1_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_year1_output.rda"))
```

#### Model 6: Year (factor) + s(dist)

```{r model_6, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(
  dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total ~ s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


save.image("rf_model_selection.RData")
# load("rf_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_sdist_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_sdist_year")))

rf_bestfitRE_sdist_year_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  # "temp_i" = rf$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_sdist_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_sdist_year_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_sdist_year_Obj = MakeADFun( data=rf_bestfitRE_sdist_year_Data,
                                                  parameters=rf_bestfitRE_sdist_year_Params,
                                                  random= c("omega_s", "epsilon_st", "b_smooth"),
                                                  DLL = "rf_bestfitRE_sdist_year")

# Optimize
rf_bestfitRE_sdist_year_Opt = nlminb( start=rf_bestfitRE_sdist_year_Obj$par, obj=rf_bestfitRE_sdist_year_Obj$fn, grad=rf_bestfitRE_sdist_year_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_sdist_year_Opt$SD = sdreport( rf_bestfitRE_sdist_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_sdist_year_report = rf_bestfitRE_sdist_year_Obj$report()

rf_bestfitRE_sdist_year_output <- list(rf_bestfitRE_sdist_year_Obj = rf_bestfitRE_sdist_year_Obj,
                                                rf_bestfitRE_sdist_year_Opt = rf_bestfitRE_sdist_year_Opt,
                                                rf_bestfitRE_sdist_year_report = rf_bestfitRE_sdist_year_report)

save(rf_bestfitRE_sdist_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_sdist_year_output.rda"))
```

#### Model 7: Year (factor)

```{r model_7, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_year")))

rf_bestfitRE_year_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  # "temp_i" = rf$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = rf$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


rf_bestfitRE_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_year_Data$n_t)
  # 
)

rf_bestfitRE_year_Obj = MakeADFun( data=rf_bestfitRE_year_Data,
                                           parameters=rf_bestfitRE_year_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "rf_bestfitRE_year")

# Optimize
rf_bestfitRE_year_Opt = nlminb( start=rf_bestfitRE_year_Obj$par, obj=rf_bestfitRE_year_Obj$fn, grad=rf_bestfitRE_year_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_year_Opt$SD = sdreport( rf_bestfitRE_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_year_report = rf_bestfitRE_year_Obj$report()

rf_bestfitRE_year_output <- list(rf_bestfitRE_year_Obj = rf_bestfitRE_year_Obj,
                                         rf_bestfitRE_year_Opt = rf_bestfitRE_year_Opt,
                                         rf_bestfitRE_year_report = rf_bestfitRE_year_report)

save(rf_bestfitRE_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "rf_bestfitRE_year_output.rda"))
```

#### Model 8: No fixed effects

```{r model_8, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_noFE.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_noFE")))

rf_bestfitRE_noFE_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  # "temp_i" = rf$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = rf$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


rf_bestfitRE_noFE_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_noFE_Data$n_t)
  # 
)

rf_bestfitRE_noFE_Obj = MakeADFun( data=rf_bestfitRE_noFE_Data,
                                           parameters=rf_bestfitRE_noFE_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "rf_bestfitRE_noFE")

# Optimize
rf_bestfitRE_noFE_Opt = nlminb( start=rf_bestfitRE_noFE_Obj$par, obj=rf_bestfitRE_noFE_Obj$fn, grad=rf_bestfitRE_noFE_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_noFE_Opt$SD = sdreport( rf_bestfitRE_noFE_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_noFE_report = rf_bestfitRE_noFE_Obj$report()

rf_bestfitRE_noFE_output <- list(rf_bestfitRE_noFE_Obj = rf_bestfitRE_noFE_Obj,
                                         rf_bestfitRE_noFE_Opt = rf_bestfitRE_noFE_Opt,
                                         rf_bestfitRE_noFE_report = rf_bestfitRE_noFE_report)

save(rf_bestfitRE_noFE_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "rf_bestfitRE_noFE_output.rda"))
```

#### Model 9: temperature (smooth spline) + distance from shore (smooth spline)

```{r model_9, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(SST_scaled) + s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total ~ s(SST_scaled) + s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("rf_model_selection.RData")
# load("rf_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_stemp_sdist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_sdist")))

rf_bestfitRE_stemp_sdist_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_stemp_sdist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_stemp_sdist_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_stemp_sdist_Obj = MakeADFun( data=rf_bestfitRE_stemp_sdist_Data,
                                                  parameters=rf_bestfitRE_stemp_sdist_Params,
                                                  random= c("omega_s", "epsilon_st", "b_smooth"),
                                                  DLL = "rf_bestfitRE_stemp_sdist")

# Optimize
rf_bestfitRE_stemp_sdist_Opt = nlminb( start=rf_bestfitRE_stemp_sdist_Obj$par, obj=rf_bestfitRE_stemp_sdist_Obj$fn, grad=rf_bestfitRE_stemp_sdist_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_stemp_sdist_Opt$SD = sdreport( rf_bestfitRE_stemp_sdist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_stemp_sdist_report = rf_bestfitRE_stemp_sdist_Obj$report()

rf_bestfitRE_stemp_sdist_output <- list(rf_bestfitRE_stemp_sdist_Obj = rf_bestfitRE_stemp_sdist_Obj,
                                                rf_bestfitRE_stemp_sdist_Opt = rf_bestfitRE_stemp_sdist_Opt,
                                                rf_bestfitRE_stemp_sdist_report = rf_bestfitRE_stemp_sdist_report)

save(rf_bestfitRE_stemp_sdist_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_sdist_output.rda"))
```

#### Model 10: temperature (linear)

```{r model_10, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp")))

rf_bestfitRE_temp_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = rf$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


rf_bestfitRE_temp_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_temp_Data$n_t)
  # 
)

rf_bestfitRE_temp_Obj = MakeADFun( data=rf_bestfitRE_temp_Data,
                                           parameters=rf_bestfitRE_temp_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "rf_bestfitRE_temp")

# Optimize
rf_bestfitRE_temp_Opt = nlminb( start=rf_bestfitRE_temp_Obj$par, obj=rf_bestfitRE_temp_Obj$fn, grad=rf_bestfitRE_temp_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_temp_Opt$SD = sdreport( rf_bestfitRE_temp_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_temp_report = rf_bestfitRE_temp_Obj$report()

rf_bestfitRE_temp_output <- list(rf_bestfitRE_temp_Obj = rf_bestfitRE_temp_Obj,
                                         rf_bestfitRE_temp_Opt = rf_bestfitRE_temp_Opt,
                                         rf_bestfitRE_temp_report = rf_bestfitRE_temp_report)

save(rf_bestfitRE_temp_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "rf_bestfitRE_temp_output.rda"))
```

#### Model 11: distance from shore (linear)

```{r model_11, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_dist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_dist")))

rf_bestfitRE_dist_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  # "temp_i" = rf$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


rf_bestfitRE_dist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_dist_Data$n_t)
  # 
)

rf_bestfitRE_dist_Obj = MakeADFun( data=rf_bestfitRE_dist_Data,
                                           parameters=rf_bestfitRE_dist_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "rf_bestfitRE_dist")

# Optimize
rf_bestfitRE_dist_Opt = nlminb( start=rf_bestfitRE_dist_Obj$par, obj=rf_bestfitRE_dist_Obj$fn, grad=rf_bestfitRE_dist_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_dist_Opt$SD = sdreport( rf_bestfitRE_dist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_dist_report = rf_bestfitRE_dist_Obj$report()

rf_bestfitRE_dist_output <- list(rf_bestfitRE_dist_Obj = rf_bestfitRE_dist_Obj,
                                         rf_bestfitRE_dist_Opt = rf_bestfitRE_dist_Opt,
                                         rf_bestfitRE_dist_report = rf_bestfitRE_dist_report)

save(rf_bestfitRE_dist_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "rf_bestfitRE_dist_output.rda"))
```

#### Model 12: temperature (smooth spline)

```{r model_12, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(SST_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix))

sm_predict <- parse_smoothers(formula = "total ~ s(SST_scaled)", data = as.data.frame(rf), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_stemp.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp")))

rf_bestfitRE_stemp_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = rf$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_stemp_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_stemp_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_stemp_Obj = MakeADFun( data=rf_bestfitRE_stemp_Data,
                                            parameters=rf_bestfitRE_stemp_Params,
                                            random= c("omega_s", "epsilon_st", "b_smooth"),
                                            DLL = "rf_bestfitRE_stemp")

# Optimize
rf_bestfitRE_stemp_Opt = nlminb( start=rf_bestfitRE_stemp_Obj$par, obj=rf_bestfitRE_stemp_Obj$fn, grad=rf_bestfitRE_stemp_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_stemp_Opt$SD = sdreport( rf_bestfitRE_stemp_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_stemp_report = rf_bestfitRE_stemp_Obj$report()

rf_bestfitRE_stemp_output <- list(rf_bestfitRE_stemp_Obj = rf_bestfitRE_stemp_Obj,
                                          rf_bestfitRE_stemp_Opt = rf_bestfitRE_stemp_Opt,
                                          rf_bestfitRE_stemp_report = rf_bestfitRE_stemp_report)

save(rf_bestfitRE_stemp_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_output.rda"))
```

#### Model 13: s(dist)

```{r model_13, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(
  dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total ~ s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("rf_model_selection.RData")
# load("rf_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_sdist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_sdist")))

rf_bestfitRE_sdist_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  # "temp_i" = rf$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_sdist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_sdist_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_sdist_Obj = MakeADFun( data=rf_bestfitRE_sdist_Data,
                                            parameters=rf_bestfitRE_sdist_Params,
                                            random= c("omega_s", "epsilon_st", "b_smooth"),
                                            DLL = "rf_bestfitRE_sdist")

# Optimize
rf_bestfitRE_sdist_Opt = nlminb( start=rf_bestfitRE_sdist_Obj$par, obj=rf_bestfitRE_sdist_Obj$fn, grad=rf_bestfitRE_sdist_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_sdist_Opt$SD = sdreport( rf_bestfitRE_sdist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_sdist_report = rf_bestfitRE_sdist_Obj$report()

rf_bestfitRE_sdist_output <- list(rf_bestfitRE_sdist_Obj = rf_bestfitRE_sdist_Obj,
                                          rf_bestfitRE_sdist_Opt = rf_bestfitRE_sdist_Opt,
                                          rf_bestfitRE_sdist_report = rf_bestfitRE_sdist_report)

save(rf_bestfitRE_sdist_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_sdist_output.rda"))
```

#### Model 14: temperature (linear) + distance from shore (linear)

```{r model_14, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_dist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_dist")))

rf_bestfitRE_temp_dist_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf
)


rf_bestfitRE_temp_dist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_temp_dist_Data$n_t)
  # 
)

rf_bestfitRE_temp_dist_Obj = MakeADFun( data=rf_bestfitRE_temp_dist_Data,
                                                            parameters=rf_bestfitRE_temp_dist_Params,
                                                            random= c("omega_s", "epsilon_st"),
                                                            DLL = "rf_bestfitRE_temp_dist")

# Optimize
rf_bestfitRE_temp_dist_Opt = nlminb( start=rf_bestfitRE_temp_dist_Obj$par, obj=rf_bestfitRE_temp_dist_Obj$fn, grad=rf_bestfitRE_temp_dist_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_temp_dist_Opt$SD = sdreport( rf_bestfitRE_temp_dist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_temp_dist_report = rf_bestfitRE_temp_dist_Obj$report()

rf_bestfitRE_temp_dist_output <- list(rf_bestfitRE_temp_dist_Obj = rf_bestfitRE_temp_dist_Obj,
                                                          rf_bestfitRE_temp_dist_Opt = rf_bestfitRE_temp_dist_Opt,
                                                          rf_bestfitRE_temp_dist_report = rf_bestfitRE_temp_dist_report)

save(rf_bestfitRE_temp_dist_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_candidate_models", "rf_bestfitRE_temp_dist_output.rda"))
```

#### Model 15: Year (factor) + temperature (smooth spline) + distance from shore (linear)

```{r model_15, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(SST_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total ~ s(SST_scaled)", data = as.data.frame(rf), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("rf_model_selection.RData")
# load("rf_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_stemp_dist_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_dist_year")))

rf_bestfitRE_stemp_dist_year_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_stemp_dist_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_stemp_dist_year_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_stemp_dist_year_Obj = MakeADFun( data=rf_bestfitRE_stemp_dist_year_Data,
                                                  parameters=rf_bestfitRE_stemp_dist_year_Params,
                                                  random= c("omega_s", "epsilon_st", "b_smooth"),
                                                  DLL = "rf_bestfitRE_stemp_dist_year")

# Optimize
rf_bestfitRE_stemp_dist_year_Opt = nlminb( start=rf_bestfitRE_stemp_dist_year_Obj$par, obj=rf_bestfitRE_stemp_dist_year_Obj$fn, grad=rf_bestfitRE_stemp_dist_year_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_stemp_dist_year_Opt$SD = sdreport( rf_bestfitRE_stemp_dist_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_stemp_dist_year_report = rf_bestfitRE_stemp_dist_year_Obj$report()

rf_bestfitRE_stemp_dist_year_output <- list(rf_bestfitRE_stemp_dist_year_Obj = rf_bestfitRE_stemp_dist_year_Obj,
                                                rf_bestfitRE_stemp_dist_year_Opt = rf_bestfitRE_stemp_dist_year_Opt,
                                                rf_bestfitRE_stemp_dist_year_report = rf_bestfitRE_stemp_dist_year_report)

save(rf_bestfitRE_stemp_dist_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_dist_year_output.rda"))
```

#### Model 16: Year (factor) + temperature (linear) + distance from shore (smooth spline)

```{r model_16, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total ~ s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("rf_model_selection.RData")
# load("rf_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_temp_sdist_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_sdist_year")))

rf_bestfitRE_temp_sdist_year_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_temp_sdist_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_temp_sdist_year_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_temp_sdist_year_Obj = MakeADFun( data=rf_bestfitRE_temp_sdist_year_Data,
                                                 parameters=rf_bestfitRE_temp_sdist_year_Params,
                                                 random= c("omega_s", "epsilon_st", "b_smooth"),
                                                 DLL = "rf_bestfitRE_temp_sdist_year")

# Optimize
rf_bestfitRE_temp_sdist_year_Opt = nlminb( start=rf_bestfitRE_temp_sdist_year_Obj$par, obj=rf_bestfitRE_temp_sdist_year_Obj$fn, grad=rf_bestfitRE_temp_sdist_year_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_temp_sdist_year_Opt$SD = sdreport( rf_bestfitRE_temp_sdist_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_temp_sdist_year_report = rf_bestfitRE_temp_sdist_year_Obj$report()

rf_bestfitRE_temp_sdist_year_output <- list(rf_bestfitRE_temp_sdist_year_Obj = rf_bestfitRE_temp_sdist_year_Obj,
                                               rf_bestfitRE_temp_sdist_year_Opt = rf_bestfitRE_temp_sdist_year_Opt,
                                               rf_bestfitRE_temp_sdist_year_report = rf_bestfitRE_temp_sdist_year_report)

save(rf_bestfitRE_temp_sdist_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_sdist_year_output.rda"))
```

#### Model 17: temperature (smooth spline) + distance from shore (linear)

```{r model_17, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(SST_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total ~ s(SST_scaled)", data = as.data.frame(rf), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("rf_model_selection.RData")
# load("rf_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_stemp_dist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_dist")))

rf_bestfitRE_stemp_dist_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_stemp_dist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(rf$year))),
  # "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_stemp_dist_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_stemp_dist_Obj = MakeADFun( data=rf_bestfitRE_stemp_dist_Data,
                                             parameters=rf_bestfitRE_stemp_dist_Params,
                                             random= c("omega_s", "epsilon_st", "b_smooth"),
                                             DLL = "rf_bestfitRE_stemp_dist")

# Optimize
rf_bestfitRE_stemp_dist_Opt = nlminb( start=rf_bestfitRE_stemp_dist_Obj$par, obj=rf_bestfitRE_stemp_dist_Obj$fn, grad=rf_bestfitRE_stemp_dist_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_stemp_dist_Opt$SD = sdreport( rf_bestfitRE_stemp_dist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_stemp_dist_report = rf_bestfitRE_stemp_dist_Obj$report()

rf_bestfitRE_stemp_dist_output <- list(rf_bestfitRE_stemp_dist_Obj = rf_bestfitRE_stemp_dist_Obj,
                                           rf_bestfitRE_stemp_dist_Opt = rf_bestfitRE_stemp_dist_Opt,
                                           rf_bestfitRE_stemp_dist_report = rf_bestfitRE_stemp_dist_report)

save(rf_bestfitRE_stemp_dist_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_dist_output.rda"))
```

#### Model 18: temperature (linear) + distance from shore (smooth spline)

```{r model_18, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total ~ s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total ~ s(dist_shore_scaled)", data = as.data.frame(rf), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("rf_model_selection.RData")
# load("rf_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models",  "rf_bestfitRE_temp_sdist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_sdist")))

rf_bestfitRE_temp_sdist_Data = list(
  # number of years
  "n_t" = length(unique(rf$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(rf$total)),
  # temperature data
  "temp_i" = rf$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = rf$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # rockfish data
  "D_i"= rf$total, "t_i" = as.integer(dplyr::dense_rank(rf$year - min(rf$year))-1),
  "A_gs"=A_gs_rf,
  "A_is"=A_is_rf,
  # "M0_b"=PRS_PWCC_spde$c0, "M1_b"=PRS_PWCC_spde$g1, "M2_b"=PRS_PWCC_spde$g2,
  "spatial_list" = spatial_list_rf,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

rf_bestfitRE_temp_sdist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(rf$year))),
  "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(PRS_PWCC_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(PRS_PWCC_spde$c0), ncol=rf_bestfitRE_temp_sdist_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

rf_bestfitRE_temp_sdist_Obj = MakeADFun( data=rf_bestfitRE_temp_sdist_Data,
                                             parameters=rf_bestfitRE_temp_sdist_Params,
                                             random= c("omega_s", "epsilon_st", "b_smooth"),
                                             DLL = "rf_bestfitRE_temp_sdist")

# Optimize
rf_bestfitRE_temp_sdist_Opt = nlminb( start=rf_bestfitRE_temp_sdist_Obj$par, obj=rf_bestfitRE_temp_sdist_Obj$fn, grad=rf_bestfitRE_temp_sdist_Obj$gr )
# add getJointPrecision for index
rf_bestfitRE_temp_sdist_Opt$SD = sdreport( rf_bestfitRE_temp_sdist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
rf_bestfitRE_temp_sdist_report = rf_bestfitRE_temp_sdist_Obj$report()

rf_bestfitRE_temp_sdist_output <- list(rf_bestfitRE_temp_sdist_Obj = rf_bestfitRE_temp_sdist_Obj,
                                           rf_bestfitRE_temp_sdist_Opt = rf_bestfitRE_temp_sdist_Opt,
                                           rf_bestfitRE_temp_sdist_report = rf_bestfitRE_temp_sdist_report)

save(rf_bestfitRE_temp_sdist_output,
     file = here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_sdist_output.rda"))
```



### Compare candidate models

```{r}
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_sdist_year1")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_year1")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_dist_year1")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_year1")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_sdist_year")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_year")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_dist")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_sdist")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_dist")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_sdist")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_dist_year")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_sdist_year")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_temp_sdist")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_stemp_dist")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "rf_fixed_effects_candidate_models", "rf_bestfitRE_noFE")))
```


```{r, cache = TRUE, include = FALSE}
rf_bestfitRE_stemp_dist_year_cAIC <- cAIC(model_data = rf_bestfitRE_stemp_dist_year_Data,
                                              model_obj = rf_bestfitRE_stemp_dist_year_Obj,  
                                              model_parameters = rf_bestfitRE_stemp_dist_year_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "rf_bestfitRE_stemp_dist_year",
                                              data_observation_names = "D_i")

rf_bestfitRE_stemp_dist_year_AIC <- 2*rf_bestfitRE_stemp_dist_year_Opt$objective + 2 * length(rf_bestfitRE_stemp_dist_year_Opt$par)

rf_bestfitRE_temp_sdist_year_cAIC <- cAIC(model_data = rf_bestfitRE_temp_sdist_year_Data,
                                              model_obj = rf_bestfitRE_temp_sdist_year_Obj,  
                                              model_parameters = rf_bestfitRE_temp_sdist_year_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "rf_bestfitRE_temp_sdist_year",
                                              data_observation_names = "D_i")

rf_bestfitRE_temp_sdist_year_AIC <- 2*rf_bestfitRE_temp_sdist_year_Opt$objective + 2 * length(rf_bestfitRE_temp_sdist_year_Opt$par)

rf_bestfitRE_temp_sdist_cAIC <- cAIC(model_data = rf_bestfitRE_temp_sdist_Data,
                                              model_obj = rf_bestfitRE_temp_sdist_Obj,  
                                              model_parameters = rf_bestfitRE_temp_sdist_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "rf_bestfitRE_temp_sdist",
                                              data_observation_names = "D_i")

rf_bestfitRE_temp_sdist_AIC <- 2*rf_bestfitRE_temp_sdist_Opt$objective + 2 * length(rf_bestfitRE_temp_sdist_Opt$par)

rf_bestfitRE_stemp_dist_cAIC <- cAIC(model_data = rf_bestfitRE_stemp_dist_Data,
                                              model_obj = rf_bestfitRE_stemp_dist_Obj,  
                                              model_parameters = rf_bestfitRE_stemp_dist_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "rf_bestfitRE_stemp_dist",
                                              data_observation_names = "D_i")

rf_bestfitRE_stemp_dist_AIC <- 2*rf_bestfitRE_stemp_dist_Opt$objective + 2 * length(rf_bestfitRE_stemp_dist_Opt$par)


```


```{r, cache = TRUE, include = FALSE}
rf_bestfitRE_stemp_sdist_cAIC <- cAIC(model_data = rf_bestfitRE_stemp_sdist_Data,
                                              model_obj = rf_bestfitRE_stemp_sdist_Obj,  
                                              model_parameters = rf_bestfitRE_stemp_sdist_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "rf_bestfitRE_stemp_sdist",
                                              data_observation_names = "D_i")

rf_bestfitRE_stemp_sdist_AIC <- 2*rf_bestfitRE_stemp_sdist_Opt$objective + 2 * length(rf_bestfitRE_stemp_sdist_Opt$par)

rf_bestfitRE_temp_cAIC <- cAIC(model_data = rf_bestfitRE_temp_Data,
                                       model_obj = rf_bestfitRE_temp_Obj,  
                                       model_parameters = rf_bestfitRE_temp_Params,
                                       random = c("omega_s", "epsilon_st"),
                                       DLL = "rf_bestfitRE_temp",
                                       data_observation_names = "D_i")

rf_bestfitRE_temp_AIC <- 2*rf_bestfitRE_temp_Opt$objective + 2 * length(rf_bestfitRE_temp_Opt$par)

rf_bestfitRE_dist_cAIC <- cAIC(model_data = rf_bestfitRE_dist_Data,
                                       model_obj = rf_bestfitRE_dist_Obj,  
                                       model_parameters = rf_bestfitRE_dist_Params,
                                       random = c("omega_s", "epsilon_st"),
                                       DLL = "rf_bestfitRE_dist",
                                       data_observation_names = "D_i")

rf_bestfitRE_dist_AIC <- 2*rf_bestfitRE_dist_Opt$objective + 2 * length(rf_bestfitRE_dist_Opt$par)

rf_bestfitRE_stemp_cAIC <- cAIC(model_data = rf_bestfitRE_stemp_Data,
                                        model_obj = rf_bestfitRE_stemp_Obj,  
                                        model_parameters = rf_bestfitRE_stemp_Params,
                                        random = c("omega_s", "epsilon_st", "b_smooth"),
                                        DLL = "rf_bestfitRE_stemp",
                                        data_observation_names = "D_i")

rf_bestfitRE_stemp_AIC <- 2*rf_bestfitRE_stemp_Opt$objective + 2 * length(rf_bestfitRE_stemp_Opt$par)

rf_bestfitRE_sdist_cAIC <- cAIC(model_data = rf_bestfitRE_sdist_Data,
                                        model_obj = rf_bestfitRE_sdist_Obj,  
                                        model_parameters = rf_bestfitRE_sdist_Params,
                                        random = c("omega_s", "epsilon_st", "b_smooth"),
                                        DLL = "rf_bestfitRE_sdist",
                                        data_observation_names = "D_i")

rf_bestfitRE_sdist_AIC <- 2*rf_bestfitRE_sdist_Opt$objective + 2 * length(rf_bestfitRE_sdist_Opt$par)

rf_bestfitRE_temp_dist_cAIC <- cAIC(model_data = rf_bestfitRE_temp_dist_Data,
                                  model_obj = rf_bestfitRE_temp_dist_Obj,  
                                  model_parameters = rf_bestfitRE_temp_dist_Params,
                                  random = c("omega_s", "epsilon_st"),
                                  DLL = "rf_bestfitRE_temp_dist",
                                  data_observation_names = "D_i")

rf_bestfitRE_temp_dist_AIC <- 2*rf_bestfitRE_temp_dist_Opt$objective + 2 * length(rf_bestfitRE_temp_dist_Opt$par)

```


```{r, cache = TRUE, include = FALSE}
rf_bestfitRE_stemp_sdist_year1_cAIC <- cAIC(model_data = rf_bestfitRE_stemp_sdist_year1_Data,
                 model_obj = rf_bestfitRE_stemp_sdist_year1_Obj,  
                 model_parameters = rf_bestfitRE_stemp_sdist_year1_Params,
                 random = c("omega_s", "epsilon_st", "b_smooth"),
                 DLL = "rf_bestfitRE_stemp_sdist_year1",
                 data_observation_names = "D_i")

rf_bestfitRE_stemp_sdist_year1_AIC <- 2*rf_bestfitRE_stemp_sdist_year1_Opt$objective + 2 * length(rf_bestfitRE_stemp_sdist_year1_Opt$par)

rf_bestfitRE_temp_year1_cAIC <- cAIC(model_data = rf_bestfitRE_temp_year1_Data,
                 model_obj = rf_bestfitRE_temp_year1_Obj,  
                 model_parameters = rf_bestfitRE_temp_year1_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "rf_bestfitRE_temp_year1",
                 data_observation_names = "D_i")

rf_bestfitRE_temp_year1_AIC <- 2*rf_bestfitRE_temp_year1_Opt$objective + 2 * length(rf_bestfitRE_temp_year1_Opt$par)

rf_bestfitRE_dist_year1_cAIC <- cAIC(model_data = rf_bestfitRE_dist_year1_Data,
                 model_obj = rf_bestfitRE_dist_year1_Obj,  
                 model_parameters = rf_bestfitRE_dist_year1_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "rf_bestfitRE_dist_year1",
                 data_observation_names = "D_i")

rf_bestfitRE_dist_year1_AIC <- 2*rf_bestfitRE_dist_year1_Opt$objective + 2 * length(rf_bestfitRE_dist_year1_Opt$par)

rf_bestfitRE_stemp_year1_cAIC <- cAIC(model_data = rf_bestfitRE_stemp_year1_Data,
                 model_obj = rf_bestfitRE_stemp_year1_Obj,  
                 model_parameters = rf_bestfitRE_stemp_year1_Params,
                 random = c("omega_s", "epsilon_st", "b_smooth"),
                 DLL = "rf_bestfitRE_stemp_year1",
                 data_observation_names = "D_i")

rf_bestfitRE_stemp_year1_AIC <- 2*rf_bestfitRE_stemp_year1_Opt$objective + 2 * length(rf_bestfitRE_stemp_year1_Opt$par)

rf_bestfitRE_sdist_year_cAIC <- cAIC(model_data = rf_bestfitRE_sdist_year_Data,
                 model_obj = rf_bestfitRE_sdist_year_Obj,  
                 model_parameters = rf_bestfitRE_sdist_year_Params,
                 random = c("omega_s", "epsilon_st", "b_smooth"),
                 DLL = "rf_bestfitRE_sdist_year",
                 data_observation_names = "D_i")

rf_bestfitRE_sdist_year_AIC <- 2*rf_bestfitRE_sdist_year_Opt$objective + 2 * length(rf_bestfitRE_sdist_year_Opt$par)

rf_bestfitRE_year_cAIC <- cAIC(model_data = rf_bestfitRE_year_Data,
                 model_obj = rf_bestfitRE_year_Obj,  
                 model_parameters = rf_bestfitRE_year_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "rf_bestfitRE_year",
                 data_observation_names = "D_i")

rf_bestfitRE_year_AIC <- 2*rf_bestfitRE_year_Opt$objective + 2 * length(rf_bestfitRE_year_Opt$par)

rf_bestfitRE_noFE_cAIC <- cAIC(model_data = rf_bestfitRE_noFE_Data,
                 model_obj = rf_bestfitRE_noFE_Obj,  
                 model_parameters = rf_bestfitRE_noFE_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "rf_bestfitRE_noFE",
                 data_observation_names = "D_i")

rf_bestfitRE_noFE_AIC <- 2*rf_bestfitRE_noFE_Opt$objective + 2 * length(rf_bestfitRE_noFE_Opt$par)

```


```{r}
fixed_effects_model_comparison_table <- data.frame(model = c("as.factor(year) + temp + dist",
                                                             "as.factor(year) + s(temp) + s(dist)",
                                                             "as.factor(year) + temp",
                                                             "as.factor(year) + dist",
                                                             "as.factor(year) + s(temp)",
                                                             "as.factor(year) + s(dist)",
                                                             "as.factor(year)",
                                                             "temp + dist",
                                                             "s(temp) + s(dist)",
                                                             "temp",
                                                             "dist",
                                                             "s(temp)",
                                                             "s(dist)",
                                                             "as.factor(year) + s(temp) + dist",
                                                             "as.factor(year) + temp + s(dist)",
                                                             "temp + s(dist)",
                                                             "s(temp) + dist",
                                                             "no fixed effects"),
                                     AIC = c(model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_AIC,
                                             rf_bestfitRE_stemp_sdist_year1_AIC,
                                             rf_bestfitRE_temp_year1_AIC,
                                             rf_bestfitRE_dist_year1_AIC,
                                             rf_bestfitRE_stemp_year1_AIC,
                                             rf_bestfitRE_sdist_year_AIC,
                                             rf_bestfitRE_year_AIC,
                                             rf_bestfitRE_temp_dist_AIC,
                                             rf_bestfitRE_stemp_sdist_AIC,
                                             rf_bestfitRE_temp_AIC,
                                             rf_bestfitRE_dist_AIC,
                                             rf_bestfitRE_stemp_AIC,
                                             rf_bestfitRE_sdist_AIC,
                                             rf_bestfitRE_stemp_dist_year_AIC,
                                             rf_bestfitRE_temp_sdist_year_AIC,
                                             rf_bestfitRE_temp_sdist_AIC,
                                             rf_bestfitRE_stemp_dist_AIC,
                                             rf_bestfitRE_noFE_AIC),
                                     cAIC = c(model3_3_spatial_aniso_spatiotemporal_IID_aniso_missing_years_cAIC,
                                              rf_bestfitRE_stemp_sdist_year1_cAIC,
                                              rf_bestfitRE_temp_year1_cAIC,
                                              rf_bestfitRE_dist_year1_cAIC,
                                              rf_bestfitRE_stemp_year1_cAIC,
                                              rf_bestfitRE_sdist_year_cAIC,
                                              rf_bestfitRE_year_cAIC,
                                              rf_bestfitRE_temp_dist_cAIC,
                                             rf_bestfitRE_stemp_sdist_cAIC,
                                             rf_bestfitRE_temp_cAIC,
                                             rf_bestfitRE_dist_cAIC,
                                             rf_bestfitRE_stemp_cAIC,
                                             rf_bestfitRE_sdist_cAIC,
                                             rf_bestfitRE_stemp_dist_year_cAIC,
                                             rf_bestfitRE_temp_sdist_year_cAIC,
                                             rf_bestfitRE_temp_sdist_cAIC,
                                             rf_bestfitRE_stemp_dist_cAIC,
                                              rf_bestfitRE_noFE_cAIC))

fixed_effects_model_comparison_table %>% 
  arrange(cAIC) %>% 
  mutate(AIC = round(AIC, 2)) %>% 
  mutate(cAIC = round(cAIC, 2)) %>% 
  kbl(caption = "AIC and cAIC scores for the fixed effect structures.") %>% 
  kable_styling(latex_options = "HOLD_position")
```

cAIC does not support the inclusion of any fixed effects.

Given that cAIC is the most appropriate model selection metric for our purposes, we will model the distribution of YOY rockfishes with the following model structure:

**density ~ spatial random fields (anisotropic) + spatiotemporal random fields (anisotropic, IID)**



#### Export the tables of AIC and cAIC for the supplemental results
```{r}
write.csv(random_effects_model_comparison_table, here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "04.3_rf_random_effects_model_comparison_table.csv"))
write.csv(fixed_effects_model_comparison_table, here::here("R", "04_SDM_model_selection", "04.3_rf_SDM", "04.3_rf_fixed_effects_model_comparison_table.csv"))
```






