---
title: "06.2.6.1_fit_seabird_two_prey_SAR"
author: "Markus Min"
date: "`r Sys.Date()`"
output: html_document
---

This model is a sensitivity analysis, where the errors-in-variables approach is not used.

This document fits the 06.2.6.1_seabird_two_prey_SAR model, which incorporates the following data sources:

1. JSOES Bongo data (biomass)
2. PWCC/PRS rockfish data
3. JSOES trawl data (for yearling and subyearling Interior Chinook)
4. JSOES seabird data (for Common Murres and Sooty Shearwaters)

The prey metrics come from 05.1, and the seabird metrics come from 05.2.

This model does not account for the abundance of salmon in the survey.

This model is fit to the combined SRF + UCSF dataset, and incorporates the effect of 
transport (and stock group) by having three terms for the effect of SRF (transported),
SRF (non-transported), and UCSF (non-transported). It also does not include
an intercept term (no beta_0), with the previous terms instead functioning as 
intercepts for the three groups.

Outmigration terms are also separate for the two stock groups.



```{r load_libraries_source_scripts, include = FALSE}
library(tidyverse)
library(here)
library(TMB)
library(ggforce)
library(sdmTMB)
library(kableExtra)
library(fmesher)
library(ggrepel)
# knitr::opts_chunk()

# source the function scripts
source("R/functions/cAIC.R")
source("R/functions/rmvnorm_prec.R")
source("R/functions/sample_var.R")
source("R/functions/plot_anisotropy_MM.R")
source("R/functions/make_anisotropy_spde.R")
source("R/functions/H_matrix_prior_predictive_check_for_ggplot.R")
source("R/functions/plot_distribution.R")
source("R/functions/plot_distribution_PRS_PWCC.R")
source("R/functions/plot_distribution_jsoes_bongo.R")
# source("R/functions/plot_distribution_hake_survey.R")
source("R/functions/generate_prediction_maps.R")
source("R/functions/generate_prediction_maps_PRS_PWCC.R")
source("R/functions/generate_prediction_maps_jsoes_bongo.R")
source("R/functions/generate_prediction_maps_hake_survey.R")
source("R/functions/generate_prediction_maps_cces.R")
source("R/functions/stage2_helper_functions.R")

# source the prep scripts for each of the surveys
source(here::here("R", "PRS_PWCC_make_mesh.R"))
source(here::here("R", "JSOES_seabirds_make_mesh.R"))
source(here::here("R", "JSOES_make_mesh.R"))

```

## Prepare survey data for SDMs

#### CSYIF + CSSIF (JSOES Trawl)
```{r csyif_cssif_data_prep}

#### prep model data ####

# extract Yearling Interior Chinook
csyif <- subset(jsoes_long, species == "chinook_salmon_yearling_interior_fa")
# extract subyearling Interior Chinook
cssif <- subset(jsoes_long, species == "chinook_salmon_subyearling_interior_fa")

# generate the projection vector/matrix for covariates - this relies
# on the survey prediction grid from the JSOES_make_mesh.R script

distance_projection_vector <- as.numeric(filter(survey_predict_grid, year == 2000)$dist_shore_scaled)

# for SST - take prediction grid, drop salinity, add SST with locations as rows and years as column
survey_predict_grid %>% 
  # select only the years that we're keeping
  filter(!(year %in% c(1998, 2022:2025))) %>% 
  dplyr::select(-c(SST, sal, sal_scaled, dist_shore, dist_shore_scaled)) %>% 
  pivot_wider(names_from = year, values_from = SST_scaled) %>% 
  dplyr::select(-c(X, Y)) %>% 
  as.matrix() -> temperature_projection_matrix

#### Create SPDE objects ####

# there are multiple meshes that are created in JSOES_make_mesh.R
# here, we will use the cutoff 15 mesh, although a higher resolution is likely going
# to improve results (at the cost of run time)

trawl_spde <- fm_fem( inla_mesh_cutoff15, 
                      order = 2 )

# create projection matrix from vertices to samples
trawl_is = fm_evaluator(inla_mesh_cutoff15, loc = as.matrix(data.frame(X = csyif$X, Y = csyif$Y)))$proj$A

# create projection matrix from vertices to prediction grid
trawl_gs = fm_evaluator( inla_mesh_cutoff15, loc=st_coordinates(survey_domain_cov_grid))$proj$A

# prep anisotropy input
# note that spatial list replaces the M0, M1, M2 objects 
spatial_list_trawl <- make_anisotropy_spde( inla_mesh_cutoff15 )
```



#### Common Murres + Sooty Shearwaters (JSOES seabirds)

```{r data_prep}
#### prep model data ####
# extract Sooty Shearwaters
sosh <- subset(birds_long, species == "sooty_shearwater")
comu <- subset(birds_long, species == "common_murre")

# generate the projection vector/matrix for covariates - this relies
# on the survey prediction grid from the JSOES_make_mesh.R script

distance_projection_vector <- as.numeric(filter(survey_predict_grid, year == 2010)$dist_shore_scaled)

# for SST - take prediction grid, drop salinity, add SST with locations as rows and years as column
survey_predict_grid %>% 
  # select only the years that we're keeping
  # filter(!(year %in% c(1998, 2022:2025))) %>% 
  filter(year %in% unique(sosh$year)) %>% 
  dplyr::select(-c(SST, dist_shore, dist_shore_scaled)) %>% 
  pivot_wider(names_from = year, values_from = SST_scaled) %>% 
  dplyr::select(-c(X, Y)) %>% 
  as.matrix() -> temperature_projection_matrix
  
# create a second temperature projection grid that contains the missing years
survey_predict_grid_birds_allyears %>% 
  # select only the years that we're keeping
  filter(!(year %in% c(1998, 2022:2025))) %>% 
  dplyr::select(-c(SST, dist_shore, dist_shore_scaled)) %>% 
  pivot_wider(names_from = year, values_from = SST_scaled) %>% 
  dplyr::select(-c(X, Y)) %>% 
  as.matrix() -> temperature_projection_matrix_birds_allyears

#### Create SPDE objects ####

# there are multiple meshes that are created in JSOES_make_mesh.R
# here, we will use the cutoff 15 mesh, although a higher resolution is likely going
# to improve results (at the cost of run time)

### June SPDE objects

birds_spde <- fm_fem( birds_inla_mesh_cutoff15, 
                      order = 2 )

# create projection matrix from vertices to samples
A_is_birds = fm_evaluator(birds_inla_mesh_cutoff15, loc = as.matrix(data.frame(X = comu$X, Y = comu$Y)))$proj$A

# create projection matrix from vertices to prediction grid
A_gs_birds = fm_evaluator( birds_inla_mesh_cutoff15, loc=st_coordinates(survey_domain_cov_grid))$proj$A

# prep anisotropy input
# note that spatial list replaces the M0, M1, M2 objects 
spatial_list_birds <- make_anisotropy_spde( birds_inla_mesh_cutoff15 )
```

#### Shrimp Larvae, Non-Cancer Crab Larvae, Crab Larvae, Hyperiid Amphipods (JSOES Bongo)

```{r data_prep}
#### load and reformat bongo data ####

jsoes_bongo_biomass_cancer_crab_larvae <- read.csv(here::here("model_inputs", "jsoes_bongo_biomass_cancer_crab_larvae.csv"))
jsoes_bongo_biomass_hyperiid_amphipods <- read.csv(here::here("model_inputs", "jsoes_bongo_biomass_hyperiid_amphipod.csv"))
jsoes_bongo_biomass_non_cancer_crab_larvae <- read.csv(here::here("model_inputs", "jsoes_bongo_biomass_non_cancer_crab_larvae.csv"))
jsoes_bongo_biomass_shrimp_larvae <- read.csv(here::here("model_inputs", "jsoes_bongo_biomass_shrimp_larvae.csv"))

# drop 1998 (bongo doesn't have this year) and 2022:2025 (adult returns aren't complete for those years yet)
jsoes_bongo_biomass_cancer_crab_larvae <- subset(jsoes_bongo_biomass_cancer_crab_larvae, !(year %in% c(1998, 2022:2025)))
jsoes_bongo_biomass_hyperiid_amphipods <- subset(jsoes_bongo_biomass_hyperiid_amphipods, !(year %in% c(1998, 2022:2025)))
jsoes_bongo_biomass_non_cancer_crab_larvae <- subset(jsoes_bongo_biomass_non_cancer_crab_larvae, !(year %in% c(1998, 2022:2025)))
jsoes_bongo_biomass_shrimp_larvae <- subset(jsoes_bongo_biomass_shrimp_larvae, !(year %in% c(1998, 2022:2025)))



# add a nmi and km from shore from shore column based on the station codes
add_nmi_km <- function(data){
  data %>% 
    mutate(nmi_from_shore = parse_number(station)) %>% 
    # change nmi to km to keep units consistent
    mutate(km_from_shore = nmi_from_shore * 1.852) -> output
  
  return(output)
}

jsoes_bongo_biomass_cancer_crab_larvae <- add_nmi_km(jsoes_bongo_biomass_cancer_crab_larvae)
jsoes_bongo_biomass_hyperiid_amphipods <- add_nmi_km(jsoes_bongo_biomass_hyperiid_amphipods)
jsoes_bongo_biomass_non_cancer_crab_larvae <- add_nmi_km(jsoes_bongo_biomass_non_cancer_crab_larvae)
jsoes_bongo_biomass_shrimp_larvae <- add_nmi_km(jsoes_bongo_biomass_shrimp_larvae)

# the other script already deals with the crs, and the X and Y columns contain
# the UTM zone 10 distances in km


#### load and reformat spatial data ####
usa_spdf <- st_read(here::here("Data", "map_files", "USA_adm0.shp"))
# load BC
CAN_spdf <- st_read(here::here("Data", "map_files", "canada", "lpr_000b16a_e.shp"))
BC_spdf <- filter(CAN_spdf, PRENAME == "British Columbia")
BC_proj <- st_transform(BC_spdf, crs = 4326)


# crop them to our desired area
US_west_coast <- sf::st_crop(usa_spdf,
                             c(xmin = -126, ymin = 44, xmax = -123, ymax = 48.5))

BC_coast <- sf::st_crop(BC_proj,
                        c(xmin = -126, ymin = 44, xmax = -123, ymax = 48.5))



# convert both shapefiles to a different projection (UTM zone 10) so that they can be plotted with the sdmTMB output
UTM_zone_10_crs <- 32610

US_west_coast_proj <- sf::st_transform(US_west_coast, crs = UTM_zone_10_crs)
BC_coast_proj <- sf::st_transform(BC_coast, crs = UTM_zone_10_crs)

# make this projection into kilometers
US_west_coast_proj_km <- st_as_sf(US_west_coast_proj$geometry/1000, crs = UTM_zone_10_crs)
BC_coast_proj_km <- st_as_sf(BC_coast_proj$geometry/1000, crs = UTM_zone_10_crs)



#### load and reformat SST data ####
SST1 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_9bdb_17ce_0aa0.csv"), skip = 1)
SST2 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_36e3_7cb3_1109.csv"), skip = 1)
SST3 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_519c_f955_0fea.csv"), skip = 1)
SST4 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_727c_5c10_e6a0.csv"), skip = 1)
SST5 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_1176_fc5f_53d2.csv"), skip = 1)
SST6 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_ae0c_9197_083f.csv"), skip = 1)
SST7 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_fe4f_9c9c_442e.csv"), skip = 1)

SST1 %>% 
  bind_rows(., SST2, SST3, SST4, SST5, SST6, SST7) -> SST

SST %>% 
  dplyr::rename(time = UTC, latitude = degrees_north, longitude = degrees_east, SST = degree_C) -> SST
SST %>% 
  mutate(time = ymd_hms(time)) %>% 
  mutate(day_of_month = mday(time)) %>% 
  mutate(year = year(time)) %>% 
  mutate(julian = yday(time))-> SST

# Looks good to me! Now just need to resolve CRS conflicts

# Convert SST to UTM zone 10
# st_as_sf(SST, coords = c("latitude", "longitude"), crs = "4326") -> SST_sf
st_as_sf(SST, coords = c("longitude", "latitude"), crs = 4326) -> SST_sf
US_west_coast

sf::st_transform(SST_sf, crs = UTM_zone_10_crs) -> SST_sf_proj

# convert to km
SST_sf_proj_km <- SST_sf_proj
SST_sf_proj_km$geometry <- SST_sf_proj$geometry/1000

SST_sf_proj_km <- st_set_crs(SST_sf_proj_km, UTM_zone_10_crs)

# Match this SST data to the JSOES bongo data
# jsoes needs to be in SF format as well

st_as_sf(jsoes_bongo_biomass_cancer_crab_larvae, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_cancer_crab_larvae_sf
st_as_sf(jsoes_bongo_biomass_hyperiid_amphipods, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_hyperiid_amphipods_sf
st_as_sf(jsoes_bongo_biomass_non_cancer_crab_larvae, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_non_cancer_crab_larvae_sf
st_as_sf(jsoes_bongo_biomass_shrimp_larvae, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_shrimp_larvae_sf

jsoes_bongo_biomass_cancer_crab_larvae_sf_split <- split(jsoes_bongo_biomass_cancer_crab_larvae_sf, jsoes_bongo_biomass_cancer_crab_larvae_sf$year)
jsoes_bongo_biomass_hyperiid_amphipods_sf_split <- split(jsoes_bongo_biomass_hyperiid_amphipods_sf, jsoes_bongo_biomass_hyperiid_amphipods_sf$year)
jsoes_bongo_biomass_non_cancer_crab_larvae_sf_split <- split(jsoes_bongo_biomass_non_cancer_crab_larvae_sf, jsoes_bongo_biomass_non_cancer_crab_larvae_sf$year)
jsoes_bongo_biomass_shrimp_larvae_sf_split <- split(jsoes_bongo_biomass_shrimp_larvae_sf, jsoes_bongo_biomass_shrimp_larvae_sf$year)
SST_sf_proj_km_split <- split(SST_sf_proj_km, SST_sf_proj_km$year) 

jsoes_years <- as.character(unique(jsoes_bongo_biomass_cancer_crab_larvae_sf$year))

jsoes_bongo_biomass_cancer_crab_larvae_remote_SST_list <- map(jsoes_years, function(year) {
  jsoes_one_year <- jsoes_bongo_biomass_cancer_crab_larvae_sf_split[[year]]
  SST_one_year <- SST_sf_proj_km_split[[year]]
  
  # Find nearest features
  SST_index_for_jsoes <- st_nearest_feature(jsoes_one_year, SST_one_year)
  
  # Combine with matched points (optional: include matched geometry or attributes)
  jsoes_one_year %>%
    mutate(matched_id = SST_index_for_jsoes) %>%
    bind_cols(
      st_drop_geometry(SST_one_year[SST_index_for_jsoes, "SST"]) %>%
        dplyr::rename(remote_SST = SST)
    ) %>% 
    dplyr::select(-matched_id) -> output
  
  # transform this to a df
  st_drop_geometry(output) %>% 
    bind_cols(st_coordinates(output)) -> output
  
  return(output)
})

# take the list and turn it back into an sf object
jsoes_bongo_biomass_cancer_crab_larvae <- list_rbind(jsoes_bongo_biomass_cancer_crab_larvae_remote_SST_list)

jsoes_bongo_biomass_hyperiid_amphipods_sf_split_remote_SST_list <- map(jsoes_years, function(year) {
  jsoes_one_year <- jsoes_bongo_biomass_hyperiid_amphipods_sf_split[[year]]
  SST_one_year <- SST_sf_proj_km_split[[year]]
  
  # Find nearest features
  SST_index_for_jsoes <- st_nearest_feature(jsoes_one_year, SST_one_year)
  
  # Combine with matched points (optional: include matched geometry or attributes)
  jsoes_one_year %>%
    mutate(matched_id = SST_index_for_jsoes) %>%
    bind_cols(
      st_drop_geometry(SST_one_year[SST_index_for_jsoes, "SST"]) %>%
        dplyr::rename(remote_SST = SST)
    ) %>% 
    dplyr::select(-matched_id) -> output
  
  # transform this to a df
  st_drop_geometry(output) %>% 
    bind_cols(st_coordinates(output)) -> output
  
  return(output)
})

# take the list and turn it back into an sf object
jsoes_bongo_biomass_hyperiid_amphipods <- list_rbind(jsoes_bongo_biomass_hyperiid_amphipods_sf_split_remote_SST_list)


jsoes_bongo_biomass_non_cancer_crab_larvae_remote_SST_list <- map(jsoes_years, function(year) {
  jsoes_one_year <- jsoes_bongo_biomass_non_cancer_crab_larvae_sf_split[[year]]
  SST_one_year <- SST_sf_proj_km_split[[year]]
  
  # Find nearest features
  SST_index_for_jsoes <- st_nearest_feature(jsoes_one_year, SST_one_year)
  
  # Combine with matched points (optional: include matched geometry or attributes)
  jsoes_one_year %>%
    mutate(matched_id = SST_index_for_jsoes) %>%
    bind_cols(
      st_drop_geometry(SST_one_year[SST_index_for_jsoes, "SST"]) %>%
        dplyr::rename(remote_SST = SST)
    ) %>% 
    dplyr::select(-matched_id) -> output
  
  # transform this to a df
  st_drop_geometry(output) %>% 
    bind_cols(st_coordinates(output)) -> output
  
  return(output)
})

# take the list and turn it back into an sf object
jsoes_bongo_biomass_non_cancer_crab_larvae <- list_rbind(jsoes_bongo_biomass_non_cancer_crab_larvae_remote_SST_list)


jsoes_bongo_biomass_shrimp_larvae_remote_SST_list <- map(jsoes_years, function(year) {
  jsoes_one_year <- jsoes_bongo_biomass_shrimp_larvae_sf_split[[year]]
  SST_one_year <- SST_sf_proj_km_split[[year]]
  
  # Find nearest features
  SST_index_for_jsoes <- st_nearest_feature(jsoes_one_year, SST_one_year)
  
  # Combine with matched points (optional: include matched geometry or attributes)
  jsoes_one_year %>%
    mutate(matched_id = SST_index_for_jsoes) %>%
    bind_cols(
      st_drop_geometry(SST_one_year[SST_index_for_jsoes, "SST"]) %>%
        dplyr::rename(remote_SST = SST)
    ) %>% 
    dplyr::select(-matched_id) -> output
  
  # transform this to a df
  st_drop_geometry(output) %>% 
    bind_cols(st_coordinates(output)) -> output
  
  return(output)
})

# take the list and turn it back into an sf object
jsoes_bongo_biomass_shrimp_larvae <- list_rbind(jsoes_bongo_biomass_shrimp_larvae_remote_SST_list)

#### Calculate distance to shore ####

### cancer crab larvae
dplyr::select(jsoes_bongo_biomass_cancer_crab_larvae, station, km_from_shore, X, Y) -> jsoes_bongo_biomass_cancer_crab_larvae_spatial
st_as_sf(jsoes_bongo_biomass_cancer_crab_larvae_spatial, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_cancer_crab_larvae_sf
st_distance(jsoes_bongo_biomass_cancer_crab_larvae_sf, US_west_coast_proj_km) -> jsoes_bongo_biomass_cancer_crab_larvae_dist_shore
jsoes_bongo_biomass_cancer_crab_larvae$sf_dist_shore <- as.numeric(jsoes_bongo_biomass_cancer_crab_larvae_dist_shore)

# confirm that these are reasonable, based on points from survey
# they're similar - but not identical
hist(as.numeric(jsoes_bongo_biomass_cancer_crab_larvae$sf_dist_shore)-jsoes_bongo_biomass_cancer_crab_larvae_spatial$km_from_shore)
# and they're pretty biased - the distances calculated using sf tend to be smaller
# Let's just use the sf distance from shore calculated for both prediction grid and
# the samples themselves
# Let's also use the SST from remote sensing, instead of the in situ data


# rescale data
jsoes_bongo_biomass_cancer_crab_larvae %>% 
  mutate(SST_scaled = as.numeric(scale(remote_SST)),
         dist_shore_scaled = as.numeric(scale(sf_dist_shore))) -> jsoes_bongo_biomass_cancer_crab_larvae

### hyperiid amphipods
dplyr::select(jsoes_bongo_biomass_hyperiid_amphipods, station, km_from_shore, X, Y) -> jsoes_bongo_biomass_hyperiid_amphipods_spatial
st_as_sf(jsoes_bongo_biomass_hyperiid_amphipods_spatial, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_hyperiid_amphipods_sf
st_distance(jsoes_bongo_biomass_hyperiid_amphipods_sf, US_west_coast_proj_km) -> jsoes_bongo_biomass_hyperiid_amphipods_dist_shore
jsoes_bongo_biomass_hyperiid_amphipods$sf_dist_shore <- as.numeric(jsoes_bongo_biomass_hyperiid_amphipods_dist_shore)

# confirm that these are reasonable, based on points from survey
# they're similar - but not identical
hist(as.numeric(jsoes_bongo_biomass_hyperiid_amphipods$sf_dist_shore)-jsoes_bongo_biomass_hyperiid_amphipods_spatial$km_from_shore)
# and they're pretty biased - the distances calculated using sf tend to be smaller
# Let's just use the sf distance from shore calculated for both prediction grid and
# the samples themselves
# Let's also use the SST from remote sensing, instead of the in situ data


# rescale data
jsoes_bongo_biomass_hyperiid_amphipods %>% 
  mutate(SST_scaled = as.numeric(scale(remote_SST)),
         dist_shore_scaled = as.numeric(scale(sf_dist_shore))) -> jsoes_bongo_biomass_hyperiid_amphipods

### non-cancer crab larvae
dplyr::select(jsoes_bongo_biomass_non_cancer_crab_larvae, station, km_from_shore, X, Y) -> jsoes_bongo_biomass_non_cancer_crab_larvae_spatial
st_as_sf(jsoes_bongo_biomass_non_cancer_crab_larvae_spatial, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_non_cancer_crab_larvae_sf
st_distance(jsoes_bongo_biomass_non_cancer_crab_larvae_sf, US_west_coast_proj_km) -> jsoes_bongo_biomass_non_cancer_crab_larvae_dist_shore
jsoes_bongo_biomass_non_cancer_crab_larvae$sf_dist_shore <- as.numeric(jsoes_bongo_biomass_non_cancer_crab_larvae_dist_shore)

# confirm that these are reasonable, based on points from survey
# they're similar - but not identical
hist(as.numeric(jsoes_bongo_biomass_non_cancer_crab_larvae$sf_dist_shore)-jsoes_bongo_biomass_non_cancer_crab_larvae_spatial$km_from_shore)
# and they're pretty biased - the distances calculated using sf tend to be smaller
# Let's just use the sf distance from shore calculated for both prediction grid and
# the samples themselves
# Let's also use the SST from remote sensing, instead of the in situ data


# rescale data
jsoes_bongo_biomass_non_cancer_crab_larvae %>% 
  mutate(SST_scaled = as.numeric(scale(remote_SST)),
         dist_shore_scaled = as.numeric(scale(sf_dist_shore))) -> jsoes_bongo_biomass_non_cancer_crab_larvae


### shrimp larvae
dplyr::select(jsoes_bongo_biomass_shrimp_larvae, station, km_from_shore, X, Y) -> jsoes_bongo_biomass_shrimp_larvae_spatial
st_as_sf(jsoes_bongo_biomass_shrimp_larvae_spatial, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_shrimp_larvae_sf
st_distance(jsoes_bongo_biomass_shrimp_larvae_sf, US_west_coast_proj_km) -> jsoes_bongo_biomass_shrimp_larvae_dist_shore
jsoes_bongo_biomass_shrimp_larvae$sf_dist_shore <- as.numeric(jsoes_bongo_biomass_shrimp_larvae_dist_shore)

# confirm that these are reasonable, based on points from survey
# they're similar - but not identical
hist(as.numeric(jsoes_bongo_biomass_shrimp_larvae$sf_dist_shore)-jsoes_bongo_biomass_shrimp_larvae_spatial$km_from_shore)
# and they're pretty biased - the distances calculated using sf tend to be smaller
# Let's just use the sf distance from shore calculated for both prediction grid and
# the samples themselves
# Let's also use the SST from remote sensing, instead of the in situ data


# rescale data
jsoes_bongo_biomass_shrimp_larvae %>% 
  mutate(SST_scaled = as.numeric(scale(remote_SST)),
         dist_shore_scaled = as.numeric(scale(sf_dist_shore))) -> jsoes_bongo_biomass_shrimp_larvae


# for SST - take prediction grid, drop distance from shore, add SST with locations as rows and years as column
survey_predict_grid %>% 
  # select only the years that we're keeping
  filter(!(year %in% c(1998, 2022:2025))) %>% 
  dplyr::select(-c(SST, sal, sal_scaled, dist_shore, dist_shore_scaled)) %>% 
  pivot_wider(names_from = year, values_from = SST_scaled) %>% 
  dplyr::select(-c(X, Y)) %>% 
  as.matrix() -> temperature_projection_matrix


distance_projection_vector <- as.numeric(filter(survey_predict_grid, year == 2010)$dist_shore_scaled)

#### Create SPDE objects ####

# note that all bongo taxa have the same samples, so any of the corresponding 
# data frames can be used for this section

bongo_spde <- fm_fem( bongo_inla_mesh_cutoff15, 
                      order = 2 )

# create projection matrix from vertices to samples
A_is_bongo = fm_evaluator(bongo_inla_mesh_cutoff15, loc = as.matrix(data.frame(X = jsoes_bongo_biomass_shrimp_larvae$X, Y = jsoes_bongo_biomass_shrimp_larvae$Y)))$proj$A

# create projection matrix from vertices to prediction grid
A_gs_bongo = fm_evaluator( bongo_inla_mesh_cutoff15, loc=st_coordinates(survey_domain_cov_grid))$proj$A

# prep anisotropy input
# note that spatial list replaces the M0, M1, M2 objects 
spatial_list_bongo <- make_anisotropy_spde( bongo_inla_mesh_cutoff15 )
```


#### Rockfishes (PWCC/PRS)

```{r}
# make sure rf density is numeric not integer
rf$total <- as.numeric(rf$total)

#### prep model data ####
# generate the projection vector/matrix for covariates - this relies
# on the survey prediction grid from the PRS_PWCC_make_mesh.R script

distance_projection_vector <- as.numeric(filter(survey_predict_grid_jsoes_rf_years, year == 2016)$dist_shore_scaled)

# for SST - take prediction grid, drop salinity, add SST with locations as rows and years as column
survey_predict_grid_jsoes_rf_years %>% 
  # select only the years that we're keeping
  filter(!(year %in% c(1998, 2022:2025))) %>% 
  dplyr::select(-c(SST, dist_shore, dist_shore_scaled)) %>% 
  pivot_wider(names_from = year, values_from = SST_scaled) %>% 
  dplyr::select(-c(X, Y)) %>% 
  as.matrix() -> temperature_projection_matrix

# create a second temperature projection grid that contains the missing years
survey_predict_grid_jsoes_rf_allyears %>% 
  # select only the years that we're keeping
  filter(!(year %in% c(1998, 2022:2025))) %>% 
  dplyr::select(-c(SST, dist_shore, dist_shore_scaled)) %>% 
  pivot_wider(names_from = year, values_from = SST_scaled) %>% 
  dplyr::select(-c(X, Y)) %>% 
  as.matrix() -> temperature_projection_matrix_rf_allyears


#### Create SPDE objects ####

# there are multiple meshes that are created in JSOES_make_mesh.R
# here, we will use the cutoff 15 mesh, although a higher resolution is likely going
# to improve results (at the cost of run time)

PRS_PWCC_spde <- fm_fem( inla_mesh_cutoff15_jsoes_domain, 
                order = 2 )

# create projection matrix from vertices to samples
A_is_rf = fm_evaluator(inla_mesh_cutoff15_jsoes_domain, loc = as.matrix(data.frame(X = rf$X, Y = rf$Y)))$proj$A

# create projection matrix from vertices to prediction grid
# use the prediction grid for only the JSOES domain here

#### Create a survey prediction grid 
# Create a survey grid from scratch, using SST data

# only keep one year, since that's all we need to make a grid
SST_sf_proj %>% 
  filter(year == 2000) -> SST_sf_proj_2000

st_as_sf(SST_sf_proj_2000$geometry/1000, crs = UTM_zone_10_crs) -> SST_sf_proj_km_2000

SST_sf_proj_km_2000$dist_shore <- as.numeric(st_distance(SST_sf_proj_km_2000, US_west_coast_proj_km))


# KEY DECISION: To what area do you want to project the SDM to?
# Some observations of the data:
# The vast majority of stations are within 60 km of shore, but there are just a couple 
# that are further out in the Columbia River Plume up to 83 km from shore
# a total of 15 tows (out of over 2000) are >= 35 nmi (65 km) offshore; many more are 30 or 31 nmi (55.5 or 57.4 km) offshore.
# We also want to exclude values that are within 0.5 km of shore, because these
# don't have measureable SST values (they're basically on shore)
# this can also be seen in the following histogram:
# hist(jsoes_long$km_from_shore)

# For the N/S dimension, the most northerly transect is typically the FS transect and 
# the most southerly transect is the NH transect.
# this corresponds to 44.67 to 48.23 N
# Let's take these as our N/S boundaries and add 10 km as a buffer.
min_Y <- mean(subset(jsoes_long, grepl("NH", station))$Y)
max_Y <- mean(subset(jsoes_long, grepl("FS", station))$Y)

# Based on this, I'm going to propose that we generate a prediction that extends:
# from 0.5 to 65 km offshore
# from 44.67 to 48.23 N (+ 10 km on either side)

# trim longitudinally
SST_sf_proj_km_2000 %>% 
  filter(dist_shore <= 65 & dist_shore >= 0.5) -> grid_within_65km_shore

# trim latitudinally
survey_domain <- st_crop(grid_within_65km_shore, xmin = 0, xmax = 100000, ymin=min_Y-10, ymax=max_Y+10)

# let's inspect the grid that we created
# ggplot(survey_domain) +
#   geom_sf()

# We will need to manually trim out some of these areas

# manually remove sections for strait, hood canal (and other inland waters), Grays Harbor, Willapa Bay, Columbia River estuary
strait <- st_as_sfc(st_bbox(c(xmin=-124.65, xmax=-122, ymin=47.9375, ymax=49), crs = "WGS84"))
strait_proj <- sf::st_transform(strait, crs = UTM_zone_10_crs)
strait_proj_km <- st_as_sf(strait_proj/1000, crs = UTM_zone_10_crs)

inland_waters <- st_as_sfc(st_bbox(c(xmin=-123.5, xmax=-120, ymin=44, ymax=49.5), crs = "WGS84"))
inland_waters_proj <- sf::st_transform(inland_waters, crs = UTM_zone_10_crs)
inland_waters_proj_km <- st_as_sf(inland_waters_proj/1000, crs = UTM_zone_10_crs)

grays_harbor <- st_as_sfc(st_bbox(c(xmin=-124.15, xmax=-123.6, ymin=46.83, ymax=47.09), crs = "WGS84"))
grays_harbor_proj <- sf::st_transform(grays_harbor, crs = UTM_zone_10_crs)
grays_harbor_proj_km <- st_as_sf(grays_harbor_proj/1000, crs = UTM_zone_10_crs)

willapa_bay <- st_as_sfc(st_bbox(c(xmin=-124.06, xmax=-123.5, ymin=46.34, ymax=46.8), crs = "WGS84"))
willapa_bay_proj <- sf::st_transform(willapa_bay, crs = UTM_zone_10_crs)
willapa_bay_proj_km <- st_as_sf(willapa_bay_proj/1000, crs = UTM_zone_10_crs)

estuary <- st_as_sfc(st_bbox(c(xmin=-124.02, xmax=-123.5, ymin=46.12, ymax=46.35), crs = "WGS84"))
estuary_proj <- sf::st_transform(estuary, crs = UTM_zone_10_crs)
estuary_proj_km <- st_as_sf(estuary_proj/1000, crs = UTM_zone_10_crs)

survey_domain %>% 
  st_difference(., strait_proj_km) %>% 
  st_difference(., inland_waters_proj_km) %>% 
  st_difference(., grays_harbor_proj_km) %>% 
  st_difference(., willapa_bay_proj_km) %>% 
  st_difference(., estuary_proj_km) -> survey_domain

# Create a concave hull around points
st_concave_hull(st_union(survey_domain), ratio = 0.1) -> survey_domain_polygon

# ggplot(survey_domain_polygon) +
#   geom_sf()

# ggplot(survey_domain) +
#   geom_sf()

# let's inspect our revised survey domain
survey_area_basemap_km +
  geom_sf(data = survey_domain_polygon, fill = "blue", alpha = 0.2)

# Take SST grid and intersect with survey domain
SST_sf_proj %>% 
  mutate(geometry = geometry/1000) %>% 
  st_set_crs(UTM_zone_10_crs) %>% 
  st_intersection(survey_domain_polygon) %>% 
  dplyr::select(SST, year, geometry)-> SST_survey_domain

survey_area_basemap_km +
  geom_sf(data = SST_survey_domain, fill = "blue", alpha = 0.2)

# add in distance from shore - this is now your survey domain, with covariates
SST_survey_domain %>% 
  mutate(sf_dist_shore = as.numeric(st_distance(geometry, US_west_coast_proj_km))) -> survey_domain_cov
survey_domain_cov <- st_as_sf(survey_domain_cov, coords = c("X", "Y"), crs = UTM_zone_10_crs)



# because the survey_domain_polygon is constructed from a concave hull, some of the borders
# are a little coarse - and as a result some of the super nearshore data snuck back it.
# filter it out
survey_domain_cov %>% 
  filter(sf_dist_shore >= 0.5) -> survey_domain_cov


# extract the grid for the projection area - the grid dimensions are the same each year, so just select one year and drop fields besides geometry
survey_domain_cov %>% 
  filter(year == 2016) %>% 
  dplyr::select(geometry) -> survey_domain_cov_grid

A_gs_rf = fm_evaluator( inla_mesh_cutoff15_jsoes_domain, loc=st_coordinates(survey_domain_cov_grid))$proj$A
# A_gs_rf = fm_evaluator( inla_mesh_cutoff15_jsoes_domain, loc=st_coordinates(survey_domain_jsoes_cov_grid))$proj$A

# prep anisotropy input
# note that spatial list replaces the M0, M1, M2 objects 
spatial_list_rf <- make_anisotropy_spde( inla_mesh_cutoff15_jsoes_domain )
```



## Estimate uncertainty for quantities that weren't ADREPORTed

### 05.1_jsoes_bongo_biomass_model
```{r}
# If this has already been run before, just load the output instead of running the sample_var code again
load(here::here("R", "05_stage1_SDM", "05.1.1_jsoes_bongo_biomass_SDM", "estimated_SE_prey_model.rda"))
SE_pianka_o_csyif_prey_field_t_prey_model <- SE_prey_model$SE_pianka_o_csyif_prey_field_t_prey_model
SE_pianka_o_cssif_prey_field_t_prey_model <- SE_prey_model$SE_pianka_o_cssif_prey_field_t_prey_model

#### 05.1.1_jsoes_bongo_biomass_SDM ####
load(here::here("R", "05_stage1_SDM", "05.1.1_jsoes_bongo_biomass_SDM", "jsoes_bongo_biomass_SDM_output.rda"))
jsoes_bongo_biomass_SDM_Obj <- jsoes_bongo_biomass_SDM_output$jsoes_bongo_biomass_SDM_Obj
jsoes_bongo_biomass_SDM_Opt <- jsoes_bongo_biomass_SDM_output$jsoes_bongo_biomass_SDM_Opt
jsoes_bongo_biomass_SDM_report <- jsoes_bongo_biomass_SDM_output$jsoes_bongo_biomass_SDM_report
# 
# ### Approximate variance for any quantities that weren't ADREPORTed
# dyn.load(dynlib(here::here("R", "05_stage1_SDM", "05.1.1_jsoes_bongo_biomass_SDM", "05_1_jsoes_bongo_biomass_SDM")))
# 
# # estimate variance for overlap metric for csyif x prey field
# SE_pianka_o_csyif_prey_field_t_prey_model = sample_var( obj=jsoes_bongo_biomass_SDM_Obj, var_name="pianka_o_csyif_prey_field_t", mu=jsoes_bongo_biomass_SDM_Obj$env$last.par.best, prec=jsoes_bongo_biomass_SDM_Opt$SD$jointPrecision )
# # estimate variance for overlap metric for cssif x prey field
# SE_pianka_o_cssif_prey_field_t_prey_model = sample_var( obj=jsoes_bongo_biomass_SDM_Obj, var_name="pianka_o_cssif_prey_field_t", mu=jsoes_bongo_biomass_SDM_Obj$env$last.par.best, prec=jsoes_bongo_biomass_SDM_Opt$SD$jointPrecision )
# 
# # Save our estimated SE
# SE_prey_model <- list(SE_pianka_o_csyif_prey_field_t_prey_model = SE_pianka_o_csyif_prey_field_t_prey_model,
#                                               SE_pianka_o_cssif_prey_field_t_prey_model = SE_pianka_o_cssif_prey_field_t_prey_model)
# # 
# save(SE_prey_model,
#      file = here::here("R", "05_stage1_SDM", "05.1.1_jsoes_bongo_biomass_SDM", "estimated_SE_prey_model.rda"))

### identify where in our covariance matrix our ADREPORTed variables are

# figure out where in the cov object our variables of interest are

# first extract the summary of the SD
prey_field_SDM_SD_summary <- summary(jsoes_bongo_biomass_SDM_Opt$SD)

# use this object to figure out what the standard errors are from our variables of interest, and then match them to the cov object
# csyif index of abundance
csyif_index_of_abundance_indices <- which(grepl("csyif_index_of_abundance", rownames(prey_field_SDM_SD_summary)))
csyif_index_of_abundance_cov_indices <- which(sqrt(diag(jsoes_bongo_biomass_SDM_Opt$SD$cov)) %in% prey_field_SDM_SD_summary[csyif_index_of_abundance_indices, "Std. Error"])

# cssif index of abundance
cssif_index_of_abundance_indices <- which(grepl("cssif_index_of_abundance", rownames(prey_field_SDM_SD_summary)))
cssif_index_of_abundance_cov_indices <- which(sqrt(diag(jsoes_bongo_biomass_SDM_Opt$SD$cov)) %in% prey_field_SDM_SD_summary[cssif_index_of_abundance_indices, "Std. Error"])

# prey field
prey_field_index_of_abundance_indices <- which(grepl("prey_field_index_of_abundance", rownames(prey_field_SDM_SD_summary)))
prey_field_index_of_abundance_cov_indices <- which(sqrt(diag(jsoes_bongo_biomass_SDM_Opt$SD$cov)) %in% prey_field_SDM_SD_summary[prey_field_index_of_abundance_indices, "Std. Error"])
```

### 05.1.2_rf_model
```{r}
# If this has already been run before, just load the output instead of running the sample_var code again
load(here::here("R", "05_stage1_SDM", "05.1.2_rf_SDM", "estimated_SE_rf_model.rda"))
SE_pianka_o_csyif_rf_t_rf_model <- SE_rf_model$SE_pianka_o_csyif_rf_t_rf_model
SE_pianka_o_cssif_rf_t_rf_model <- SE_rf_model$SE_pianka_o_cssif_rf_t_rf_model

#### 05.1.2_rf_SDM ####
load(here::here("R", "05_stage1_SDM", "05.1.2_rf_SDM", "rf_SDM_output.rda"))
rf_SDM_Obj <- rf_SDM_output$rf_SDM_Obj
rf_SDM_Opt <- rf_SDM_output$rf_SDM_Opt
rf_SDM_report <- rf_SDM_output$rf_SDM_report
# 
### Approximate variance for any quantities that weren't ADREPORTed
# dyn.load(dynlib(here::here("R", "05_stage1_SDM", "05.1.2_rf_SDM", "05_1_rf_SDM")))
# 
# # estimate variance for overlap metric for csyif x rf
# SE_pianka_o_csyif_rf_t_rf_model = sample_var( obj=rf_SDM_Obj, var_name="pianka_o_csyif_rf_t", mu=rf_SDM_Obj$env$last.par.best, prec=rf_SDM_Opt$SD$jointPrecision )
# # estimate variance for overlap metric for cssif x rf
# SE_pianka_o_cssif_rf_t_rf_model = sample_var( obj=rf_SDM_Obj, var_name="pianka_o_cssif_rf_t", mu=rf_SDM_Obj$env$last.par.best, prec=rf_SDM_Opt$SD$jointPrecision )
# 
# # Save our estimated SE
# SE_rf_model <- list(SE_pianka_o_csyif_rf_t_rf_model = SE_pianka_o_csyif_rf_t_rf_model,
#                                               SE_pianka_o_cssif_rf_t_rf_model = SE_pianka_o_cssif_rf_t_rf_model)
# # 
# save(SE_rf_model,
#      file = here::here("R", "05_stage1_SDM", "05.1.2_rf_SDM", "estimated_SE_rf_model.rda"))

### identify where in our covariance matrix our ADREPORTed variables are

# figure out where in the cov object our variables of interest are

# first extract the summary of the SD
rf_SDM_SD_summary <- summary(rf_SDM_Opt$SD)

# use this object to figure out what the standard errors are from our variables of interest, and then match them to the cov object

# rf
rf_index_of_abundance_indices <- which(grepl("rf_index_of_abundance", rownames(rf_SDM_SD_summary)))
rf_index_of_abundance_cov_indices <- which(sqrt(diag(rf_SDM_Opt$SD$cov)) %in% rf_SDM_SD_summary[rf_index_of_abundance_indices, "Std. Error"])
```

### 05.2_seabird_model
```{r}
# If this has already been run before, just load the output instead of running the sample_var code again
load(here::here("R", "05_stage1_SDM", "05.2_seabird_SDM", "estimated_SE_seabird_model.rda"))
SE_pianka_o_cssif_sosh_t_seabird_model <- SE_seabird_model$SE_pianka_o_cssif_sosh_t_seabird_model
SE_pianka_o_cssif_comu_t_seabird_model <- SE_seabird_model$SE_pianka_o_cssif_comu_t_seabird_model
SE_pianka_o_csyif_sosh_t_seabird_model <- SE_seabird_model$SE_pianka_o_csyif_sosh_t_seabird_model
SE_pianka_o_csyif_comu_t_seabird_model <- SE_seabird_model$SE_pianka_o_csyif_comu_t_seabird_model

# #### 05.2_seabird_SDM ####
load(here::here("R", "05_stage1_SDM", "05.2_seabird_SDM", "seabird_SDM_output.rda"))
seabird_SDM_Obj <- seabird_SDM_output$seabird_SDM_Obj
seabird_SDM_Opt <- seabird_SDM_output$seabird_SDM_Opt
seabird_SDM_report <- seabird_SDM_output$seabird_SDM_report
# 
# ### Approximate variance for any quantities that weren't ADREPORTed
# dyn.load(dynlib(here::here("R", "05_stage1_SDM", "05.2_seabird_SDM", "05_2_seabird_SDM_v2")))
# 
# # estimate variance for overlap metric for cssif x sosh
# SE_pianka_o_cssif_sosh_t_seabird_model = sample_var( obj=seabird_SDM_Obj, var_name="pianka_o_cssif_sosh_t", mu=seabird_SDM_Obj$env$last.par.best, prec=seabird_SDM_Opt$SD$jointPrecision)
# # estimate variance for overlap metric for cssif x comu
# SE_pianka_o_cssif_comu_t_seabird_model = sample_var( obj=seabird_SDM_Obj, var_name="pianka_o_cssif_comu_t", mu=seabird_SDM_Obj$env$last.par.best, prec=seabird_SDM_Opt$SD$jointPrecision)
# 
# # estimate variance for overlap metric for csyif x sosh
# SE_pianka_o_csyif_sosh_t_seabird_model = sample_var( obj=seabird_SDM_Obj, var_name="pianka_o_csyif_sosh_t", mu=seabird_SDM_Obj$env$last.par.best, prec=seabird_SDM_Opt$SD$jointPrecision)
# # estimate variance for overlap metric for csyif x comu
# SE_pianka_o_csyif_comu_t_seabird_model = sample_var( obj=seabird_SDM_Obj, var_name="pianka_o_csyif_comu_t", mu=seabird_SDM_Obj$env$last.par.best, prec=seabird_SDM_Opt$SD$jointPrecision)
# 
# # Save our estimated SE
# SE_seabird_model <- list(SE_pianka_o_cssif_sosh_t_seabird_model = SE_pianka_o_cssif_sosh_t_seabird_model,
#                                               SE_pianka_o_cssif_comu_t_seabird_model = SE_pianka_o_cssif_comu_t_seabird_model,
#                                               SE_pianka_o_csyif_sosh_t_seabird_model = SE_pianka_o_csyif_sosh_t_seabird_model,
#                          SE_pianka_o_csyif_comu_t_seabird_model = SE_pianka_o_csyif_comu_t_seabird_model)
# 
# 
# save(SE_seabird_model,
#      file = here::here("R", "05_stage1_SDM", "05.2_seabird_SDM", "estimated_SE_seabird_model.rda"))



### identify where in our covariance matrix our ADREPORTed variables are

# figure out where in the cov object our variables of interest are

# first extract the summary of the SD
seabird_SDM_SD_summary <- summary(seabird_SDM_Opt$SD)

# use this object to figure out what the standard errors are from our variables of interest, and then match them to the cov object
# comu index of abundance
comu_index_of_abundance_indices <- which(grepl("comu_index_of_abundance", rownames(seabird_SDM_SD_summary)))
comu_index_of_abundance_cov_indices <- which(sqrt(diag(seabird_SDM_Opt$SD$cov)) %in% seabird_SDM_SD_summary[comu_index_of_abundance_indices, "Std. Error"])

sosh_index_of_abundance_indices <- which(grepl("sosh_index_of_abundance", rownames(seabird_SDM_SD_summary)))
sosh_index_of_abundance_cov_indices <- which(sqrt(diag(seabird_SDM_Opt$SD$cov)) %in% seabird_SDM_SD_summary[sosh_index_of_abundance_indices, "Std. Error"])
```



# Fit SAR model - prey and seabirds

### Prepare UC and SR SAR data

```{r prep_SAR_data}
common_years_seabird_prey <- intersect(intersect(unique(csyif$year), # jsoes trawl
                          unique(jsoes_bongo_biomass_cancer_crab_larvae$year)), # jsoes bongo
                          intersect(unique(rf$year), # PRS/PWCC
                          unique(sosh$year)))

common_years_prey <- intersect(
                          unique(jsoes_bongo_biomass_cancer_crab_larvae$year), # jsoes bongo
                          unique(rf$year) # PRS/PWCC
                          )

# load the SAR data
SAR_data <- read.csv(here::here("model_inputs", "chinook_det_hist.csv"))

# subset to only the common years
SAR_data <- subset(SAR_data, run_year %in% common_years_seabird_prey)

# quick look at data
table(SAR_data$run_name, SAR_data$group)

# inspect distribution of outmigration timing

SAR_data %>% 
  mutate(BON_juv_det_date = substr(BON_juv_det_time, 1, 10)) %>% 
  mutate(BON_juv_det_date = ymd(BON_juv_det_date)) %>% 
  mutate(outmigration_date = yday(BON_juv_det_date)) -> SAR_data

# plot distribution of outmigration timing
outmigration_timing_plot <- ggplot(SAR_data, aes(x = outmigration_date)) +
  geom_histogram() +
  facet_wrap(~group + run_name)

ggsave(here::here("figures", "outmigration_timing_plot_by_group.png"), outmigration_timing_plot,
       height = 12, width = 12)


# inspect distribution of years with adult returns
subset(SAR_data, adult_det == 1) -> SAR_adult_det

adult_returns_plot <- ggplot(SAR_adult_det, aes(x = run_year)) +
  geom_histogram(breaks = seq(1998, 2025, 1)) +
  facet_wrap(~group + run_name)

ggsave(here::here("figures", "adult_returns_plot.png"), adult_returns_plot,
       height = 12, width = 12)

SAR_adult_det %>% 
  group_by(group, run_name, run_year) %>% 
  summarise(N = sum(adult_det)) -> adult_det_by_year

#### Snake River Fall Chinook ####

# extract Snake River Fall Chinook from SAR data
SRF <- subset(SAR_data, run_name == "Fall" & group == "Snake River")

#### subset by outmigration date
hist(SRF$outmigration_date)
summary(SRF$outmigration_date)

# restrict the analysis to fish that could have theoretically been caught in the June JSOES survey.
# we will call this April 15 - June 15 (for now!)
# day 105 is April 15 in a non-leap year and day 167 is June 15 in a leap year
# SRF_subset <- filter(SRF, outmigration_date >= 105 & outmigration_date <= 167)

# let's just have a final cutoff - based on the fact that fish can hang out
# off the coast for a while, it's less justified to have a front-end cutoff date
SRF_subset <- filter(SRF, outmigration_date <= 167)
nrow(SRF_subset)/nrow(SRF)
# this leaves us with 54% of our initial dataset


#### Join with the transport data

transport_data_export <- read.csv(here::here("model_inputs", "chinook_transport.csv"))

# join the juvenile transport data with the SAR data
SRF_subset %>% 
  left_join(transport_data_export, join_by(tag_code == tag_id)) -> SRF_subset

SRF_subset$transport[is.na(SRF_subset$transport)] <- 0

# inspect number of transported fish
table(SRF_subset$transport)


#### Drop the few natural origin fish

# fish with rear type U or H are hatchery fish
SRF_subset %>% 
  mutate(rear_numeric = ifelse(rear_type_code %in% c("H", "U"), 1, 0)) -> SRF_subset

# ok, so they're all hatchery. That's perhaps not surprising, given what we see in the JSOES survey is basically all hatchery
# Let's only keep hatchery fish
SRF_subset <- subset(SRF_subset, rear_type_code != "W")


#### Upper Columbia Summer/Fall Chinook ####

# extract Snake River Fall Chinook from SAR data
UCSF <- subset(SAR_data, run_name %in% c("Summer", "Fall") & group == "Upper Columbia")

#### subset by outmigration date
hist(UCSF$outmigration_date)
summary(UCSF$outmigration_date)

# restrict the analysis to fish that could have theoretically been caught in the June JSOES survey.
# we will call this April 15 - June 15 (for now!)
# day 105 is April 15 in a non-leap year and day 167 is June 15 in a leap year
# UCSF_subset <- filter(UCSF, outmigration_date >= 105 & outmigration_date <= 167)

# let's just have a final cutoff - based on the fact that fish can hang out
# off the coast for a while, it's less justified to have a front-end cutoff date
UCSF_subset <- filter(UCSF, outmigration_date <= 167)
nrow(UCSF_subset)/nrow(UCSF)
# this leaves us with 72% of our initial dataset

#### Drop the few natural origin fish (28 of ~97,500)

table(UCSF_subset$rear_type_code)

# fish with rear type U or H are hatchery fish
UCSF_subset %>% 
  mutate(rear_numeric = ifelse(rear_type_code %in% c("H", "U"), 1, 0)) -> UCSF_subset

# ok, so they're all hatchery. That's perhaps not surprising, given what we see in the JSOES survey is basically all hatchery
# Let's only keep hatchery fish
UCSF_subset <- subset(UCSF_subset, rear_type_code != "W")

# set transport to 0 for UCSF fish
UCSF_subset$transport <- 0

# JOIN SRF AND UCSF TOGETHER
bind_rows(SRF_subset, UCSF_subset) -> interior_subset
```

### Set up predictors for SAR model

#### Predictor estimates
```{r}
#### combine all predictors into a data frame ####
## prey model outputs
csyif_index_of_abundance_df <- data.frame(year = sort(unique(csyif$year)),
                                          csyif_index_of_abundance = jsoes_bongo_biomass_SDM_report$csyif_index_of_abundance)
cssif_index_of_abundance_df <- data.frame(year = sort(unique(cssif$year)),
                                          cssif_index_of_abundance = jsoes_bongo_biomass_SDM_report$cssif_index_of_abundance)

# jsoes bongo biomass prey field
prey_field_index_of_abundance_df <- data.frame(year = unique(jsoes_bongo_biomass_cancer_crab_larvae$year),
                                               prey_field_index_of_abundance = jsoes_bongo_biomass_SDM_report$prey_field_index_of_abundance)
pianka_o_csyif_prey_field_df <- data.frame(year = unique(jsoes_bongo_biomass_cancer_crab_larvae$year),
                                           pianka_o_csyif_prey_field_t = jsoes_bongo_biomass_SDM_report$pianka_o_csyif_prey_field_t)
pianka_o_cssif_prey_field_df <- data.frame(year = unique(jsoes_bongo_biomass_cancer_crab_larvae$year),
                                           pianka_o_cssif_prey_field_t = jsoes_bongo_biomass_SDM_report$pianka_o_cssif_prey_field_t)

# rf
rf_index_of_abundance_df <- data.frame(year = unique(rf$year),
                                       rf_index_of_abundance = rf_SDM_report$rf_index_of_abundance)
pianka_o_csyif_rf_df <- data.frame(year = unique(rf$year),
                                   pianka_o_csyif_rf_t = rf_SDM_report$pianka_o_csyif_rf_t)
pianka_o_cssif_rf_df <- data.frame(year = unique(rf$year),
                                   pianka_o_cssif_rf_t = rf_SDM_report$pianka_o_cssif_rf_t)

## seabird model outputs
comu_index_of_abundance_df <- data.frame(year = sort(unique(comu$year)),
                                         comu_index_of_abundance = seabird_SDM_report$comu_index_of_abundance)
sosh_index_of_abundance_df <- data.frame(year = min(sosh$year):max(sosh$year),
                                         sosh_index_of_abundance = seabird_SDM_report$sosh_index_of_abundance)
pianka_o_csyif_comu_df <- data.frame(year = sort(unique(comu$year)),
                                     pianka_o_csyif_comu_t = seabird_SDM_report$pianka_o_csyif_comu_t)
pianka_o_csyif_sosh_df <- data.frame(year = sort(unique(sosh$year)),
                                     pianka_o_csyif_sosh_t = seabird_SDM_report$pianka_o_csyif_sosh_t)
pianka_o_cssif_comu_df <- data.frame(year = sort(unique(comu$year)),
                                     pianka_o_cssif_comu_t = seabird_SDM_report$pianka_o_cssif_comu_t)
pianka_o_cssif_sosh_df <- data.frame(year = sort(unique(sosh$year)),
                                     pianka_o_cssif_sosh_t = seabird_SDM_report$pianka_o_cssif_sosh_t)


csyif_index_of_abundance_df %>% 
  left_join(cssif_index_of_abundance_df, by = "year") %>% 
  left_join(prey_field_index_of_abundance_df, by = "year") %>% 
  left_join(pianka_o_csyif_prey_field_df, by = "year") %>% 
  left_join(pianka_o_cssif_prey_field_df, by = "year") %>%  
  left_join(rf_index_of_abundance_df, by = "year") %>% 
  left_join(pianka_o_csyif_rf_df, by = "year") %>% 
  left_join(pianka_o_cssif_rf_df, by = "year") %>%  
  left_join(comu_index_of_abundance_df, by = "year") %>% 
  left_join(sosh_index_of_abundance_df, by = "year") %>% 
  left_join(pianka_o_csyif_comu_df, by = "year") %>% 
  left_join(pianka_o_csyif_sosh_df, by = "year") %>% 
  left_join(pianka_o_cssif_comu_df, by = "year") %>% 
  left_join(pianka_o_cssif_sosh_df, by = "year") -> SDM_predictors_df
```


#### Predictor uncertainty
```{r}

#### Trim variance terms to common years ####

# APPROXIMATE STANDARD ERRORS
names(SE_pianka_o_csyif_prey_field_t_prey_model) <- unique(jsoes_bongo_biomass_cancer_crab_larvae$year)
names(SE_pianka_o_cssif_prey_field_t_prey_model) <- unique(jsoes_bongo_biomass_cancer_crab_larvae$year)

names(SE_pianka_o_csyif_rf_t_rf_model) <- unique(rf$year)
names(SE_pianka_o_cssif_rf_t_rf_model) <- unique(rf$year)

names(SE_pianka_o_cssif_sosh_t_seabird_model) <- sort(unique(sosh$year))
names(SE_pianka_o_cssif_comu_t_seabird_model) <- sort(unique(sosh$year))
names(SE_pianka_o_csyif_sosh_t_seabird_model) <- sort(unique(sosh$year))
names(SE_pianka_o_csyif_comu_t_seabird_model) <- sort(unique(sosh$year))

SE_pianka_o_csyif_prey_field_t_prey_model_common_years_seabird_prey <- SE_pianka_o_csyif_prey_field_t_prey_model[which(names(SE_pianka_o_csyif_prey_field_t_prey_model) %in% common_years_seabird_prey)]
SE_pianka_o_cssif_prey_field_t_prey_model_common_years_seabird_prey <- SE_pianka_o_cssif_prey_field_t_prey_model[which(names(SE_pianka_o_cssif_prey_field_t_prey_model) %in% common_years_seabird_prey)]

SE_pianka_o_csyif_rf_t_rf_model_common_years_seabird_prey <- SE_pianka_o_csyif_rf_t_rf_model[which(names(SE_pianka_o_csyif_rf_t_rf_model) %in% common_years_seabird_prey)]
SE_pianka_o_cssif_rf_t_rf_model_common_years_seabird_prey <- SE_pianka_o_cssif_rf_t_rf_model[which(names(SE_pianka_o_cssif_rf_t_rf_model) %in% common_years_seabird_prey)]

SE_pianka_o_cssif_sosh_t_seabird_model_common_years_seabird_prey <- SE_pianka_o_cssif_sosh_t_seabird_model[which(names(SE_pianka_o_cssif_sosh_t_seabird_model) %in% common_years_seabird_prey)]
SE_pianka_o_cssif_comu_t_seabird_model_common_years_seabird_prey <- SE_pianka_o_cssif_comu_t_seabird_model[which(names(SE_pianka_o_cssif_comu_t_seabird_model) %in% common_years_seabird_prey)]
SE_pianka_o_csyif_sosh_t_seabird_model_common_years_seabird_prey <- SE_pianka_o_csyif_sosh_t_seabird_model[which(names(SE_pianka_o_csyif_sosh_t_seabird_model) %in% common_years_seabird_prey)]
SE_pianka_o_csyif_comu_t_seabird_model_common_years_seabird_prey <- SE_pianka_o_csyif_comu_t_seabird_model[which(names(SE_pianka_o_csyif_comu_t_seabird_model) %in% common_years_seabird_prey)]



# APPROXIMATE STANDARD ERRORS - as diagonals in covariance matrices
# csyif x prey (JSOES bongo biomass) overlap
pianka_o_csyif_prey_field_cov_matrix <- matrix(data = 0, nrow = length(SE_pianka_o_csyif_prey_field_t_prey_model),
                                               ncol = length(SE_pianka_o_csyif_prey_field_t_prey_model))
diag(pianka_o_csyif_prey_field_cov_matrix) <- SE_pianka_o_csyif_prey_field_t_prey_model^2
rownames(pianka_o_csyif_prey_field_cov_matrix) <- unique(jsoes_bongo_biomass_cancer_crab_larvae$year)
colnames(pianka_o_csyif_prey_field_cov_matrix) <- unique(jsoes_bongo_biomass_cancer_crab_larvae$year)

pianka_o_csyif_prey_field_cov_matrix_common_years_seabird_prey <- pianka_o_csyif_prey_field_cov_matrix[which(rownames(pianka_o_csyif_prey_field_cov_matrix) %in% common_years_seabird_prey), which(colnames(pianka_o_csyif_prey_field_cov_matrix) %in% common_years_seabird_prey)]

# cssif x prey (JSOES bongo biomass) overlap
pianka_o_cssif_prey_field_cov_matrix <- matrix(data = 0, nrow = length(SE_pianka_o_cssif_prey_field_t_prey_model),
                                               ncol = length(SE_pianka_o_cssif_prey_field_t_prey_model))
diag(pianka_o_cssif_prey_field_cov_matrix) <- SE_pianka_o_cssif_prey_field_t_prey_model^2
rownames(pianka_o_cssif_prey_field_cov_matrix) <- unique(jsoes_bongo_biomass_cancer_crab_larvae$year)
colnames(pianka_o_cssif_prey_field_cov_matrix) <- unique(jsoes_bongo_biomass_cancer_crab_larvae$year)

pianka_o_cssif_prey_field_cov_matrix_common_years_seabird_prey <- pianka_o_cssif_prey_field_cov_matrix[which(rownames(pianka_o_cssif_prey_field_cov_matrix) %in% common_years_seabird_prey), which(colnames(pianka_o_cssif_prey_field_cov_matrix) %in% common_years_seabird_prey)]

# csyif x rf overlap
pianka_o_csyif_rf_cov_matrix <- matrix(data = 0, nrow = length(SE_pianka_o_csyif_rf_t_rf_model),
                                       ncol = length(SE_pianka_o_csyif_rf_t_rf_model))
diag(pianka_o_csyif_rf_cov_matrix) <- SE_pianka_o_csyif_rf_t_rf_model^2
rownames(pianka_o_csyif_rf_cov_matrix) <- unique(rf$year)
colnames(pianka_o_csyif_rf_cov_matrix) <- unique(rf$year)

pianka_o_csyif_rf_cov_matrix_common_years_seabird_prey <- pianka_o_csyif_rf_cov_matrix[which(rownames(pianka_o_csyif_rf_cov_matrix) %in% common_years_seabird_prey), which(colnames(pianka_o_csyif_rf_cov_matrix) %in% common_years_seabird_prey)]

# cssif x rf overlap
pianka_o_cssif_rf_cov_matrix <- matrix(data = 0, nrow = length(SE_pianka_o_cssif_rf_t_rf_model),
                                       ncol = length(SE_pianka_o_cssif_rf_t_rf_model))
diag(pianka_o_cssif_rf_cov_matrix) <- SE_pianka_o_cssif_rf_t_rf_model^2
rownames(pianka_o_cssif_rf_cov_matrix) <- unique(rf$year)
colnames(pianka_o_cssif_rf_cov_matrix) <- unique(rf$year)

pianka_o_cssif_rf_cov_matrix_common_years_seabird_prey <- pianka_o_cssif_rf_cov_matrix[which(rownames(pianka_o_cssif_rf_cov_matrix) %in% common_years_seabird_prey), which(colnames(pianka_o_cssif_rf_cov_matrix) %in% common_years_seabird_prey)]

# cssif x sosh overlap
pianka_o_cssif_sosh_cov_matrix <- matrix(data = 0, nrow = length(SE_pianka_o_cssif_sosh_t_seabird_model),
                                         ncol = length(SE_pianka_o_cssif_sosh_t_seabird_model))
diag(pianka_o_cssif_sosh_cov_matrix) <- SE_pianka_o_cssif_sosh_t_seabird_model^2
rownames(pianka_o_cssif_sosh_cov_matrix) <- sort(unique(sosh$year))
colnames(pianka_o_cssif_sosh_cov_matrix) <- sort(unique(sosh$year))

pianka_o_cssif_sosh_cov_matrix_common_years_seabird_prey <- pianka_o_cssif_sosh_cov_matrix[which(rownames(pianka_o_cssif_sosh_cov_matrix) %in% common_years_seabird_prey), which(colnames(pianka_o_cssif_sosh_cov_matrix) %in% common_years_seabird_prey)]

# cssif x comu overlap
pianka_o_cssif_comu_cov_matrix <- matrix(data = 0, nrow = length(SE_pianka_o_cssif_comu_t_seabird_model),
                                         ncol = length(SE_pianka_o_cssif_comu_t_seabird_model))
diag(pianka_o_cssif_comu_cov_matrix) <- SE_pianka_o_cssif_comu_t_seabird_model^2
rownames(pianka_o_cssif_comu_cov_matrix) <- sort(unique(sosh$year))
colnames(pianka_o_cssif_comu_cov_matrix) <- sort(unique(sosh$year))

pianka_o_cssif_comu_cov_matrix_common_years_seabird_prey <- pianka_o_cssif_comu_cov_matrix[which(rownames(pianka_o_cssif_comu_cov_matrix) %in% common_years_seabird_prey), which(colnames(pianka_o_cssif_comu_cov_matrix) %in% common_years_seabird_prey)]

# csyif x sosh overlap
pianka_o_csyif_sosh_cov_matrix <- matrix(data = 0, nrow = length(SE_pianka_o_csyif_sosh_t_seabird_model),
                                         ncol = length(SE_pianka_o_csyif_sosh_t_seabird_model))
diag(pianka_o_csyif_sosh_cov_matrix) <- SE_pianka_o_csyif_sosh_t_seabird_model^2
rownames(pianka_o_csyif_sosh_cov_matrix) <- sort(unique(sosh$year))
colnames(pianka_o_csyif_sosh_cov_matrix) <- sort(unique(sosh$year))

pianka_o_csyif_sosh_cov_matrix_common_years_seabird_prey <- pianka_o_csyif_sosh_cov_matrix[which(rownames(pianka_o_csyif_sosh_cov_matrix) %in% common_years_seabird_prey), which(colnames(pianka_o_csyif_sosh_cov_matrix) %in% common_years_seabird_prey)]

# csyif x comu overlap
pianka_o_csyif_comu_cov_matrix <- matrix(data = 0, nrow = length(SE_pianka_o_csyif_comu_t_seabird_model),
                                         ncol = length(SE_pianka_o_csyif_comu_t_seabird_model))
diag(pianka_o_csyif_comu_cov_matrix) <- SE_pianka_o_csyif_comu_t_seabird_model^2
rownames(pianka_o_csyif_comu_cov_matrix) <- sort(unique(sosh$year))
colnames(pianka_o_csyif_comu_cov_matrix) <- sort(unique(sosh$year))

pianka_o_csyif_comu_cov_matrix_common_years_seabird_prey <- pianka_o_csyif_comu_cov_matrix[which(rownames(pianka_o_csyif_comu_cov_matrix) %in% common_years_seabird_prey), which(colnames(pianka_o_csyif_comu_cov_matrix) %in% common_years_seabird_prey)]




# COVARIANCE MATRICES

#### trim variance terms from prey field model to common years for joint seabird + prey model
csyif_index_of_abundance_cov_matrix <- as.matrix(jsoes_bongo_biomass_SDM_Opt$SD$cov[csyif_index_of_abundance_cov_indices, csyif_index_of_abundance_cov_indices])
colnames(csyif_index_of_abundance_cov_matrix) <- sort(unique(csyif$year))
rownames(csyif_index_of_abundance_cov_matrix) <- sort(unique(csyif$year))

csyif_index_of_abundance_cov_matrix_common_years_seabird_prey <- csyif_index_of_abundance_cov_matrix[which(rownames(csyif_index_of_abundance_cov_matrix) %in% common_years_seabird_prey), which(colnames(csyif_index_of_abundance_cov_matrix) %in% common_years_seabird_prey)]

cssif_index_of_abundance_cov_matrix <- as.matrix(jsoes_bongo_biomass_SDM_Opt$SD$cov[cssif_index_of_abundance_cov_indices, cssif_index_of_abundance_cov_indices])
colnames(cssif_index_of_abundance_cov_matrix) <- sort(unique(cssif$year))
rownames(cssif_index_of_abundance_cov_matrix) <- sort(unique(cssif$year))

cssif_index_of_abundance_cov_matrix_common_years_seabird_prey <- cssif_index_of_abundance_cov_matrix[which(rownames(cssif_index_of_abundance_cov_matrix) %in% common_years_seabird_prey), which(colnames(cssif_index_of_abundance_cov_matrix) %in% common_years_seabird_prey)]

prey_field_index_of_abundance_cov_matrix <- as.matrix(jsoes_bongo_biomass_SDM_Opt$SD$cov[prey_field_index_of_abundance_cov_indices, prey_field_index_of_abundance_cov_indices])
colnames(prey_field_index_of_abundance_cov_matrix) <- unique(jsoes_bongo_biomass_cancer_crab_larvae$year)
rownames(prey_field_index_of_abundance_cov_matrix) <- unique(jsoes_bongo_biomass_cancer_crab_larvae$year)

prey_field_index_of_abundance_cov_matrix_common_years_seabird_prey <- prey_field_index_of_abundance_cov_matrix[which(rownames(prey_field_index_of_abundance_cov_matrix) %in% common_years_seabird_prey), which(colnames(prey_field_index_of_abundance_cov_matrix) %in% common_years_seabird_prey)]

rf_index_of_abundance_cov_matrix <- as.matrix(rf_SDM_Opt$SD$cov[rf_index_of_abundance_cov_indices, rf_index_of_abundance_cov_indices])
colnames(rf_index_of_abundance_cov_matrix) <- unique(rf$year)
rownames(rf_index_of_abundance_cov_matrix) <- unique(rf$year)

rf_index_of_abundance_cov_matrix_common_years_seabird_prey <- rf_index_of_abundance_cov_matrix[which(rownames(rf_index_of_abundance_cov_matrix) %in% common_years_seabird_prey), which(colnames(rf_index_of_abundance_cov_matrix) %in% common_years_seabird_prey)]

#### trim variance terms from seabird model to common years for joint seabird + prey model
comu_index_of_abundance_cov_matrix <- as.matrix(seabird_SDM_Opt$SD$cov[comu_index_of_abundance_cov_indices, comu_index_of_abundance_cov_indices])
colnames(comu_index_of_abundance_cov_matrix) <- sort(unique(comu$year))
rownames(comu_index_of_abundance_cov_matrix) <- sort(unique(comu$year))

comu_index_of_abundance_cov_matrix_common_years_seabird_prey <- comu_index_of_abundance_cov_matrix[which(rownames(comu_index_of_abundance_cov_matrix) %in% common_years_seabird_prey), which(colnames(comu_index_of_abundance_cov_matrix) %in% common_years_seabird_prey)]

sosh_index_of_abundance_cov_matrix <- as.matrix(seabird_SDM_Opt$SD$cov[sosh_index_of_abundance_cov_indices, sosh_index_of_abundance_cov_indices])
colnames(sosh_index_of_abundance_cov_matrix) <- min(sosh$year):max(sosh$year)
rownames(sosh_index_of_abundance_cov_matrix) <- min(sosh$year):max(sosh$year)

sosh_index_of_abundance_cov_matrix_common_years_seabird_prey <- sosh_index_of_abundance_cov_matrix[which(rownames(sosh_index_of_abundance_cov_matrix) %in% common_years_seabird_prey), which(colnames(sosh_index_of_abundance_cov_matrix) %in% common_years_seabird_prey)]



```


### Fit model for interior - z-scored

#### Version 1 - only CSYIF as predictors - but using sample_var estimated SEs as diagonals in covariance matrix
```{r}
# trim the SAR data to only the years of the SDM data
interior_v1_data_prey_seabirds <- subset(interior_subset, run_year %in% common_years_seabird_prey)

# calculate outmigration date separately for UCSF and SRF
mean_SRF_outmigration_date <- mean(subset(interior_v1_data_prey_seabirds, group == "Snake River")$outmigration_date)

mean_UCSF_outmigration_date <- mean(subset(interior_v1_data_prey_seabirds, group == "Upper Columbia")$outmigration_date)

interior_v1_data_prey_seabirds %>% 
  mutate(outmigration_scaled = ifelse(group == "Snake River", outmigration_date - mean_SRF_outmigration_date,
                                      ifelse(group == "Upper Columbia", outmigration_date - mean_UCSF_outmigration_date, NA))) -> interior_v1_data_prey_seabirds

# create new indicators for group x transport
interior_v1_data_prey_seabirds %>% 
  mutate(SRF_transported_i = ifelse(group == "Snake River" & transport == 1, 1, 0)) %>% 
  mutate(SRF_non_transported_i = ifelse(group == "Snake River" & transport == 0, 1, 0)) %>% 
  mutate(UCSF_i = ifelse(group == "Upper Columbia", 1, 0)) -> interior_v1_data_prey_seabirds

# trim the predictors df to only the years in common
interior_v1_seabird_prey_predictors <- subset(SDM_predictors_df, year %in% common_years_seabird_prey)

compile(here::here("R", "06_stage2_SAR", "06.2.6.1_seabird_two_prey_no_salmon_index_one_model_SAR_no_EiV", "06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov.cpp"))
dyn.load(dynlib(here::here("R", "06_stage2_SAR", "06.2.6.1_seabird_two_prey_no_salmon_index_one_model_SAR_no_EiV", "06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov")))

interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Data = list(
  ## SAR data
  "z_i" = interior_v1_data_prey_seabirds$adult_det, 
  "t_i" = as.integer(dplyr::dense_rank(interior_v1_data_prey_seabirds$run_year - min(interior_v1_data_prey_seabirds$run_year))-1),
  "n_t" = length(unique(interior_v1_data_prey_seabirds$run_year)),
  "SRF_transported_i" = interior_v1_data_prey_seabirds$SRF_transported_i,
  "SRF_non_transported_i" = interior_v1_data_prey_seabirds$SRF_non_transported_i,
  "UCSF_i" = interior_v1_data_prey_seabirds$UCSF_i,
  "outmigration_i" = interior_v1_data_prey_seabirds$outmigration_scaled,
  
  ## indices of abundance from stage 1 model
  # "csyif_index_t" = (interior_v1_seabird_prey_predictors$csyif_index_of_abundance - mean(interior_v1_seabird_prey_predictors$csyif_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$csyif_index_of_abundance),
  "prey_field_index_t" = (interior_v1_seabird_prey_predictors$prey_field_index_of_abundance - mean(interior_v1_seabird_prey_predictors$prey_field_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$prey_field_index_of_abundance),
  "rf_index_t" = (interior_v1_seabird_prey_predictors$rf_index_of_abundance - mean(interior_v1_seabird_prey_predictors$rf_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$rf_index_of_abundance),
  "comu_index_t" = (interior_v1_seabird_prey_predictors$comu_index_of_abundance - mean(interior_v1_seabird_prey_predictors$comu_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$comu_index_of_abundance),
  "sosh_index_t" = (interior_v1_seabird_prey_predictors$sosh_index_of_abundance - mean(interior_v1_seabird_prey_predictors$sosh_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$sosh_index_of_abundance),
  
  ## overlap metrics from stage 1 model
  "pianka_o_csyif_prey_field_t" = (interior_v1_seabird_prey_predictors$pianka_o_csyif_prey_field_t - mean(interior_v1_seabird_prey_predictors$pianka_o_csyif_prey_field_t))/sd(interior_v1_seabird_prey_predictors$pianka_o_csyif_prey_field_t),
  "pianka_o_csyif_rf_t" = (interior_v1_seabird_prey_predictors$pianka_o_csyif_rf_t - mean(interior_v1_seabird_prey_predictors$pianka_o_csyif_rf_t))/sd(interior_v1_seabird_prey_predictors$pianka_o_csyif_rf_t),
  "pianka_o_csyif_sosh_t" = (interior_v1_seabird_prey_predictors$pianka_o_csyif_sosh_t - mean(interior_v1_seabird_prey_predictors$pianka_o_csyif_sosh_t))/sd(interior_v1_seabird_prey_predictors$pianka_o_csyif_sosh_t),
  "pianka_o_csyif_comu_t" = (interior_v1_seabird_prey_predictors$pianka_o_csyif_comu_t - mean(interior_v1_seabird_prey_predictors$pianka_o_csyif_comu_t))/sd(interior_v1_seabird_prey_predictors$pianka_o_csyif_comu_t)
)

interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Params <- list(
  ## SAR parameters
  "beta_UCSF" = 0, # intercept term for UCSF
  "beta_SRF_transported" = 0, # intercept term for SRF transported
  "beta_SRF_non_transported" = 0, # intercept term for SRF non-transported
  
  "beta_outmigration_SRF" = 0,
  "beta_outmigration_SRF2" = 0,
  "beta_outmigration_UCSF" = 0,
  "beta_outmigration_UCSF2" = 0,
  
  # "beta_csyif_index" = 0, # term to account for abundance in the survey
  
  # parameters for the influence of abundance of prey/predators
  "beta_prey_field_index" = 0,
  "beta_rf_index" = 0,
  "beta_comu_index" = 0,
  "beta_sosh_index" = 0,
  
  # parameters for the influence of overlap with prey/predators
  "beta_ov_prey_field" = 0,
  "beta_ov_rf" = 0,
  "beta_ov_comu" = 0,
  "beta_ov_sosh" = 0
  
)

# save workspace, in case model crashes
# save.image(here::here("R", "06_stage2_SAR", "06.2.6.1_seabird_two_prey_no_salmon_index_one_model_SAR_no_EiV", "06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_workspace.rda"))
# load("hake_SAR_workspace.RData")

interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Obj = MakeADFun( data=interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Data, 
                                                              parameters=interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Params, 
                                                              random= NULL,
                                                              DLL = "06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov")


# Optimize
interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Opt = nlminb( start=interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Obj$par, obj=interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Obj$fn, grad=interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Obj$gr )
interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Opt$SD = sdreport( interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Obj)
interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_report = interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Obj$report()

interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_output <- list(interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Obj = interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Obj,
                                                            interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Opt = interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_Opt,
                                                            interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_report = interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_report)


save(interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_output,
     file = here::here("R", "06_stage2_SAR", "06.2.6.1_seabird_two_prey_no_salmon_index_one_model_SAR_no_EiV", "interior_06_2_6_1_seabird_prey_csyif_only_one_model_SAR_no_EiV_SEcov_output_zscored.rda"))
```

#### Version 2 - only CSSIF as predictors - but using sample_var estimated SEs as diagonals in covariance matrix
```{r}
# trim the SAR data to only the years of the SDM data
interior_v1_data_prey_seabirds <- subset(interior_subset, run_year %in% common_years_seabird_prey)

# calculate outmigration date separately for UCSF and SRF
mean_SRF_outmigration_date <- mean(subset(interior_v1_data_prey_seabirds, group == "Snake River")$outmigration_date)

mean_UCSF_outmigration_date <- mean(subset(interior_v1_data_prey_seabirds, group == "Upper Columbia")$outmigration_date)

interior_v1_data_prey_seabirds %>% 
  mutate(outmigration_scaled = ifelse(group == "Snake River", outmigration_date - mean_SRF_outmigration_date,
                                      ifelse(group == "Upper Columbia", outmigration_date - mean_UCSF_outmigration_date, NA))) -> interior_v1_data_prey_seabirds

# create new indicators for group x transport
interior_v1_data_prey_seabirds %>% 
  mutate(SRF_transported_i = ifelse(group == "Snake River" & transport == 1, 1, 0)) %>% 
  mutate(SRF_non_transported_i = ifelse(group == "Snake River" & transport == 0, 1, 0)) %>% 
  mutate(UCSF_i = ifelse(group == "Upper Columbia", 1, 0)) -> interior_v1_data_prey_seabirds


# trim the predictors df to only the years in common
interior_v1_seabird_prey_predictors <- subset(SDM_predictors_df, year %in% common_years_seabird_prey)

compile(here::here("R", "06_stage2_SAR", "06.2.6.1_seabird_two_prey_no_salmon_index_one_model_SAR_no_EiV", "06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov.cpp"))
dyn.load(dynlib(here::here("R", "06_stage2_SAR", "06.2.6.1_seabird_two_prey_no_salmon_index_one_model_SAR_no_EiV", "06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov")))

interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Data = list(
  ## SAR data
  "z_i" = interior_v1_data_prey_seabirds$adult_det, 
  "t_i" = as.integer(dplyr::dense_rank(interior_v1_data_prey_seabirds$run_year - min(interior_v1_data_prey_seabirds$run_year))-1),
  "n_t" = length(unique(interior_v1_data_prey_seabirds$run_year)),
  "SRF_transported_i" = interior_v1_data_prey_seabirds$SRF_transported_i,
  "SRF_non_transported_i" = interior_v1_data_prey_seabirds$SRF_non_transported_i,
  "UCSF_i" = interior_v1_data_prey_seabirds$UCSF_i,
  "outmigration_i" = interior_v1_data_prey_seabirds$outmigration_scaled,
  
  ## indices of abundance from stage 1 model
  # "cssif_index_t" = (interior_v1_seabird_prey_predictors$cssif_index_of_abundance - mean(interior_v1_seabird_prey_predictors$cssif_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$cssif_index_of_abundance),
  "prey_field_index_t" = (interior_v1_seabird_prey_predictors$prey_field_index_of_abundance - mean(interior_v1_seabird_prey_predictors$prey_field_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$prey_field_index_of_abundance),
  "rf_index_t" = (interior_v1_seabird_prey_predictors$rf_index_of_abundance - mean(interior_v1_seabird_prey_predictors$rf_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$rf_index_of_abundance),
  "comu_index_t" = (interior_v1_seabird_prey_predictors$comu_index_of_abundance - mean(interior_v1_seabird_prey_predictors$comu_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$comu_index_of_abundance),
  "sosh_index_t" = (interior_v1_seabird_prey_predictors$sosh_index_of_abundance - mean(interior_v1_seabird_prey_predictors$sosh_index_of_abundance))/sd(interior_v1_seabird_prey_predictors$sosh_index_of_abundance),
  
  ## overlap metrics from stage 1 model
  "pianka_o_cssif_prey_field_t" = (interior_v1_seabird_prey_predictors$pianka_o_cssif_prey_field_t - mean(interior_v1_seabird_prey_predictors$pianka_o_cssif_prey_field_t))/sd(interior_v1_seabird_prey_predictors$pianka_o_cssif_prey_field_t),
  "pianka_o_cssif_rf_t" = (interior_v1_seabird_prey_predictors$pianka_o_cssif_rf_t - mean(interior_v1_seabird_prey_predictors$pianka_o_cssif_rf_t))/sd(interior_v1_seabird_prey_predictors$pianka_o_cssif_rf_t),
  "pianka_o_cssif_sosh_t" = (interior_v1_seabird_prey_predictors$pianka_o_cssif_sosh_t - mean(interior_v1_seabird_prey_predictors$pianka_o_cssif_sosh_t))/sd(interior_v1_seabird_prey_predictors$pianka_o_cssif_sosh_t),
  "pianka_o_cssif_comu_t" = (interior_v1_seabird_prey_predictors$pianka_o_cssif_comu_t - mean(interior_v1_seabird_prey_predictors$pianka_o_cssif_comu_t))/sd(interior_v1_seabird_prey_predictors$pianka_o_cssif_comu_t)
)

interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Params <- list(
  ## SAR parameters
  "beta_UCSF" = 0, # intercept term for UCSF
  "beta_SRF_transported" = 0, # intercept term for SRF transported
  "beta_SRF_non_transported" = 0, # intercept term for SRF non-transported
  
  "beta_outmigration_SRF" = 0,
  "beta_outmigration_SRF2" = 0,
  "beta_outmigration_UCSF" = 0,
  "beta_outmigration_UCSF2" = 0,
  
  # "beta_cssif_index" = 0, # term to account for abundance in the survey
  
  # parameters for the influence of abundance of prey/predators
  "beta_prey_field_index" = 0,
  "beta_rf_index" = 0,
  "beta_comu_index" = 0,
  "beta_sosh_index" = 0,
  
  # parameters for the influence of overlap with prey/predators
  "beta_ov_prey_field" = 0,
  "beta_ov_rf" = 0,
  "beta_ov_comu" = 0,
  "beta_ov_sosh" = 0
  
)

# save workspace, in case model crashes
# save.image(here::here("R", "06_stage2_SAR", "06.2.6.1_seabird_two_prey_no_salmon_index_one_model_SAR_no_EiV", "06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_workspace.rda"))
# load("hake_SAR_workspace.RData")

interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Obj = MakeADFun( data=interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Data, 
                                                              parameters=interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Params, 
                                                              random= NULL,
                                                              DLL = "06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov")


# Optimize
interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Opt = nlminb( start=interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Obj$par, obj=interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Obj$fn, grad=interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Obj$gr )
interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Opt$SD = sdreport( interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Obj)
interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_report = interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Obj$report()

interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_output <- list(interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Obj = interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Obj,
                                                            interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Opt = interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_Opt,
                                                            interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_report = interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_report)


save(interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_output,
     file = here::here("R", "06_stage2_SAR", "06.2.6.1_seabird_two_prey_no_salmon_index_one_model_SAR_no_EiV", "interior_06_2_6_1_seabird_prey_cssif_only_one_model_SAR_no_EiV_SEcov_output_zscored.rda"))
```
