---
title: "04.6.1_non_cancer_crab_larvae_biomass_SDM"
author: "Markus Min"
date: "`r Sys.Date()`"
output: html_document
---

## Description

In this document, we conduct the model selection process for Cancer crab larvae from the JSOES Bongo data. We identify the best fit model, which is then used to model the distribution of Cancer crab larvae in the joint distribution models found in the folder `/05_stage1_SDM`.


### Model selection steps

1. Fit a "base" model, that includes all fixed effects but no random effects.

2. Use this base model to tune the structure of the spatial and spatiotemporal random effects using cAIC as the information criterion.

3. Keeping the random effects structure identified in step 2 fixed, identify the fixed effect structure for our covariates (year, temperature, distance from shore), again using cAIC as the information criterion. This includes exploring two alternative functional forms for temperature and distance from shore: linear and p-splines.

```{r chunk-options, message = FALSE, echo = FALSE, warning = FALSE}
# set chunk options
knitr::opts_chunk$set(message = FALSE, echo = FALSE, warning = FALSE)
```


## Data preparation for model selection process

```{r load_libraries_source_scripts, include = FALSE}
library(tidyverse)
library(here)
library(TMB)
library(ggforce)
library(sdmTMB)
library(kableExtra)
library(mgcv)

# source the function scripts
source("R/functions/cAIC.R")
source("R/functions/rmvnorm_prec.R")
source("R/functions/sample_var.R")
source("R/functions/plot_anisotropy_MM.R")
source("R/functions/make_anisotropy_spde.R")
source("R/functions/H_matrix_prior_predictive_check_for_ggplot.R")
source("R/functions/plot_distribution.R")
source("R/functions/plot_distribution_jsoes_bongo.R")
source("R/functions/generate_prediction_maps.R")
source("R/functions/sdmTMB_smoothing_functions.R")

# source the script to load spatial data and create the mesh
source(here::here("R", "JSOES_make_mesh.R"))
# note that this script also does some processing of the trawl data
```

```{r data_prep}
#### load and reformat bongo data ####

jsoes_bongo_biomass_cancer_crab_larvae <- read.csv(here::here("model_inputs", "jsoes_bongo_biomass_cancer_crab_larvae.csv"))
jsoes_bongo_biomass_hyperiid_amphipods <- read.csv(here::here("model_inputs", "jsoes_bongo_biomass_hyperiid_amphipod.csv"))
jsoes_bongo_biomass_non_cancer_crab_larvae <- read.csv(here::here("model_inputs", "jsoes_bongo_biomass_non_cancer_crab_larvae.csv"))
jsoes_bongo_biomass_shrimp_larvae <- read.csv(here::here("model_inputs", "jsoes_bongo_biomass_shrimp_larvae.csv"))

# drop 1998 (bongo doesn't have this year) and 2022:2025 (adult returns aren't complete for those years yet)
jsoes_bongo_biomass_cancer_crab_larvae <- subset(jsoes_bongo_biomass_cancer_crab_larvae, !(year %in% c(1998, 2022:2025)))
jsoes_bongo_biomass_hyperiid_amphipods <- subset(jsoes_bongo_biomass_hyperiid_amphipods, !(year %in% c(1998, 2022:2025)))
jsoes_bongo_biomass_non_cancer_crab_larvae <- subset(jsoes_bongo_biomass_non_cancer_crab_larvae, !(year %in% c(1998, 2022:2025)))
jsoes_bongo_biomass_shrimp_larvae <- subset(jsoes_bongo_biomass_shrimp_larvae, !(year %in% c(1998, 2022:2025)))



# add a nmi and km from shore from shore column based on the station codes
add_nmi_km <- function(data){
  data %>% 
    mutate(nmi_from_shore = parse_number(station)) %>% 
    # change nmi to km to keep units consistent
    mutate(km_from_shore = nmi_from_shore * 1.852) -> output
  
  return(output)
}

jsoes_bongo_biomass_cancer_crab_larvae <- add_nmi_km(jsoes_bongo_biomass_cancer_crab_larvae)
jsoes_bongo_biomass_hyperiid_amphipods <- add_nmi_km(jsoes_bongo_biomass_hyperiid_amphipods)
jsoes_bongo_biomass_non_cancer_crab_larvae <- add_nmi_km(jsoes_bongo_biomass_non_cancer_crab_larvae)
jsoes_bongo_biomass_shrimp_larvae <- add_nmi_km(jsoes_bongo_biomass_shrimp_larvae)

# the other script already deals with the crs, and the X and Y columns contain
# the UTM zone 10 distances in km


#### load and reformat spatial data ####
usa_spdf <- st_read(here::here("Data", "map_files", "USA_adm0.shp"))
# load BC
CAN_spdf <- st_read(here::here("Data", "map_files", "canada", "lpr_000b16a_e.shp"))
BC_spdf <- filter(CAN_spdf, PRENAME == "British Columbia")
BC_proj <- st_transform(BC_spdf, crs = 4326)


# crop them to our desired area
US_west_coast <- sf::st_crop(usa_spdf,
                             c(xmin = -126, ymin = 44, xmax = -123, ymax = 48.5))

BC_coast <- sf::st_crop(BC_proj,
                        c(xmin = -126, ymin = 44, xmax = -123, ymax = 48.5))



# convert both shapefiles to a different projection (UTM zone 10) so that they can be plotted with the sdmTMB output
UTM_zone_10_crs <- 32610

US_west_coast_proj <- sf::st_transform(US_west_coast, crs = UTM_zone_10_crs)
BC_coast_proj <- sf::st_transform(BC_coast, crs = UTM_zone_10_crs)

# make this projection into kilometers
US_west_coast_proj_km <- st_as_sf(US_west_coast_proj$geometry/1000, crs = UTM_zone_10_crs)
BC_coast_proj_km <- st_as_sf(BC_coast_proj$geometry/1000, crs = UTM_zone_10_crs)



#### load and reformat SST data ####
SST1 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_9bdb_17ce_0aa0.csv"), skip = 1)
SST2 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_36e3_7cb3_1109.csv"), skip = 1)
SST3 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_519c_f955_0fea.csv"), skip = 1)
SST4 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_727c_5c10_e6a0.csv"), skip = 1)
SST5 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_1176_fc5f_53d2.csv"), skip = 1)
SST6 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_ae0c_9197_083f.csv"), skip = 1)
SST7 <- read.csv(here::here("Data", "NOAA_SST", "noaacrwsstDaily_fe4f_9c9c_442e.csv"), skip = 1)

SST1 %>% 
  bind_rows(., SST2, SST3, SST4, SST5, SST6, SST7) -> SST

SST %>% 
  dplyr::rename(time = UTC, latitude = degrees_north, longitude = degrees_east, SST = degree_C) -> SST
SST %>% 
  mutate(time = ymd_hms(time)) %>% 
  mutate(day_of_month = mday(time)) %>% 
  mutate(year = year(time)) %>% 
  mutate(julian = yday(time))-> SST

# Looks good to me! Now just need to resolve CRS conflicts

# Convert SST to UTM zone 10
# st_as_sf(SST, coords = c("latitude", "longitude"), crs = "4326") -> SST_sf
st_as_sf(SST, coords = c("longitude", "latitude"), crs = 4326) -> SST_sf
US_west_coast

sf::st_transform(SST_sf, crs = UTM_zone_10_crs) -> SST_sf_proj

# convert to km
SST_sf_proj_km <- SST_sf_proj
SST_sf_proj_km$geometry <- SST_sf_proj$geometry/1000

SST_sf_proj_km <- st_set_crs(SST_sf_proj_km, UTM_zone_10_crs)

# Match this SST data to the JSOES bongo data
# jsoes needs to be in SF format as well

st_as_sf(jsoes_bongo_biomass_cancer_crab_larvae, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_cancer_crab_larvae_sf
st_as_sf(jsoes_bongo_biomass_hyperiid_amphipods, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_hyperiid_amphipods_sf
st_as_sf(jsoes_bongo_biomass_non_cancer_crab_larvae, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_non_cancer_crab_larvae_sf
st_as_sf(jsoes_bongo_biomass_shrimp_larvae, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_shrimp_larvae_sf

jsoes_bongo_biomass_cancer_crab_larvae_sf_split <- split(jsoes_bongo_biomass_cancer_crab_larvae_sf, jsoes_bongo_biomass_cancer_crab_larvae_sf$year)
jsoes_bongo_biomass_hyperiid_amphipods_sf_split <- split(jsoes_bongo_biomass_hyperiid_amphipods_sf, jsoes_bongo_biomass_hyperiid_amphipods_sf$year)
jsoes_bongo_biomass_non_cancer_crab_larvae_sf_split <- split(jsoes_bongo_biomass_non_cancer_crab_larvae_sf, jsoes_bongo_biomass_non_cancer_crab_larvae_sf$year)
jsoes_bongo_biomass_shrimp_larvae_sf_split <- split(jsoes_bongo_biomass_shrimp_larvae_sf, jsoes_bongo_biomass_shrimp_larvae_sf$year)
SST_sf_proj_km_split <- split(SST_sf_proj_km, SST_sf_proj_km$year) 

jsoes_years <- as.character(unique(jsoes_bongo_biomass_cancer_crab_larvae_sf$year))

jsoes_bongo_biomass_cancer_crab_larvae_remote_SST_list <- map(jsoes_years, function(year) {
  jsoes_one_year <- jsoes_bongo_biomass_cancer_crab_larvae_sf_split[[year]]
  SST_one_year <- SST_sf_proj_km_split[[year]]
  
  # Find nearest features
  SST_index_for_jsoes <- st_nearest_feature(jsoes_one_year, SST_one_year)
  
  # Combine with matched points (optional: include matched geometry or attributes)
  jsoes_one_year %>%
    mutate(matched_id = SST_index_for_jsoes) %>%
    bind_cols(
      st_drop_geometry(SST_one_year[SST_index_for_jsoes, "SST"]) %>%
        dplyr::rename(remote_SST = SST)
    ) %>% 
    dplyr::select(-matched_id) -> output
  
  # transform this to a df
  st_drop_geometry(output) %>% 
    bind_cols(st_coordinates(output)) -> output
  
  return(output)
})

# take the list and turn it back into an sf object
jsoes_bongo_biomass_cancer_crab_larvae <- list_rbind(jsoes_bongo_biomass_cancer_crab_larvae_remote_SST_list)

jsoes_bongo_biomass_hyperiid_amphipods_sf_split_remote_SST_list <- map(jsoes_years, function(year) {
  jsoes_one_year <- jsoes_bongo_biomass_hyperiid_amphipods_sf_split[[year]]
  SST_one_year <- SST_sf_proj_km_split[[year]]
  
  # Find nearest features
  SST_index_for_jsoes <- st_nearest_feature(jsoes_one_year, SST_one_year)
  
  # Combine with matched points (optional: include matched geometry or attributes)
  jsoes_one_year %>%
    mutate(matched_id = SST_index_for_jsoes) %>%
    bind_cols(
      st_drop_geometry(SST_one_year[SST_index_for_jsoes, "SST"]) %>%
        dplyr::rename(remote_SST = SST)
    ) %>% 
    dplyr::select(-matched_id) -> output
  
  # transform this to a df
  st_drop_geometry(output) %>% 
    bind_cols(st_coordinates(output)) -> output
  
  return(output)
})

# take the list and turn it back into an sf object
jsoes_bongo_biomass_hyperiid_amphipods <- list_rbind(jsoes_bongo_biomass_hyperiid_amphipods_sf_split_remote_SST_list)


jsoes_bongo_biomass_non_cancer_crab_larvae_remote_SST_list <- map(jsoes_years, function(year) {
  jsoes_one_year <- jsoes_bongo_biomass_non_cancer_crab_larvae_sf_split[[year]]
  SST_one_year <- SST_sf_proj_km_split[[year]]
  
  # Find nearest features
  SST_index_for_jsoes <- st_nearest_feature(jsoes_one_year, SST_one_year)
  
  # Combine with matched points (optional: include matched geometry or attributes)
  jsoes_one_year %>%
    mutate(matched_id = SST_index_for_jsoes) %>%
    bind_cols(
      st_drop_geometry(SST_one_year[SST_index_for_jsoes, "SST"]) %>%
        dplyr::rename(remote_SST = SST)
    ) %>% 
    dplyr::select(-matched_id) -> output
  
  # transform this to a df
  st_drop_geometry(output) %>% 
    bind_cols(st_coordinates(output)) -> output
  
  return(output)
})

# take the list and turn it back into an sf object
jsoes_bongo_biomass_non_cancer_crab_larvae <- list_rbind(jsoes_bongo_biomass_non_cancer_crab_larvae_remote_SST_list)


jsoes_bongo_biomass_shrimp_larvae_remote_SST_list <- map(jsoes_years, function(year) {
  jsoes_one_year <- jsoes_bongo_biomass_shrimp_larvae_sf_split[[year]]
  SST_one_year <- SST_sf_proj_km_split[[year]]
  
  # Find nearest features
  SST_index_for_jsoes <- st_nearest_feature(jsoes_one_year, SST_one_year)
  
  # Combine with matched points (optional: include matched geometry or attributes)
  jsoes_one_year %>%
    mutate(matched_id = SST_index_for_jsoes) %>%
    bind_cols(
      st_drop_geometry(SST_one_year[SST_index_for_jsoes, "SST"]) %>%
        dplyr::rename(remote_SST = SST)
    ) %>% 
    dplyr::select(-matched_id) -> output
  
  # transform this to a df
  st_drop_geometry(output) %>% 
    bind_cols(st_coordinates(output)) -> output
  
  return(output)
})

# take the list and turn it back into an sf object
jsoes_bongo_biomass_shrimp_larvae <- list_rbind(jsoes_bongo_biomass_shrimp_larvae_remote_SST_list)

#### Calculate distance to shore ####

### cancer crab larvae
dplyr::select(jsoes_bongo_biomass_cancer_crab_larvae, station, km_from_shore, X, Y) -> jsoes_bongo_biomass_cancer_crab_larvae_spatial
st_as_sf(jsoes_bongo_biomass_cancer_crab_larvae_spatial, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_cancer_crab_larvae_sf
st_distance(jsoes_bongo_biomass_cancer_crab_larvae_sf, US_west_coast_proj_km) -> jsoes_bongo_biomass_cancer_crab_larvae_dist_shore
jsoes_bongo_biomass_cancer_crab_larvae$sf_dist_shore <- as.numeric(jsoes_bongo_biomass_cancer_crab_larvae_dist_shore)

# confirm that these are reasonable, based on points from survey
# they're similar - but not identical
hist(as.numeric(jsoes_bongo_biomass_cancer_crab_larvae$sf_dist_shore)-jsoes_bongo_biomass_cancer_crab_larvae_spatial$km_from_shore)
# and they're pretty biased - the distances calculated using sf tend to be smaller
# Let's just use the sf distance from shore calculated for both prediction grid and
# the samples themselves
# Let's also use the SST from remote sensing, instead of the in situ data


# rescale data
jsoes_bongo_biomass_cancer_crab_larvae %>% 
  mutate(SST_scaled = as.numeric(scale(remote_SST)),
         dist_shore_scaled = as.numeric(scale(sf_dist_shore))) -> jsoes_bongo_biomass_cancer_crab_larvae

### hyperiid amphipods
dplyr::select(jsoes_bongo_biomass_hyperiid_amphipods, station, km_from_shore, X, Y) -> jsoes_bongo_biomass_hyperiid_amphipods_spatial
st_as_sf(jsoes_bongo_biomass_hyperiid_amphipods_spatial, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_hyperiid_amphipods_sf
st_distance(jsoes_bongo_biomass_hyperiid_amphipods_sf, US_west_coast_proj_km) -> jsoes_bongo_biomass_hyperiid_amphipods_dist_shore
jsoes_bongo_biomass_hyperiid_amphipods$sf_dist_shore <- as.numeric(jsoes_bongo_biomass_hyperiid_amphipods_dist_shore)

# confirm that these are reasonable, based on points from survey
# they're similar - but not identical
hist(as.numeric(jsoes_bongo_biomass_hyperiid_amphipods$sf_dist_shore)-jsoes_bongo_biomass_hyperiid_amphipods_spatial$km_from_shore)
# and they're pretty biased - the distances calculated using sf tend to be smaller
# Let's just use the sf distance from shore calculated for both prediction grid and
# the samples themselves
# Let's also use the SST from remote sensing, instead of the in situ data


# rescale data
jsoes_bongo_biomass_hyperiid_amphipods %>% 
  mutate(SST_scaled = as.numeric(scale(remote_SST)),
         dist_shore_scaled = as.numeric(scale(sf_dist_shore))) -> jsoes_bongo_biomass_hyperiid_amphipods

### non-cancer crab larvae
dplyr::select(jsoes_bongo_biomass_non_cancer_crab_larvae, station, km_from_shore, X, Y) -> jsoes_bongo_biomass_non_cancer_crab_larvae_spatial
st_as_sf(jsoes_bongo_biomass_non_cancer_crab_larvae_spatial, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_non_cancer_crab_larvae_sf
st_distance(jsoes_bongo_biomass_non_cancer_crab_larvae_sf, US_west_coast_proj_km) -> jsoes_bongo_biomass_non_cancer_crab_larvae_dist_shore
jsoes_bongo_biomass_non_cancer_crab_larvae$sf_dist_shore <- as.numeric(jsoes_bongo_biomass_non_cancer_crab_larvae_dist_shore)

# confirm that these are reasonable, based on points from survey
# they're similar - but not identical
hist(as.numeric(jsoes_bongo_biomass_non_cancer_crab_larvae$sf_dist_shore)-jsoes_bongo_biomass_non_cancer_crab_larvae_spatial$km_from_shore)
# and they're pretty biased - the distances calculated using sf tend to be smaller
# Let's just use the sf distance from shore calculated for both prediction grid and
# the samples themselves
# Let's also use the SST from remote sensing, instead of the in situ data


# rescale data
jsoes_bongo_biomass_non_cancer_crab_larvae %>% 
  mutate(SST_scaled = as.numeric(scale(remote_SST)),
         dist_shore_scaled = as.numeric(scale(sf_dist_shore))) -> jsoes_bongo_biomass_non_cancer_crab_larvae


### shrimp larvae
dplyr::select(jsoes_bongo_biomass_shrimp_larvae, station, km_from_shore, X, Y) -> jsoes_bongo_biomass_shrimp_larvae_spatial
st_as_sf(jsoes_bongo_biomass_shrimp_larvae_spatial, coords = c("X", "Y"), crs = UTM_zone_10_crs) -> jsoes_bongo_biomass_shrimp_larvae_sf
st_distance(jsoes_bongo_biomass_shrimp_larvae_sf, US_west_coast_proj_km) -> jsoes_bongo_biomass_shrimp_larvae_dist_shore
jsoes_bongo_biomass_shrimp_larvae$sf_dist_shore <- as.numeric(jsoes_bongo_biomass_shrimp_larvae_dist_shore)

# confirm that these are reasonable, based on points from survey
# they're similar - but not identical
hist(as.numeric(jsoes_bongo_biomass_shrimp_larvae$sf_dist_shore)-jsoes_bongo_biomass_shrimp_larvae_spatial$km_from_shore)
# and they're pretty biased - the distances calculated using sf tend to be smaller
# Let's just use the sf distance from shore calculated for both prediction grid and
# the samples themselves
# Let's also use the SST from remote sensing, instead of the in situ data


# rescale data
jsoes_bongo_biomass_shrimp_larvae %>% 
  mutate(SST_scaled = as.numeric(scale(remote_SST)),
         dist_shore_scaled = as.numeric(scale(sf_dist_shore))) -> jsoes_bongo_biomass_shrimp_larvae


# for SST - take prediction grid, drop distance from shore, add SST with locations as rows and years as column
survey_predict_grid %>% 
  # select only the years that we're keeping
  filter(!(year %in% c(1998, 2022:2025))) %>% 
  dplyr::select(-c(SST, sal, sal_scaled, dist_shore, dist_shore_scaled)) %>% 
  pivot_wider(names_from = year, values_from = SST_scaled) %>% 
  dplyr::select(-c(X, Y)) %>% 
  as.matrix() -> temperature_projection_matrix


distance_projection_vector <- as.numeric(filter(survey_predict_grid, year == 2010)$dist_shore_scaled)

#### Create SPDE objects ####

bongo_biomass_spde <- fm_fem( bongo_inla_mesh_cutoff15, 
                      order = 2 )

# create projection matrix from vertices to samples
A_is_non_cancer_crab_larvae = fm_evaluator(bongo_inla_mesh_cutoff15, loc = as.matrix(data.frame(X = jsoes_bongo_biomass_non_cancer_crab_larvae$X, Y = jsoes_bongo_biomass_non_cancer_crab_larvae$Y)))$proj$A

# create projection matrix from vertices to prediction grid
A_gs_non_cancer_crab_larvae = fm_evaluator( bongo_inla_mesh_cutoff15, loc=st_coordinates(survey_domain_cov_grid))$proj$A

# prep anisotropy input
# note that spatial list replaces the M0, M1, M2 objects 
spatial_list_non_cancer_crab_larvae <- make_anisotropy_spde( bongo_inla_mesh_cutoff15 )
spatial_list <- make_anisotropy_spde( inla_mesh_cutoff15 )
```
The JSOES Bongo data for this taxon is visualized below:

```{r plot_distribtion, fig.cap = "Density of non-Cancer crab larvae caught in the JSOES Bongol.", fig.height = 8, fig.width = 12}
non_cancer_crab_larvae_june_biomass_plot <- plot_distribution_jsoes_bongo_biomass(data = jsoes_bongo_biomass_non_cancer_crab_larvae, taxon_name = "Non-Cancer Crab Larvae")

ggsave(here::here("figures",  "non_cancer_crab_larvae_june_biomass_plot.png"), non_cancer_crab_larvae_june_biomass_plot,  
       height = 10, width = 16)

non_cancer_crab_larvae_june_biomass_plot
```


# Model selection process

## Step 1: Fit the base model with saturated fixed effects

We will first fit a model with all fixed effects included: temperature, distance from shore, and year.

- Year will be modeled as a factor
- Temperature and distance from shore will be modeled as linear predictors.

```{r fit_saturated_fixed_effects, cache = TRUE, include = FALSE}
# save.image("non_cancer_crab_larvae_model_selection.RData")
# load("shrimp_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "candidate_models", "saturated_fixed_effects_year_temp_dist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models", "saturated_fixed_effects_year_temp_dist")))

saturated_fixed_effects_year_temp_dist_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  # "spatial_list" = spatial_list_non_cancer_crab_larvae
)


saturated_fixed_effects_year_temp_dist_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  # "ln_tau_omega"=0, # "ln_tau_epsilon"=0,
  # "ln_kappa"=0, 
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(bongo_biomass_spde$c0))
  # "epsilon_st_b"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=saturated_fixed_effects_year_temp_dist_Data$n_t)
  # 
)

saturated_fixed_effects_year_temp_dist_Obj = MakeADFun( data=saturated_fixed_effects_year_temp_dist_Data,
                                                        parameters=saturated_fixed_effects_year_temp_dist_Params,
                                                        # random= c("omega_s"),
                                                        DLL = "saturated_fixed_effects_year_temp_dist")

# Optimize
saturated_fixed_effects_year_temp_dist_Opt = nlminb( start=saturated_fixed_effects_year_temp_dist_Obj$par, obj=saturated_fixed_effects_year_temp_dist_Obj$fn, grad=saturated_fixed_effects_year_temp_dist_Obj$gr )
# add getJointPrecision for index
# there are no random effects so turn off bias.correct and getJointPrecision
# saturated_fixed_effects_year_temp_dist_Opt$SD = sdreport( saturated_fixed_effects_year_temp_dist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
saturated_fixed_effects_year_temp_dist_Opt$SD = sdreport( saturated_fixed_effects_year_temp_dist_Obj, bias.correct=FALSE, getJointPrecision = FALSE )
saturated_fixed_effects_year_temp_dist_report = saturated_fixed_effects_year_temp_dist_Obj$report()

saturated_fixed_effects_year_temp_dist_output <- list(saturated_fixed_effects_year_temp_dist_Obj = saturated_fixed_effects_year_temp_dist_Obj,
                                                      saturated_fixed_effects_year_temp_dist_Opt = saturated_fixed_effects_year_temp_dist_Opt,
                                                      saturated_fixed_effects_year_temp_dist_report = saturated_fixed_effects_year_temp_dist_report)

save(saturated_fixed_effects_year_temp_dist_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "saturated_fixed_effects_year_temp_dist_output.rda"))
```


## Step 2: Tune the random effects, based on the model with saturated fixed effects

The primary random effects in this model are spatial and spatiotemporal random effects.

The options for the spatial random effects are as follows:
  
  1. No spatial random fields
2. Spatial random fields (isotropic)
3. Spatial random fields (anisotropic)

The options for the spatiotemporal random effects are as follows:
  
  1. No spatiotemporal random fields
2. Spatiotemporal random fields (isotropic, IID)
3. Spatiotemporal random fields (anisotropic, IID)
4. Spatiotemporal random fields (isotropic, AR1)
5. Spatiotemporal random fields (anisotropic, AR1)

However, it doesn't make ecological sense for spatial fields to be anisotropic and spatiotemporal fields to be isotropic, or vice versa (spatial fields isotropic and spatiotemporal fields anisotropic). So, the result of crossing these options with this constraint are the following models:

- 1-1 No spatial random fields, no spatiotemporal random fields
- 1-2 No spatial random fields, spatiotemporal random fields (isotropic, IID)
- 1-3 No spatial random fields, spatiotemporal random fields (anisotropic, IID)
- 1-4 No spatial random fields, spatiotemporal random fields (isotropic, AR1)
- 1-5 No spatial random fields, spatiotemporal random fields (anisotropic, AR1)

- 2-1 Spatial random fields (isotropic), no spatiotemporal random fields
- 2-2 Spatial random fields (isotropic), spatiotemporal random fields (isotropic, IID)
- 2-4 Spatial random fields (isotropic), spatiotemporal random fields (isotropic, AR1)

- 3-1 Spatial random fields (anisotropic), no spatiotemporal random fields
- 3-3 Spatial random fields (anisotropic), spatiotemporal random fields (anisotropic, IID)
- 3-5 Spatial random fields (anisotropic), spatiotemporal random fields (anisotropic, AR1)



The first cross (Model 1-1, no spatial random fields and no spatiotemporal random fields) is the saturated fixed effects model. Therefore, we will proceed with the ten other candidate random effect structures.

### Fit candidate models

#### Model 1-2: No spatial random fields, spatiotemporal random fields (isotropic, IID)

```{r model_1_2, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_2_spatial_off_spatiotemporal_IID_iso.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_2_spatial_off_spatiotemporal_IID_iso")))

model1_2_spatial_off_spatiotemporal_IID_iso_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
    # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model1_2_spatial_off_spatiotemporal_IID_iso_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0, # "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(bongo_biomass_spde$c0))
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model1_2_spatial_off_spatiotemporal_IID_iso_Data$n_t)
  # 
)

model1_2_spatial_off_spatiotemporal_IID_iso_Obj = MakeADFun( data=model1_2_spatial_off_spatiotemporal_IID_iso_Data,
                                         parameters=model1_2_spatial_off_spatiotemporal_IID_iso_Params,
                                         # random= c("omega_s"),
                                         random= c("epsilon_st"),
                                         DLL = "model1_2_spatial_off_spatiotemporal_IID_iso")

# Optimize
model1_2_spatial_off_spatiotemporal_IID_iso_Opt = nlminb( start=model1_2_spatial_off_spatiotemporal_IID_iso_Obj$par, obj=model1_2_spatial_off_spatiotemporal_IID_iso_Obj$fn, grad=model1_2_spatial_off_spatiotemporal_IID_iso_Obj$gr )
# add getJointPrecision for index
model1_2_spatial_off_spatiotemporal_IID_iso_Opt$SD = sdreport( model1_2_spatial_off_spatiotemporal_IID_iso_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model1_2_spatial_off_spatiotemporal_IID_iso_report = model1_2_spatial_off_spatiotemporal_IID_iso_Obj$report()

model1_2_spatial_off_spatiotemporal_IID_iso_output <- list(model1_2_spatial_off_spatiotemporal_IID_iso_Obj = model1_2_spatial_off_spatiotemporal_IID_iso_Obj,
                                       model1_2_spatial_off_spatiotemporal_IID_iso_Opt = model1_2_spatial_off_spatiotemporal_IID_iso_Opt,
                                       model1_2_spatial_off_spatiotemporal_IID_iso_report = model1_2_spatial_off_spatiotemporal_IID_iso_report)

save(model1_2_spatial_off_spatiotemporal_IID_iso_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model1_2_spatial_off_spatiotemporal_IID_iso_output.rda"))
```


#### Model 1-3: No spatial random fields, spatiotemporal random fields (anisotropic, IID)

```{r model_1_3, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_3_spatial_off_spatiotemporal_IID_aniso.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_3_spatial_off_spatiotemporal_IID_aniso")))

model1_3_spatial_off_spatiotemporal_IID_aniso_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model1_3_spatial_off_spatiotemporal_IID_aniso_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0, # "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(bongo_biomass_spde$c0))
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model1_3_spatial_off_spatiotemporal_IID_aniso_Data$n_t)
  # 
)

model1_3_spatial_off_spatiotemporal_IID_aniso_Obj = MakeADFun( data=model1_3_spatial_off_spatiotemporal_IID_aniso_Data,
                                                             parameters=model1_3_spatial_off_spatiotemporal_IID_aniso_Params,
                                                             # random= c("omega_s"),
                                                             random= c("epsilon_st"),
                                                             DLL = "model1_3_spatial_off_spatiotemporal_IID_aniso")

# Optimize
model1_3_spatial_off_spatiotemporal_IID_aniso_Opt = nlminb( start=model1_3_spatial_off_spatiotemporal_IID_aniso_Obj$par, obj=model1_3_spatial_off_spatiotemporal_IID_aniso_Obj$fn, grad=model1_3_spatial_off_spatiotemporal_IID_aniso_Obj$gr )
# add getJointPrecision for index
model1_3_spatial_off_spatiotemporal_IID_aniso_Opt$SD = sdreport( model1_3_spatial_off_spatiotemporal_IID_aniso_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model1_3_spatial_off_spatiotemporal_IID_aniso_report = model1_3_spatial_off_spatiotemporal_IID_aniso_Obj$report()

model1_3_spatial_off_spatiotemporal_IID_aniso_output <- list(model1_3_spatial_off_spatiotemporal_IID_aniso_Obj = model1_3_spatial_off_spatiotemporal_IID_aniso_Obj,
                                                           model1_3_spatial_off_spatiotemporal_IID_aniso_Opt = model1_3_spatial_off_spatiotemporal_IID_aniso_Opt,
                                                           model1_3_spatial_off_spatiotemporal_IID_aniso_report = model1_3_spatial_off_spatiotemporal_IID_aniso_report)

save(model1_3_spatial_off_spatiotemporal_IID_aniso_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model1_3_spatial_off_spatiotemporal_IID_aniso_output.rda"))
```

#### Model 1-4: No spatial random fields, spatiotemporal random fields (isotropic, AR1)

```{r model_1_4, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_4_spatial_off_spatiotemporal_AR1_iso.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_4_spatial_off_spatiotemporal_AR1_iso")))

model1_4_spatial_off_spatiotemporal_AR1_iso_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model1_4_spatial_off_spatiotemporal_AR1_iso_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0, # "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(bongo_biomass_spde$c0))
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model1_4_spatial_off_spatiotemporal_AR1_iso_Data$n_t)
  # 
)

model1_4_spatial_off_spatiotemporal_AR1_iso_Obj = MakeADFun( data=model1_4_spatial_off_spatiotemporal_AR1_iso_Data,
                                                             parameters=model1_4_spatial_off_spatiotemporal_AR1_iso_Params,
                                                             # random= c("omega_s"),
                                                             random= c("epsilon_st"),
                                                             DLL = "model1_4_spatial_off_spatiotemporal_AR1_iso")

# Optimize
model1_4_spatial_off_spatiotemporal_AR1_iso_Opt = nlminb( start=model1_4_spatial_off_spatiotemporal_AR1_iso_Obj$par, obj=model1_4_spatial_off_spatiotemporal_AR1_iso_Obj$fn, grad=model1_4_spatial_off_spatiotemporal_AR1_iso_Obj$gr )
# add getJointPrecision for index
model1_4_spatial_off_spatiotemporal_AR1_iso_Opt$SD = sdreport( model1_4_spatial_off_spatiotemporal_AR1_iso_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model1_4_spatial_off_spatiotemporal_AR1_iso_report = model1_4_spatial_off_spatiotemporal_AR1_iso_Obj$report()

model1_4_spatial_off_spatiotemporal_AR1_iso_output <- list(model1_4_spatial_off_spatiotemporal_AR1_iso_Obj = model1_4_spatial_off_spatiotemporal_AR1_iso_Obj,
                                                           model1_4_spatial_off_spatiotemporal_AR1_iso_Opt = model1_4_spatial_off_spatiotemporal_AR1_iso_Opt,
                                                           model1_4_spatial_off_spatiotemporal_AR1_iso_report = model1_4_spatial_off_spatiotemporal_AR1_iso_report)

save(model1_4_spatial_off_spatiotemporal_AR1_iso_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model1_4_spatial_off_spatiotemporal_AR1_iso_output.rda"))
```

#### Model 1-5: No spatial random fields, spatiotemporal random fields (anisotropic, AR1)

```{r model_1_5, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_5_spatial_off_spatiotemporal_AR1_aniso.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_5_spatial_off_spatiotemporal_AR1_aniso")))

model1_5_spatial_off_spatiotemporal_AR1_aniso_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model1_5_spatial_off_spatiotemporal_AR1_aniso_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0, # "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  # "omega_s"=rnorm(nrow(bongo_biomass_spde$c0))
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model1_5_spatial_off_spatiotemporal_AR1_aniso_Data$n_t)
  # 
)

model1_5_spatial_off_spatiotemporal_AR1_aniso_Obj = MakeADFun( data=model1_5_spatial_off_spatiotemporal_AR1_aniso_Data,
                                                             parameters=model1_5_spatial_off_spatiotemporal_AR1_aniso_Params,
                                                             # random= c("omega_s"),
                                                             random= c("epsilon_st"),
                                                             DLL = "model1_5_spatial_off_spatiotemporal_AR1_aniso")

# Optimize
model1_5_spatial_off_spatiotemporal_AR1_aniso_Opt = nlminb( start=model1_5_spatial_off_spatiotemporal_AR1_aniso_Obj$par, obj=model1_5_spatial_off_spatiotemporal_AR1_aniso_Obj$fn, grad=model1_5_spatial_off_spatiotemporal_AR1_aniso_Obj$gr )
# add getJointPrecision for index
model1_5_spatial_off_spatiotemporal_AR1_aniso_Opt$SD = sdreport( model1_5_spatial_off_spatiotemporal_AR1_aniso_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model1_5_spatial_off_spatiotemporal_AR1_aniso_report = model1_5_spatial_off_spatiotemporal_AR1_aniso_Obj$report()

model1_5_spatial_off_spatiotemporal_AR1_aniso_output <- list(model1_5_spatial_off_spatiotemporal_AR1_aniso_Obj = model1_5_spatial_off_spatiotemporal_AR1_aniso_Obj,
                                                           model1_5_spatial_off_spatiotemporal_AR1_aniso_Opt = model1_5_spatial_off_spatiotemporal_AR1_aniso_Opt,
                                                           model1_5_spatial_off_spatiotemporal_AR1_aniso_report = model1_5_spatial_off_spatiotemporal_AR1_aniso_report)

save(model1_5_spatial_off_spatiotemporal_AR1_aniso_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model1_5_spatial_off_spatiotemporal_AR1_aniso_output.rda"))
```


#### Model 2-1: Spatial random fields (isotropic), no spatiotemporal random fields


```{r model_2_1, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model2_1_spatial_on_spatiotemporal_off.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model2_1_spatial_on_spatiotemporal_off")))

model2_1_spatial_on_spatiotemporal_off_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model2_1_spatial_on_spatiotemporal_off_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  # "ln_tau_epsilon"=0, 
  "ln_tau_omega"=0,
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0))
  # "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model2_1_spatial_on_spatiotemporal_off_Data$n_t)
  # 
)

model2_1_spatial_on_spatiotemporal_off_Obj = MakeADFun( data=model2_1_spatial_on_spatiotemporal_off_Data,
                                                             parameters=model2_1_spatial_on_spatiotemporal_off_Params,
                                                             random= c("omega_s"),
                                                             # random= c("epsilon_st"),
                                                             DLL = "model2_1_spatial_on_spatiotemporal_off")

# Optimize
model2_1_spatial_on_spatiotemporal_off_Opt = nlminb( start=model2_1_spatial_on_spatiotemporal_off_Obj$par, obj=model2_1_spatial_on_spatiotemporal_off_Obj$fn, grad=model2_1_spatial_on_spatiotemporal_off_Obj$gr )
# add getJointPrecision for index
model2_1_spatial_on_spatiotemporal_off_Opt$SD = sdreport( model2_1_spatial_on_spatiotemporal_off_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model2_1_spatial_on_spatiotemporal_off_report = model2_1_spatial_on_spatiotemporal_off_Obj$report()

model2_1_spatial_on_spatiotemporal_off_output <- list(model2_1_spatial_on_spatiotemporal_off_Obj = model2_1_spatial_on_spatiotemporal_off_Obj,
                                                           model2_1_spatial_on_spatiotemporal_off_Opt = model2_1_spatial_on_spatiotemporal_off_Opt,
                                                           model2_1_spatial_on_spatiotemporal_off_report = model2_1_spatial_on_spatiotemporal_off_report)

save(model2_1_spatial_on_spatiotemporal_off_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model2_1_spatial_on_spatiotemporal_off_output.rda"))
```


#### Model 2-2: Spatial random fields (isotropic), spatiotemporal random fields (isotropic, IID)


```{r model_2_2, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model2_2_spatial_on_spatiotemporal_IID_iso.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model2_2_spatial_on_spatiotemporal_IID_iso")))

model2_2_spatial_on_spatiotemporal_IID_iso_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model2_2_spatial_on_spatiotemporal_IID_iso_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model2_2_spatial_on_spatiotemporal_IID_iso_Data$n_t)
  # 
)

model2_2_spatial_on_spatiotemporal_IID_iso_Obj = MakeADFun( data=model2_2_spatial_on_spatiotemporal_IID_iso_Data,
                                                             parameters=model2_2_spatial_on_spatiotemporal_IID_iso_Params,
                                                             random= c("omega_s", "epsilon_st"),
                                                             DLL = "model2_2_spatial_on_spatiotemporal_IID_iso")

# Optimize
model2_2_spatial_on_spatiotemporal_IID_iso_Opt = nlminb( start=model2_2_spatial_on_spatiotemporal_IID_iso_Obj$par, obj=model2_2_spatial_on_spatiotemporal_IID_iso_Obj$fn, grad=model2_2_spatial_on_spatiotemporal_IID_iso_Obj$gr )
# add getJointPrecision for index
model2_2_spatial_on_spatiotemporal_IID_iso_Opt$SD = sdreport( model2_2_spatial_on_spatiotemporal_IID_iso_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model2_2_spatial_on_spatiotemporal_IID_iso_report = model2_2_spatial_on_spatiotemporal_IID_iso_Obj$report()

model2_2_spatial_on_spatiotemporal_IID_iso_output <- list(model2_2_spatial_on_spatiotemporal_IID_iso_Obj = model2_2_spatial_on_spatiotemporal_IID_iso_Obj,
                                                           model2_2_spatial_on_spatiotemporal_IID_iso_Opt = model2_2_spatial_on_spatiotemporal_IID_iso_Opt,
                                                           model2_2_spatial_on_spatiotemporal_IID_iso_report = model2_2_spatial_on_spatiotemporal_IID_iso_report)

save(model2_2_spatial_on_spatiotemporal_IID_iso_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model2_2_spatial_on_spatiotemporal_IID_iso_output.rda"))
```


#### Model 2-4: Spatial random fields (isotropic), spatiotemporal random fields (isotropic, AR1)

```{r model_2_4, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model2_4_spatial_on_spatiotemporal_AR1_iso.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model2_4_spatial_on_spatiotemporal_AR1_iso")))

model2_4_spatial_on_spatiotemporal_AR1_iso_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model2_4_spatial_on_spatiotemporal_AR1_iso_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model2_4_spatial_on_spatiotemporal_AR1_iso_Data$n_t)
  # 
)

model2_4_spatial_on_spatiotemporal_AR1_iso_Obj = MakeADFun( data=model2_4_spatial_on_spatiotemporal_AR1_iso_Data,
                                                              parameters=model2_4_spatial_on_spatiotemporal_AR1_iso_Params,
                                                              random= c("omega_s", "epsilon_st"),
                                                              DLL = "model2_4_spatial_on_spatiotemporal_AR1_iso")

# Optimize
model2_4_spatial_on_spatiotemporal_AR1_iso_Opt = nlminb( start=model2_4_spatial_on_spatiotemporal_AR1_iso_Obj$par, obj=model2_4_spatial_on_spatiotemporal_AR1_iso_Obj$fn, grad=model2_4_spatial_on_spatiotemporal_AR1_iso_Obj$gr )
# add getJointPrecision for index
model2_4_spatial_on_spatiotemporal_AR1_iso_Opt$SD = sdreport( model2_4_spatial_on_spatiotemporal_AR1_iso_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model2_4_spatial_on_spatiotemporal_AR1_iso_report = model2_4_spatial_on_spatiotemporal_AR1_iso_Obj$report()

model2_4_spatial_on_spatiotemporal_AR1_iso_output <- list(model2_4_spatial_on_spatiotemporal_AR1_iso_Obj = model2_4_spatial_on_spatiotemporal_AR1_iso_Obj,
                                                            model2_4_spatial_on_spatiotemporal_AR1_iso_Opt = model2_4_spatial_on_spatiotemporal_AR1_iso_Opt,
                                                            model2_4_spatial_on_spatiotemporal_AR1_iso_report = model2_4_spatial_on_spatiotemporal_AR1_iso_report)

save(model2_4_spatial_on_spatiotemporal_AR1_iso_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model2_4_spatial_on_spatiotemporal_AR1_iso_output.rda"))
```

#### Model 3-1: Spatial random fields (anisotropic), no spatiotemporal random fields


```{r model_3_1, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model3_1_spatial_aniso_spatiotemporal_off.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model3_1_spatial_aniso_spatiotemporal_off")))

model3_1_spatial_aniso_spatiotemporal_off_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model3_1_spatial_aniso_spatiotemporal_off_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  # "ln_tau_epsilon"=0, 
  "ln_tau_omega"=0,
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0))
  # "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model3_1_spatial_aniso_spatiotemporal_off_Data$n_t)
  # 
)

model3_1_spatial_aniso_spatiotemporal_off_Obj = MakeADFun( data=model3_1_spatial_aniso_spatiotemporal_off_Data,
                                                        parameters=model3_1_spatial_aniso_spatiotemporal_off_Params,
                                                        random= c("omega_s"),
                                                        # random= c("epsilon_st"),
                                                        DLL = "model3_1_spatial_aniso_spatiotemporal_off")

# Optimize
model3_1_spatial_aniso_spatiotemporal_off_Opt = nlminb( start=model3_1_spatial_aniso_spatiotemporal_off_Obj$par, obj=model3_1_spatial_aniso_spatiotemporal_off_Obj$fn, grad=model3_1_spatial_aniso_spatiotemporal_off_Obj$gr )
# add getJointPrecision for index
model3_1_spatial_aniso_spatiotemporal_off_Opt$SD = sdreport( model3_1_spatial_aniso_spatiotemporal_off_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model3_1_spatial_aniso_spatiotemporal_off_report = model3_1_spatial_aniso_spatiotemporal_off_Obj$report()

model3_1_spatial_aniso_spatiotemporal_off_output <- list(model3_1_spatial_aniso_spatiotemporal_off_Obj = model3_1_spatial_aniso_spatiotemporal_off_Obj,
                                                      model3_1_spatial_aniso_spatiotemporal_off_Opt = model3_1_spatial_aniso_spatiotemporal_off_Opt,
                                                      model3_1_spatial_aniso_spatiotemporal_off_report = model3_1_spatial_aniso_spatiotemporal_off_report)

save(model3_1_spatial_aniso_spatiotemporal_off_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model3_1_spatial_aniso_spatiotemporal_off_output.rda"))
```




#### Model 3-3: Spatial random fields (anisotropic), spatiotemporal random fields (anisotropic, IID)

```{r model_3_3, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model3_3_spatial_aniso_spatiotemporal_IID_aniso_1.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model3_3_spatial_aniso_spatiotemporal_IID_aniso_1")))

model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Data$n_t)
  # 
)

model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Obj = MakeADFun( data=model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Data,
                                                            parameters=model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Params,
                                                            random= c("omega_s", "epsilon_st"),
                                                            DLL = "model3_3_spatial_aniso_spatiotemporal_IID_aniso_1")

# Optimize
model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Opt = nlminb( start=model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Obj$par, obj=model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Obj$fn, grad=model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Obj$gr )
# add getJointPrecision for index
model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Opt$SD = sdreport( model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_report = model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Obj$report()

model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_output <- list(model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Obj = model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Obj,
                                                          model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Opt = model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Opt,
                                                          model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_report = model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_report)

save(model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_output.rda"))
```



#### Model 3-5: Spatial random fields (anisotropic), spatiotemporal random fields (anisotropic, AR1)


```{r model_3_5, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "candidate_models",  "model3_5_spatial_aniso_spatiotemporal_AR1_aniso.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model3_5_spatial_aniso_spatiotemporal_AR1_aniso")))

model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Data$n_t)
  # 
)

model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Obj = MakeADFun( data=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Data,
                                                            parameters=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Params,
                                                            random= c("omega_s", "epsilon_st"),
                                                            DLL = "model3_5_spatial_aniso_spatiotemporal_AR1_aniso")

# Optimize
model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Opt = nlminb( start=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Obj$par, obj=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Obj$fn, grad=model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Obj$gr )
# add getJointPrecision for index
model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Opt$SD = sdreport( model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
model3_5_spatial_aniso_spatiotemporal_AR1_aniso_report = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Obj$report()

model3_5_spatial_aniso_spatiotemporal_AR1_aniso_output <- list(model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Obj = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Obj,
                                                          model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Opt = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Opt,
                                                          model3_5_spatial_aniso_spatiotemporal_AR1_aniso_report = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_report)

save(model3_5_spatial_aniso_spatiotemporal_AR1_aniso_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "model3_5_spatial_aniso_spatiotemporal_AR1_aniso_output.rda"))
```

### Compare candidate models

```{r random_effects_model_selection, cache = TRUE, include = FALSE}
# make sure that objects are loaded - need to do this if chunk outputs were cached
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models", "saturated_fixed_effects_year_temp_dist")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_2_spatial_off_spatiotemporal_IID_iso")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_3_spatial_off_spatiotemporal_IID_aniso")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_4_spatial_off_spatiotemporal_AR1_iso")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model1_5_spatial_off_spatiotemporal_AR1_aniso")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model2_1_spatial_on_spatiotemporal_off")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model2_2_spatial_on_spatiotemporal_IID_iso")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model2_4_spatial_on_spatiotemporal_AR1_iso")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model3_1_spatial_aniso_spatiotemporal_off")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model3_3_spatial_aniso_spatiotemporal_IID_aniso_1")))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "candidate_models",  "model3_5_spatial_aniso_spatiotemporal_AR1_aniso")))

# Calculate cAIC and AIC for each model

### Fixed effects
saturated_fixed_effects_year_temp_dist_cAIC <- cAIC(model_data = saturated_fixed_effects_year_temp_dist_Data,
                 model_obj = saturated_fixed_effects_year_temp_dist_Obj,  
                 model_parameters = saturated_fixed_effects_year_temp_dist_Params,
                 random = NULL,
                 DLL = "saturated_fixed_effects_year_temp_dist",
                 data_observation_names = "D_i")

saturated_fixed_effects_year_temp_dist_AIC <- 2*saturated_fixed_effects_year_temp_dist_Opt$objective + 2 * length(saturated_fixed_effects_year_temp_dist_Opt$par)

### Model 1-2
model1_2_spatial_off_spatiotemporal_IID_iso_cAIC <- cAIC(model_data = model1_2_spatial_off_spatiotemporal_IID_iso_Data,
                 model_obj = model1_2_spatial_off_spatiotemporal_IID_iso_Obj,  
                 model_parameters = model1_2_spatial_off_spatiotemporal_IID_iso_Params,
                 random = "epsilon_st",
                 DLL = "model1_2_spatial_off_spatiotemporal_IID_iso",
                 data_observation_names = "D_i")

model1_2_spatial_off_spatiotemporal_IID_iso_AIC <- 2*model1_2_spatial_off_spatiotemporal_IID_iso_Opt$objective + 2 * length(model1_2_spatial_off_spatiotemporal_IID_iso_Opt$par)

### Model 1-3
model1_3_spatial_off_spatiotemporal_IID_aniso_cAIC <- cAIC(model_data = model1_3_spatial_off_spatiotemporal_IID_aniso_Data,
                 model_obj = model1_3_spatial_off_spatiotemporal_IID_aniso_Obj,  
                 model_parameters = model1_3_spatial_off_spatiotemporal_IID_aniso_Params,
                 random = "epsilon_st",
                 DLL = "model1_3_spatial_off_spatiotemporal_IID_aniso",
                 data_observation_names = "D_i")

model1_3_spatial_off_spatiotemporal_IID_aniso_AIC <- 2*model1_3_spatial_off_spatiotemporal_IID_aniso_Opt$objective + 2 * length(model1_3_spatial_off_spatiotemporal_IID_aniso_Opt$par)

### Model 1-4
model1_4_spatial_off_spatiotemporal_AR1_iso_cAIC <- cAIC(model_data = model1_4_spatial_off_spatiotemporal_AR1_iso_Data,
                 model_obj = model1_4_spatial_off_spatiotemporal_AR1_iso_Obj,  
                 model_parameters = model1_4_spatial_off_spatiotemporal_AR1_iso_Params,
                 random = "epsilon_st",
                 DLL = "model1_4_spatial_off_spatiotemporal_AR1_iso",
                 data_observation_names = "D_i")

model1_4_spatial_off_spatiotemporal_AR1_iso_AIC <- 2*model1_4_spatial_off_spatiotemporal_AR1_iso_Opt$objective + 2 * length(model1_4_spatial_off_spatiotemporal_AR1_iso_Opt$par)

### Model 1-5
model1_5_spatial_off_spatiotemporal_AR1_aniso_cAIC <- cAIC(model_data = model1_5_spatial_off_spatiotemporal_AR1_aniso_Data,
                 model_obj = model1_5_spatial_off_spatiotemporal_AR1_aniso_Obj,  
                 model_parameters = model1_5_spatial_off_spatiotemporal_AR1_aniso_Params,
                 random = "epsilon_st",
                 DLL = "model1_5_spatial_off_spatiotemporal_AR1_aniso",
                 data_observation_names = "D_i")

model1_5_spatial_off_spatiotemporal_AR1_aniso_AIC <- 2*model1_5_spatial_off_spatiotemporal_AR1_aniso_Opt$objective + 2 * length(model1_5_spatial_off_spatiotemporal_AR1_aniso_Opt$par)

### Model 2-1
model2_1_spatial_on_spatiotemporal_off_cAIC <- cAIC(model_data = model2_1_spatial_on_spatiotemporal_off_Data,
                 model_obj = model2_1_spatial_on_spatiotemporal_off_Obj,  
                 model_parameters = model2_1_spatial_on_spatiotemporal_off_Params,
                 random = "omega_s",
                 DLL = "model2_1_spatial_on_spatiotemporal_off",
                 data_observation_names = "D_i")

model2_1_spatial_on_spatiotemporal_off_AIC <- 2*model2_1_spatial_on_spatiotemporal_off_Opt$objective + 2 * length(model2_1_spatial_on_spatiotemporal_off_Opt$par)

### Model 2-2
model2_2_spatial_on_spatiotemporal_IID_iso_cAIC <- cAIC(model_data = model2_2_spatial_on_spatiotemporal_IID_iso_Data,
                 model_obj = model2_2_spatial_on_spatiotemporal_IID_iso_Obj,  
                 model_parameters = model2_2_spatial_on_spatiotemporal_IID_iso_Params,
                 random = "omega_s",
                 DLL = "model2_2_spatial_on_spatiotemporal_IID_iso",
                 data_observation_names = "D_i")

model2_2_spatial_on_spatiotemporal_IID_iso_AIC <- 2*model2_2_spatial_on_spatiotemporal_IID_iso_Opt$objective + 2 * length(model2_2_spatial_on_spatiotemporal_IID_iso_Opt$par)

### Model 2-4
model2_4_spatial_on_spatiotemporal_AR1_iso_cAIC <- cAIC(model_data = model2_4_spatial_on_spatiotemporal_AR1_iso_Data,
                 model_obj = model2_4_spatial_on_spatiotemporal_AR1_iso_Obj,  
                 model_parameters = model2_4_spatial_on_spatiotemporal_AR1_iso_Params,
                 random = "omega_s",
                 DLL = "model2_4_spatial_on_spatiotemporal_AR1_iso",
                 data_observation_names = "D_i")

model2_4_spatial_on_spatiotemporal_AR1_iso_AIC <- 2*model2_4_spatial_on_spatiotemporal_AR1_iso_Opt$objective + 2 * length(model2_4_spatial_on_spatiotemporal_AR1_iso_Opt$par)

### Model 3-1
model3_1_spatial_aniso_spatiotemporal_off_cAIC <- cAIC(model_data = model3_1_spatial_aniso_spatiotemporal_off_Data,
                 model_obj = model3_1_spatial_aniso_spatiotemporal_off_Obj,  
                 model_parameters = model3_1_spatial_aniso_spatiotemporal_off_Params,
                 random = c("omega_s"),
                 DLL = "model3_1_spatial_aniso_spatiotemporal_off",
                 data_observation_names = "D_i")

model3_1_spatial_aniso_spatiotemporal_off_AIC <- 2*model3_1_spatial_aniso_spatiotemporal_off_Opt$objective + 2 * length(model3_1_spatial_aniso_spatiotemporal_off_Opt$par)

### Model 3-3
model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_cAIC <- cAIC(model_data = model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Data,
                 model_obj = model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Obj,  
                 model_parameters = model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "model3_3_spatial_aniso_spatiotemporal_IID_aniso_1",
                 data_observation_names = "D_i")

model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_AIC <- 2*model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Opt$objective + 2 * length(model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_Opt$par)

### Model 3-5
model3_5_spatial_aniso_spatiotemporal_AR1_aniso_cAIC <- cAIC(model_data = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Data,
                 model_obj = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Obj,  
                 model_parameters = model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "model3_5_spatial_aniso_spatiotemporal_AR1_aniso",
                 data_observation_names = "D_i")

model3_5_spatial_aniso_spatiotemporal_AR1_aniso_AIC <- 2*model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Opt$objective + 2 * length(model3_5_spatial_aniso_spatiotemporal_AR1_aniso_Opt$par)

# summarize this all in a table
random_effects_model_comparison_table <- data.frame(model = c("Spatial OFF + Spatiotemporal OFF",
                                                              "Spatial OFF + Spatiotemporal IID (isotropic)",
                                                              "Spatial OFF + Spatiotemporal IID (anisotropic)",
                                                              "Spatial OFF + Spatiotemporal AR1 (isotropic)",
                                                              "Spatial OFF + Spatiotemporal AR1 (anisotropic)",
                                                              "Spatial ON + Spatiotemporal OFF (isotropic)",
                                                              "Spatial ON + Spatiotemporal IID (isotropic)",
                                                              "Spatial ON + Spatiotemporal AR1 (isotropic)",
                                               "Spatial ON + Spatiotemporal OFF (anisotropic)", 
                                               "Spatial ON + Spatiotemporal IID (anisotropic)", 
                                               "Spatial ON + Spatiotemporal AR1 (anisotropic)"),
                                     AIC = c(saturated_fixed_effects_year_temp_dist_AIC,
                                             model1_2_spatial_off_spatiotemporal_IID_iso_AIC,
                                             model1_3_spatial_off_spatiotemporal_IID_aniso_AIC,
                                             model1_4_spatial_off_spatiotemporal_AR1_iso_AIC,
                                             model1_5_spatial_off_spatiotemporal_AR1_aniso_AIC,
                                             model2_1_spatial_on_spatiotemporal_off_AIC,
                                             model2_2_spatial_on_spatiotemporal_IID_iso_AIC,
                                             model2_4_spatial_on_spatiotemporal_AR1_iso_AIC,
                                             model3_1_spatial_aniso_spatiotemporal_off_AIC,
                                             model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_AIC,
                                             model3_5_spatial_aniso_spatiotemporal_AR1_aniso_AIC),
                                     cAIC = c(NA,
                                              model1_2_spatial_off_spatiotemporal_IID_iso_cAIC,
                                              model1_3_spatial_off_spatiotemporal_IID_aniso_cAIC,
                                              model1_4_spatial_off_spatiotemporal_AR1_iso_cAIC,
                                              model1_5_spatial_off_spatiotemporal_AR1_aniso_cAIC,
                                              model2_1_spatial_on_spatiotemporal_off_cAIC,
                                              model2_2_spatial_on_spatiotemporal_IID_iso_cAIC,
                                              model2_4_spatial_on_spatiotemporal_AR1_iso_cAIC,
                                              model3_1_spatial_aniso_spatiotemporal_off_cAIC,
                                              model3_3_spatial_aniso_spatiotemporal_IID_aniso_1_cAIC,
                                              model3_5_spatial_aniso_spatiotemporal_AR1_aniso_cAIC))
```

```{r random_effects_model_comparison_table}
random_effects_model_comparison_table %>% 
  arrange(cAIC) %>% 
  mutate(AIC = round(AIC, 2)) %>% 
  mutate(cAIC = round(cAIC, 2)) %>% 
  kbl(caption = "AIC and cAIC scores for the random effect structures.") %>% 
  kable_styling(latex_options = "HOLD_position")
```


## Step 3: Tune fixed effects based on random effects selected in step 2

We have three covariates that we are exploring:

- Year
- Temperature
- Distance from shore

Year we will only model as a factor representing a year-wide intercept for the full survey region, without an autoregressive component. This is because we do not expect the abundance of salmon to be an autoregressive process (successive run years are not related in terms of abundance, given the salmon life history).

Temperature we can model as a linear fixed effect or a smooth spline (technically a random effect).

Distance from shore we can model as a linear fixed effect or a smooth spline (technically a random effect).

Therefore, our options for fixed effects are as follows:

- Year, as a factor (on/off)
- Temperature (off/linear/spline)
- Distance from shore (off/linear/spline)


Crossing these options (2 x 3 x 3) results in the following 18 candidate fixed effect structures:

1. Year (factor) + temperature (linear) + distance from shore (linear) - this is the base model that we used to test random effect structures.
2. Year (factor) + temperature (smooth spline) + distance from shore (smooth spline)
3. Year (factor) + temperature (linear) + distance from shore (smooth spline)
4. Year (factor) + temperature (smooth spline) + distance from shore (linear)
5. Year (factor) + temperature (linear)
6. Year (factor) + distance from shore (linear)
7. Year (factor) + temperature (smooth spline)
8. Year (factor) + distance from shore (smooth spline)
9. Year (factor)
10. Distance from shore (smooth spline) + temperature (smooth spline)
11. Distance from shore (smooth spline) + temperature (linear)
12. Distance from shore (linear) + temperature (smooth spline)
13. Distance from shore (linear) + temperature (linear)
14. Distance from shore (smooth spline)
15. Distance from shore (linear)
16. Temperature (smooth spline)
17. Temperature (linear)
18. No fixed effects


### Fit candidate models

#### Model 2: Year (factor) + temperature (smooth spline) + distance from shore (smooth spline)

```{r model_2, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled) + s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                             dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled) + s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
      newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")
# load("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_stemp_sdist_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_sdist_year")))

non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
    # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Data,
                                                            parameters=non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Params,
                                                            random= c("omega_s", "epsilon_st", "b_smooth"),
                                                            DLL = "non_cancer_crab_larvae_bestfitRE_stemp_sdist_year")

# Optimize
non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_report = non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Obj$report()

non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_output <- list(non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Obj = non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Obj,
                                                          non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Opt = non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Opt,
                                                          non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_report = non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_report)

save(non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_output.rda"))
```

#### Model 3: Year (factor) + temperature (linear)

```{r model_3, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_year")))

non_cancer_crab_larvae_bestfitRE_temp_year_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


non_cancer_crab_larvae_bestfitRE_temp_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_temp_year_Data$n_t)
  # 
)

non_cancer_crab_larvae_bestfitRE_temp_year_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_temp_year_Data,
                                                                 parameters=non_cancer_crab_larvae_bestfitRE_temp_year_Params,
                                                                 random= c("omega_s", "epsilon_st"),
                                                                 DLL = "non_cancer_crab_larvae_bestfitRE_temp_year")

# Optimize
non_cancer_crab_larvae_bestfitRE_temp_year_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_temp_year_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_temp_year_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_temp_year_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_temp_year_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_temp_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_temp_year_report = non_cancer_crab_larvae_bestfitRE_temp_year_Obj$report()

non_cancer_crab_larvae_bestfitRE_temp_year_output <- list(non_cancer_crab_larvae_bestfitRE_temp_year_Obj = non_cancer_crab_larvae_bestfitRE_temp_year_Obj,
                                                               non_cancer_crab_larvae_bestfitRE_temp_year_Opt = non_cancer_crab_larvae_bestfitRE_temp_year_Opt,
                                                               non_cancer_crab_larvae_bestfitRE_temp_year_report = non_cancer_crab_larvae_bestfitRE_temp_year_report)

save(non_cancer_crab_larvae_bestfitRE_temp_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_year_output.rda"))
```

#### Model 4: Year (factor) + distance from shore (linear)

```{r model_4, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_dist_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_dist_year")))

non_cancer_crab_larvae_bestfitRE_dist_year_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  # "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


non_cancer_crab_larvae_bestfitRE_dist_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_dist_year_Data$n_t)
  # 
)

non_cancer_crab_larvae_bestfitRE_dist_year_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_dist_year_Data,
                                           parameters=non_cancer_crab_larvae_bestfitRE_dist_year_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "non_cancer_crab_larvae_bestfitRE_dist_year")

# Optimize
non_cancer_crab_larvae_bestfitRE_dist_year_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_dist_year_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_dist_year_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_dist_year_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_dist_year_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_dist_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_dist_year_report = non_cancer_crab_larvae_bestfitRE_dist_year_Obj$report()

non_cancer_crab_larvae_bestfitRE_dist_year_output <- list(non_cancer_crab_larvae_bestfitRE_dist_year_Obj = non_cancer_crab_larvae_bestfitRE_dist_year_Obj,
                                         non_cancer_crab_larvae_bestfitRE_dist_year_Opt = non_cancer_crab_larvae_bestfitRE_dist_year_Opt,
                                         non_cancer_crab_larvae_bestfitRE_dist_year_report = non_cancer_crab_larvae_bestfitRE_dist_year_report)

save(non_cancer_crab_larvae_bestfitRE_dist_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "non_cancer_crab_larvae_bestfitRE_dist_year_output.rda"))
```

#### Model 5: Year (factor) + temperature (smooth spline)

```{r model_5, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_stemp_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_year")))

non_cancer_crab_larvae_bestfitRE_stemp_year_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_stemp_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_stemp_year_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_stemp_year_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_stemp_year_Data,
                                                  parameters=non_cancer_crab_larvae_bestfitRE_stemp_year_Params,
                                                  random= c("omega_s", "epsilon_st", "b_smooth"),
                                                  DLL = "non_cancer_crab_larvae_bestfitRE_stemp_year")

# Optimize
non_cancer_crab_larvae_bestfitRE_stemp_year_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_stemp_year_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_stemp_year_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_stemp_year_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_stemp_year_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_stemp_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_stemp_year_report = non_cancer_crab_larvae_bestfitRE_stemp_year_Obj$report()

non_cancer_crab_larvae_bestfitRE_stemp_year_output <- list(non_cancer_crab_larvae_bestfitRE_stemp_year_Obj = non_cancer_crab_larvae_bestfitRE_stemp_year_Obj,
                                                non_cancer_crab_larvae_bestfitRE_stemp_year_Opt = non_cancer_crab_larvae_bestfitRE_stemp_year_Opt,
                                                non_cancer_crab_larvae_bestfitRE_stemp_year_report = non_cancer_crab_larvae_bestfitRE_stemp_year_report)

save(non_cancer_crab_larvae_bestfitRE_stemp_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_year_output.rda"))
```

#### Model 6: Year (factor) + s(dist)

```{r model_6, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(
  dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


save.image("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")
# load("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_sdist_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_sdist_year")))

non_cancer_crab_larvae_bestfitRE_sdist_year_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  # "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_sdist_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_sdist_year_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_sdist_year_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_sdist_year_Data,
                                                  parameters=non_cancer_crab_larvae_bestfitRE_sdist_year_Params,
                                                  random= c("omega_s", "epsilon_st", "b_smooth"),
                                                  DLL = "non_cancer_crab_larvae_bestfitRE_sdist_year")

# Optimize
non_cancer_crab_larvae_bestfitRE_sdist_year_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_sdist_year_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_sdist_year_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_sdist_year_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_sdist_year_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_sdist_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_sdist_year_report = non_cancer_crab_larvae_bestfitRE_sdist_year_Obj$report()

non_cancer_crab_larvae_bestfitRE_sdist_year_output <- list(non_cancer_crab_larvae_bestfitRE_sdist_year_Obj = non_cancer_crab_larvae_bestfitRE_sdist_year_Obj,
                                                non_cancer_crab_larvae_bestfitRE_sdist_year_Opt = non_cancer_crab_larvae_bestfitRE_sdist_year_Opt,
                                                non_cancer_crab_larvae_bestfitRE_sdist_year_report = non_cancer_crab_larvae_bestfitRE_sdist_year_report)

save(non_cancer_crab_larvae_bestfitRE_sdist_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_sdist_year_output.rda"))
```

#### Model 7: Year (factor)

```{r model_7, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_year")))

non_cancer_crab_larvae_bestfitRE_year_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  # "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


non_cancer_crab_larvae_bestfitRE_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_year_Data$n_t)
  # 
)

non_cancer_crab_larvae_bestfitRE_year_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_year_Data,
                                           parameters=non_cancer_crab_larvae_bestfitRE_year_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "non_cancer_crab_larvae_bestfitRE_year")

# Optimize
non_cancer_crab_larvae_bestfitRE_year_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_year_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_year_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_year_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_year_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_year_report = non_cancer_crab_larvae_bestfitRE_year_Obj$report()

non_cancer_crab_larvae_bestfitRE_year_output <- list(non_cancer_crab_larvae_bestfitRE_year_Obj = non_cancer_crab_larvae_bestfitRE_year_Obj,
                                         non_cancer_crab_larvae_bestfitRE_year_Opt = non_cancer_crab_larvae_bestfitRE_year_Opt,
                                         non_cancer_crab_larvae_bestfitRE_year_report = non_cancer_crab_larvae_bestfitRE_year_report)

save(non_cancer_crab_larvae_bestfitRE_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "non_cancer_crab_larvae_bestfitRE_year_output.rda"))
```

#### Model 8: No fixed effects

```{r model_8, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_noFE.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_noFE")))

non_cancer_crab_larvae_bestfitRE_noFE_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  # "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


non_cancer_crab_larvae_bestfitRE_noFE_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_noFE_Data$n_t)
  # 
)

non_cancer_crab_larvae_bestfitRE_noFE_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_noFE_Data,
                                           parameters=non_cancer_crab_larvae_bestfitRE_noFE_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "non_cancer_crab_larvae_bestfitRE_noFE")

# Optimize
non_cancer_crab_larvae_bestfitRE_noFE_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_noFE_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_noFE_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_noFE_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_noFE_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_noFE_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_noFE_report = non_cancer_crab_larvae_bestfitRE_noFE_Obj$report()

non_cancer_crab_larvae_bestfitRE_noFE_output <- list(non_cancer_crab_larvae_bestfitRE_noFE_Obj = non_cancer_crab_larvae_bestfitRE_noFE_Obj,
                                         non_cancer_crab_larvae_bestfitRE_noFE_Opt = non_cancer_crab_larvae_bestfitRE_noFE_Opt,
                                         non_cancer_crab_larvae_bestfitRE_noFE_report = non_cancer_crab_larvae_bestfitRE_noFE_report)

save(non_cancer_crab_larvae_bestfitRE_noFE_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "non_cancer_crab_larvae_bestfitRE_noFE_output.rda"))
```

#### Model 9: temperature (smooth spline) + distance from shore (smooth spline)

```{r model_9, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled) + s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled) + s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")
# load("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_stemp_sdist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_sdist")))

non_cancer_crab_larvae_bestfitRE_stemp_sdist_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_stemp_sdist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_stemp_sdist_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_stemp_sdist_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_stemp_sdist_Data,
                                                  parameters=non_cancer_crab_larvae_bestfitRE_stemp_sdist_Params,
                                                  random= c("omega_s", "epsilon_st", "b_smooth"),
                                                  DLL = "non_cancer_crab_larvae_bestfitRE_stemp_sdist")

# Optimize
non_cancer_crab_larvae_bestfitRE_stemp_sdist_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_stemp_sdist_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_stemp_sdist_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_stemp_sdist_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_stemp_sdist_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_stemp_sdist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_stemp_sdist_report = non_cancer_crab_larvae_bestfitRE_stemp_sdist_Obj$report()

non_cancer_crab_larvae_bestfitRE_stemp_sdist_output <- list(non_cancer_crab_larvae_bestfitRE_stemp_sdist_Obj = non_cancer_crab_larvae_bestfitRE_stemp_sdist_Obj,
                                                non_cancer_crab_larvae_bestfitRE_stemp_sdist_Opt = non_cancer_crab_larvae_bestfitRE_stemp_sdist_Opt,
                                                non_cancer_crab_larvae_bestfitRE_stemp_sdist_report = non_cancer_crab_larvae_bestfitRE_stemp_sdist_report)

save(non_cancer_crab_larvae_bestfitRE_stemp_sdist_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_sdist_output.rda"))
```

#### Model 10: temperature (linear)

```{r model_10, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp")))

non_cancer_crab_larvae_bestfitRE_temp_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


non_cancer_crab_larvae_bestfitRE_temp_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_temp_Data$n_t)
  # 
)

non_cancer_crab_larvae_bestfitRE_temp_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_temp_Data,
                                           parameters=non_cancer_crab_larvae_bestfitRE_temp_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "non_cancer_crab_larvae_bestfitRE_temp")

# Optimize
non_cancer_crab_larvae_bestfitRE_temp_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_temp_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_temp_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_temp_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_temp_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_temp_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_temp_report = non_cancer_crab_larvae_bestfitRE_temp_Obj$report()

non_cancer_crab_larvae_bestfitRE_temp_output <- list(non_cancer_crab_larvae_bestfitRE_temp_Obj = non_cancer_crab_larvae_bestfitRE_temp_Obj,
                                         non_cancer_crab_larvae_bestfitRE_temp_Opt = non_cancer_crab_larvae_bestfitRE_temp_Opt,
                                         non_cancer_crab_larvae_bestfitRE_temp_report = non_cancer_crab_larvae_bestfitRE_temp_report)

save(non_cancer_crab_larvae_bestfitRE_temp_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_output.rda"))
```

#### Model 11: distance from shore (linear)

```{r model_11, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_dist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_dist")))

non_cancer_crab_larvae_bestfitRE_dist_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  # "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


non_cancer_crab_larvae_bestfitRE_dist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_dist_Data$n_t)
  # 
)

non_cancer_crab_larvae_bestfitRE_dist_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_dist_Data,
                                           parameters=non_cancer_crab_larvae_bestfitRE_dist_Params,
                                           random= c("omega_s", "epsilon_st"),
                                           DLL = "non_cancer_crab_larvae_bestfitRE_dist")

# Optimize
non_cancer_crab_larvae_bestfitRE_dist_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_dist_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_dist_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_dist_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_dist_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_dist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_dist_report = non_cancer_crab_larvae_bestfitRE_dist_Obj$report()

non_cancer_crab_larvae_bestfitRE_dist_output <- list(non_cancer_crab_larvae_bestfitRE_dist_Obj = non_cancer_crab_larvae_bestfitRE_dist_Obj,
                                         non_cancer_crab_larvae_bestfitRE_dist_Opt = non_cancer_crab_larvae_bestfitRE_dist_Opt,
                                         non_cancer_crab_larvae_bestfitRE_dist_report = non_cancer_crab_larvae_bestfitRE_dist_report)

save(non_cancer_crab_larvae_bestfitRE_dist_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "non_cancer_crab_larvae_bestfitRE_dist_output.rda"))
```

#### Model 12: temperature (smooth spline)

```{r model_12, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_stemp.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp")))

non_cancer_crab_larvae_bestfitRE_stemp_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  # "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  # "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_stemp_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_stemp_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_stemp_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_stemp_Data,
                                            parameters=non_cancer_crab_larvae_bestfitRE_stemp_Params,
                                            random= c("omega_s", "epsilon_st", "b_smooth"),
                                            DLL = "non_cancer_crab_larvae_bestfitRE_stemp")

# Optimize
non_cancer_crab_larvae_bestfitRE_stemp_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_stemp_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_stemp_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_stemp_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_stemp_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_stemp_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_stemp_report = non_cancer_crab_larvae_bestfitRE_stemp_Obj$report()

non_cancer_crab_larvae_bestfitRE_stemp_output <- list(non_cancer_crab_larvae_bestfitRE_stemp_Obj = non_cancer_crab_larvae_bestfitRE_stemp_Obj,
                                          non_cancer_crab_larvae_bestfitRE_stemp_Opt = non_cancer_crab_larvae_bestfitRE_stemp_Opt,
                                          non_cancer_crab_larvae_bestfitRE_stemp_report = non_cancer_crab_larvae_bestfitRE_stemp_report)

save(non_cancer_crab_larvae_bestfitRE_stemp_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_output.rda"))
```

#### Model 13: s(dist)

```{r model_13, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(
  dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")
# load("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_sdist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_sdist")))

non_cancer_crab_larvae_bestfitRE_sdist_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  # "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  # "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_sdist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_sdist_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_sdist_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_sdist_Data,
                                            parameters=non_cancer_crab_larvae_bestfitRE_sdist_Params,
                                            random= c("omega_s", "epsilon_st", "b_smooth"),
                                            DLL = "non_cancer_crab_larvae_bestfitRE_sdist")

# Optimize
non_cancer_crab_larvae_bestfitRE_sdist_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_sdist_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_sdist_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_sdist_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_sdist_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_sdist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_sdist_report = non_cancer_crab_larvae_bestfitRE_sdist_Obj$report()

non_cancer_crab_larvae_bestfitRE_sdist_output <- list(non_cancer_crab_larvae_bestfitRE_sdist_Obj = non_cancer_crab_larvae_bestfitRE_sdist_Obj,
                                          non_cancer_crab_larvae_bestfitRE_sdist_Opt = non_cancer_crab_larvae_bestfitRE_sdist_Opt,
                                          non_cancer_crab_larvae_bestfitRE_sdist_report = non_cancer_crab_larvae_bestfitRE_sdist_report)

save(non_cancer_crab_larvae_bestfitRE_sdist_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_sdist_output.rda"))
```

#### Model 14: temperature (linear) + distance from shore (linear)

```{r model_14, cache = TRUE, include = FALSE}
compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_dist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_dist")))

non_cancer_crab_larvae_bestfitRE_temp_dist_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae
)


non_cancer_crab_larvae_bestfitRE_temp_dist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_temp_dist_Data$n_t)
  # 
)

non_cancer_crab_larvae_bestfitRE_temp_dist_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_temp_dist_Data,
                                                            parameters=non_cancer_crab_larvae_bestfitRE_temp_dist_Params,
                                                            random= c("omega_s", "epsilon_st"),
                                                            DLL = "non_cancer_crab_larvae_bestfitRE_temp_dist")

# Optimize
non_cancer_crab_larvae_bestfitRE_temp_dist_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_temp_dist_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_temp_dist_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_temp_dist_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_temp_dist_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_temp_dist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_temp_dist_report = non_cancer_crab_larvae_bestfitRE_temp_dist_Obj$report()

non_cancer_crab_larvae_bestfitRE_temp_dist_output <- list(non_cancer_crab_larvae_bestfitRE_temp_dist_Obj = non_cancer_crab_larvae_bestfitRE_temp_dist_Obj,
                                                          non_cancer_crab_larvae_bestfitRE_temp_dist_Opt = non_cancer_crab_larvae_bestfitRE_temp_dist_Opt,
                                                          non_cancer_crab_larvae_bestfitRE_temp_dist_report = non_cancer_crab_larvae_bestfitRE_temp_dist_report)

save(non_cancer_crab_larvae_bestfitRE_temp_dist_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_dist_output.rda"))
```

#### Model 15: Year (factor) + temperature (smooth spline) + distance from shore (linear)

```{r model_15, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")
# load("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_stemp_dist_year.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_dist_year")))

non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Data,
                                                  parameters=non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Params,
                                                  random= c("omega_s", "epsilon_st", "b_smooth"),
                                                  DLL = "non_cancer_crab_larvae_bestfitRE_stemp_dist_year")

# Optimize
non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_stemp_dist_year_report = non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Obj$report()

non_cancer_crab_larvae_bestfitRE_stemp_dist_year_output <- list(non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Obj = non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Obj,
                                                non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Opt = non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Opt,
                                                non_cancer_crab_larvae_bestfitRE_stemp_dist_year_report = non_cancer_crab_larvae_bestfitRE_stemp_dist_year_report)

save(non_cancer_crab_larvae_bestfitRE_stemp_dist_year_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_dist_year_output.rda"))
```

#### Model 16: Year (factor) + temperature (linear) + distance from shore (smooth spline)

```{r model_16, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")
# load("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1")))

non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Params <- list(
  
  ## SDM parameters
  "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Data,
                                                 parameters=non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Params,
                                                 random= c("omega_s", "epsilon_st", "b_smooth"),
                                                 DLL = "non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1")

# Optimize
non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_report = non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Obj$report()

non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_output <- list(non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Obj = non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Obj,
                                               non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Opt = non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Opt,
                                               non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_report = non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_report)

save(non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_output.rda"))
```

#### Model 17: temperature (smooth spline) + distance from shore (linear)

```{r model_17, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(SST_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")
# load("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_stemp_dist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_dist")))

non_cancer_crab_larvae_bestfitRE_stemp_dist_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_stemp_dist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  # "beta_temp" = 0,
  "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_stemp_dist_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_stemp_dist_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_stemp_dist_Data,
                                             parameters=non_cancer_crab_larvae_bestfitRE_stemp_dist_Params,
                                             random= c("omega_s", "epsilon_st", "b_smooth"),
                                             DLL = "non_cancer_crab_larvae_bestfitRE_stemp_dist")

# Optimize
non_cancer_crab_larvae_bestfitRE_stemp_dist_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_stemp_dist_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_stemp_dist_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_stemp_dist_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_stemp_dist_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_stemp_dist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_stemp_dist_report = non_cancer_crab_larvae_bestfitRE_stemp_dist_Obj$report()

non_cancer_crab_larvae_bestfitRE_stemp_dist_output <- list(non_cancer_crab_larvae_bestfitRE_stemp_dist_Obj = non_cancer_crab_larvae_bestfitRE_stemp_dist_Obj,
                                           non_cancer_crab_larvae_bestfitRE_stemp_dist_Opt = non_cancer_crab_larvae_bestfitRE_stemp_dist_Opt,
                                           non_cancer_crab_larvae_bestfitRE_stemp_dist_report = non_cancer_crab_larvae_bestfitRE_stemp_dist_report)

save(non_cancer_crab_larvae_bestfitRE_stemp_dist_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_stemp_dist_output.rda"))
```

#### Model 18: temperature (linear) + distance from shore (smooth spline)

```{r model_18, cache = TRUE, include = FALSE}
# to get the sdmTMB code to work, you just need to make sure that you provide the smooth terms in a formula (no need to write the complete formula)
sm <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL)

# get smoother values for predict grid
predict_grid <- data.frame(SST_scaled = as.vector(temperature_projection_matrix),
                           dist_shore_scaled = rep(distance_projection_vector, ncol(temperature_projection_matrix)))

sm_predict <- parse_smoothers(formula = "total_sum_of_final_carbon_mg_m3 ~ s(dist_shore_scaled)", data = as.data.frame(jsoes_bongo_biomass_non_cancer_crab_larvae), knots = NULL,
                              newdata = predict_grid, basis_prev = sm$basis_out)


# save.image("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")
# load("jsoes_bongo_biomass_non_cancer_crab_larvae_model_selection.RData")

compile(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models",  "non_cancer_crab_larvae_bestfitRE_temp_sdist.cpp"))
dyn.load(dynlib(here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_sdist")))

non_cancer_crab_larvae_bestfitRE_temp_sdist_Data = list(
  # number of years
  "n_t" = length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year)),
  # weights (all 1)
  "weights_i" = rep(1, length(jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3)),
  # temperature data
  "temp_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$SST_scaled,
  "temp_gt" = temperature_projection_matrix,
  # distance data
  "dist_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$dist_shore_scaled,
  "dist_g" = distance_projection_vector,
  
  # Subyearling Interior Fall Chinook data
  "D_i"= jsoes_bongo_biomass_non_cancer_crab_larvae$total_sum_of_final_carbon_mg_m3, "t_i" = jsoes_bongo_biomass_non_cancer_crab_larvae$year - min(jsoes_bongo_biomass_non_cancer_crab_larvae$year),
  "A_gs"=A_gs_non_cancer_crab_larvae,
  "A_is"=A_is_non_cancer_crab_larvae,
  # "M0_b"=bongo_biomass_spde$c0, "M1_b"=bongo_biomass_spde$g1, "M2_b"=bongo_biomass_spde$g2,
  "spatial_list" = spatial_list_non_cancer_crab_larvae,
  
  # smooth function data
  "Zs" = sm$Zs, # smoother basis function matrices
  "Xs" = sm$Xs, # smoother linear effect matrix
  "b_smooth_start" = sm$b_smooth_start,
  proj_Zs = sm_predict$Zs,
  proj_Xs = sm_predict$Xs
)

non_cancer_crab_larvae_bestfitRE_temp_sdist_Params <- list(
  
  ## SDM parameters
  # "beta_t"=rep(0, length(unique(jsoes_bongo_biomass_non_cancer_crab_larvae$year))),
  "beta_temp" = 0,
  # "beta_dist" = 0,
  "ln_tau_epsilon"=0,
  "ln_tau_omega"=0, 
  "ln_kappa"=0,
  "ln_phi" = 0,
  # "ln_H_input" = c(0, 0),
  "finv_power" = 0,
  
  "logit_rhoE" = 0,
  
  # SDM random effects
  "omega_s"=rnorm(nrow(bongo_biomass_spde$c0)),
  "epsilon_st"=matrix(0, nrow=nrow(bongo_biomass_spde$c0), ncol=non_cancer_crab_larvae_bestfitRE_temp_sdist_Data$n_t),
  
  # smooth functions
  bs = rep(0, ncol(sm$Xs)),
  b_smooth = rep(0, sum(sm$sm_dims)),
  ln_smooth_sigma = rep(0, length(sm$sm_dims))
)

non_cancer_crab_larvae_bestfitRE_temp_sdist_Obj = MakeADFun( data=non_cancer_crab_larvae_bestfitRE_temp_sdist_Data,
                                             parameters=non_cancer_crab_larvae_bestfitRE_temp_sdist_Params,
                                             random= c("omega_s", "epsilon_st", "b_smooth"),
                                             DLL = "non_cancer_crab_larvae_bestfitRE_temp_sdist")

# Optimize
non_cancer_crab_larvae_bestfitRE_temp_sdist_Opt = nlminb( start=non_cancer_crab_larvae_bestfitRE_temp_sdist_Obj$par, obj=non_cancer_crab_larvae_bestfitRE_temp_sdist_Obj$fn, grad=non_cancer_crab_larvae_bestfitRE_temp_sdist_Obj$gr )
# add getJointPrecision for index
non_cancer_crab_larvae_bestfitRE_temp_sdist_Opt$SD = sdreport( non_cancer_crab_larvae_bestfitRE_temp_sdist_Obj, bias.correct=TRUE, getJointPrecision = TRUE )
non_cancer_crab_larvae_bestfitRE_temp_sdist_report = non_cancer_crab_larvae_bestfitRE_temp_sdist_Obj$report()

non_cancer_crab_larvae_bestfitRE_temp_sdist_output <- list(non_cancer_crab_larvae_bestfitRE_temp_sdist_Obj = non_cancer_crab_larvae_bestfitRE_temp_sdist_Obj,
                                           non_cancer_crab_larvae_bestfitRE_temp_sdist_Opt = non_cancer_crab_larvae_bestfitRE_temp_sdist_Opt,
                                           non_cancer_crab_larvae_bestfitRE_temp_sdist_report = non_cancer_crab_larvae_bestfitRE_temp_sdist_report)

save(non_cancer_crab_larvae_bestfitRE_temp_sdist_output,
     file = here::here("R", "04_SDM_model_selection", "04.6_non_cancer_crab_larvae_SDM", "non_cancer_crab_larvae_fixed_effects_candidate_models", "non_cancer_crab_larvae_bestfitRE_temp_sdist_output.rda"))
```



### Compare candidate models

```{r, cache = TRUE, include = FALSE}
non_cancer_crab_larvae_bestfitRE_stemp_dist_year_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Data,
                                              model_obj = non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Obj,  
                                              model_parameters = non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "non_cancer_crab_larvae_bestfitRE_stemp_dist_year",
                                              data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_stemp_dist_year_AIC <- 2*non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_stemp_dist_year_Opt$par)

non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Data,
                                              model_obj = non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Obj,  
                                              model_parameters = non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1",
                                              data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_AIC <- 2*non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_Opt$par)

non_cancer_crab_larvae_bestfitRE_temp_sdist_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_temp_sdist_Data,
                                              model_obj = non_cancer_crab_larvae_bestfitRE_temp_sdist_Obj,  
                                              model_parameters = non_cancer_crab_larvae_bestfitRE_temp_sdist_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "non_cancer_crab_larvae_bestfitRE_temp_sdist",
                                              data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_temp_sdist_AIC <- 2*non_cancer_crab_larvae_bestfitRE_temp_sdist_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_temp_sdist_Opt$par)

non_cancer_crab_larvae_bestfitRE_stemp_dist_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_stemp_dist_Data,
                                              model_obj = non_cancer_crab_larvae_bestfitRE_stemp_dist_Obj,  
                                              model_parameters = non_cancer_crab_larvae_bestfitRE_stemp_dist_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "non_cancer_crab_larvae_bestfitRE_stemp_dist",
                                              data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_stemp_dist_AIC <- 2*non_cancer_crab_larvae_bestfitRE_stemp_dist_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_stemp_dist_Opt$par)


```


```{r, cache = TRUE, include = FALSE}
non_cancer_crab_larvae_bestfitRE_stemp_sdist_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_stemp_sdist_Data,
                                              model_obj = non_cancer_crab_larvae_bestfitRE_stemp_sdist_Obj,  
                                              model_parameters = non_cancer_crab_larvae_bestfitRE_stemp_sdist_Params,
                                              random = c("omega_s", "epsilon_st", "b_smooth"),
                                              DLL = "non_cancer_crab_larvae_bestfitRE_stemp_sdist",
                                              data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_stemp_sdist_AIC <- 2*non_cancer_crab_larvae_bestfitRE_stemp_sdist_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_stemp_sdist_Opt$par)

non_cancer_crab_larvae_bestfitRE_temp_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_temp_Data,
                                       model_obj = non_cancer_crab_larvae_bestfitRE_temp_Obj,  
                                       model_parameters = non_cancer_crab_larvae_bestfitRE_temp_Params,
                                       random = c("omega_s", "epsilon_st"),
                                       DLL = "non_cancer_crab_larvae_bestfitRE_temp",
                                       data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_temp_AIC <- 2*non_cancer_crab_larvae_bestfitRE_temp_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_temp_Opt$par)

non_cancer_crab_larvae_bestfitRE_dist_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_dist_Data,
                                       model_obj = non_cancer_crab_larvae_bestfitRE_dist_Obj,  
                                       model_parameters = non_cancer_crab_larvae_bestfitRE_dist_Params,
                                       random = c("omega_s", "epsilon_st"),
                                       DLL = "non_cancer_crab_larvae_bestfitRE_dist",
                                       data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_dist_AIC <- 2*non_cancer_crab_larvae_bestfitRE_dist_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_dist_Opt$par)

non_cancer_crab_larvae_bestfitRE_stemp_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_stemp_Data,
                                        model_obj = non_cancer_crab_larvae_bestfitRE_stemp_Obj,  
                                        model_parameters = non_cancer_crab_larvae_bestfitRE_stemp_Params,
                                        random = c("omega_s", "epsilon_st", "b_smooth"),
                                        DLL = "non_cancer_crab_larvae_bestfitRE_stemp",
                                        data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_stemp_AIC <- 2*non_cancer_crab_larvae_bestfitRE_stemp_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_stemp_Opt$par)

non_cancer_crab_larvae_bestfitRE_sdist_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_sdist_Data,
                                        model_obj = non_cancer_crab_larvae_bestfitRE_sdist_Obj,  
                                        model_parameters = non_cancer_crab_larvae_bestfitRE_sdist_Params,
                                        random = c("omega_s", "epsilon_st", "b_smooth"),
                                        DLL = "non_cancer_crab_larvae_bestfitRE_sdist",
                                        data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_sdist_AIC <- 2*non_cancer_crab_larvae_bestfitRE_sdist_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_sdist_Opt$par)

non_cancer_crab_larvae_bestfitRE_temp_dist_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_temp_dist_Data,
                                  model_obj = non_cancer_crab_larvae_bestfitRE_temp_dist_Obj,  
                                  model_parameters = non_cancer_crab_larvae_bestfitRE_temp_dist_Params,
                                  random = c("omega_s", "epsilon_st"),
                                  DLL = "non_cancer_crab_larvae_bestfitRE_temp_dist",
                                  data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_temp_dist_AIC <- 2*non_cancer_crab_larvae_bestfitRE_temp_dist_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_temp_dist_Opt$par)

```


```{r, cache = TRUE, include = FALSE}
non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Data,
                 model_obj = non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Obj,  
                 model_parameters = non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Params,
                 random = c("omega_s", "epsilon_st", "b_smooth"),
                 DLL = "non_cancer_crab_larvae_bestfitRE_stemp_sdist_year",
                 data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_AIC <- 2*non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_Opt$par)

non_cancer_crab_larvae_bestfitRE_temp_year_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_temp_year_Data,
                 model_obj = non_cancer_crab_larvae_bestfitRE_temp_year_Obj,  
                 model_parameters = non_cancer_crab_larvae_bestfitRE_temp_year_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "non_cancer_crab_larvae_bestfitRE_temp_year",
                 data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_temp_year_AIC <- 2*non_cancer_crab_larvae_bestfitRE_temp_year_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_temp_year_Opt$par)

non_cancer_crab_larvae_bestfitRE_dist_year_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_dist_year_Data,
                 model_obj = non_cancer_crab_larvae_bestfitRE_dist_year_Obj,  
                 model_parameters = non_cancer_crab_larvae_bestfitRE_dist_year_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "non_cancer_crab_larvae_bestfitRE_dist_year",
                 data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_dist_year_AIC <- 2*non_cancer_crab_larvae_bestfitRE_dist_year_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_dist_year_Opt$par)

non_cancer_crab_larvae_bestfitRE_stemp_year_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_stemp_year_Data,
                 model_obj = non_cancer_crab_larvae_bestfitRE_stemp_year_Obj,  
                 model_parameters = non_cancer_crab_larvae_bestfitRE_stemp_year_Params,
                 random = c("omega_s", "epsilon_st", "b_smooth"),
                 DLL = "non_cancer_crab_larvae_bestfitRE_stemp_year",
                 data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_stemp_year_AIC <- 2*non_cancer_crab_larvae_bestfitRE_stemp_year_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_stemp_year_Opt$par)

non_cancer_crab_larvae_bestfitRE_sdist_year_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_sdist_year_Data,
                 model_obj = non_cancer_crab_larvae_bestfitRE_sdist_year_Obj,  
                 model_parameters = non_cancer_crab_larvae_bestfitRE_sdist_year_Params,
                 random = c("omega_s", "epsilon_st", "b_smooth"),
                 DLL = "non_cancer_crab_larvae_bestfitRE_sdist_year",
                 data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_sdist_year_AIC <- 2*non_cancer_crab_larvae_bestfitRE_sdist_year_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_sdist_year_Opt$par)

non_cancer_crab_larvae_bestfitRE_year_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_year_Data,
                 model_obj = non_cancer_crab_larvae_bestfitRE_year_Obj,  
                 model_parameters = non_cancer_crab_larvae_bestfitRE_year_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "non_cancer_crab_larvae_bestfitRE_year",
                 data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_year_AIC <- 2*non_cancer_crab_larvae_bestfitRE_year_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_year_Opt$par)

non_cancer_crab_larvae_bestfitRE_noFE_cAIC <- cAIC(model_data = non_cancer_crab_larvae_bestfitRE_noFE_Data,
                 model_obj = non_cancer_crab_larvae_bestfitRE_noFE_Obj,  
                 model_parameters = non_cancer_crab_larvae_bestfitRE_noFE_Params,
                 random = c("omega_s", "epsilon_st"),
                 DLL = "non_cancer_crab_larvae_bestfitRE_noFE",
                 data_observation_names = "D_i")

non_cancer_crab_larvae_bestfitRE_noFE_AIC <- 2*non_cancer_crab_larvae_bestfitRE_noFE_Opt$objective + 2 * length(non_cancer_crab_larvae_bestfitRE_noFE_Opt$par)

```


```{r}
fixed_effects_model_comparison_table <- data.frame(model = c("as.factor(year) + temp + dist",
                                                             "as.factor(year) + s(temp) + s(dist)",
                                                             "as.factor(year) + temp",
                                                             "as.factor(year) + dist",
                                                             "as.factor(year) + s(temp)",
                                                             "as.factor(year) + s(dist)",
                                                             "as.factor(year)",
                                                             "temp + dist",
                                                             "s(temp) + s(dist)",
                                                             "temp",
                                                             "dist",
                                                             "s(temp)",
                                                             "s(dist)",
                                                             "as.factor(year) + s(temp) + dist",
                                                             "as.factor(year) + temp + s(dist)",
                                                             "temp + s(dist)",
                                                             "s(temp) + dist",
                                                             "no fixed effects"),
                                     AIC = c(model2_4_spatial_on_spatiotemporal_AR1_iso_AIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_AIC,
                                             non_cancer_crab_larvae_bestfitRE_temp_year_AIC,
                                             non_cancer_crab_larvae_bestfitRE_dist_year_AIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_year_AIC,
                                             non_cancer_crab_larvae_bestfitRE_sdist_year_AIC,
                                             non_cancer_crab_larvae_bestfitRE_year_AIC,
                                             non_cancer_crab_larvae_bestfitRE_temp_dist_AIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_sdist_AIC,
                                             non_cancer_crab_larvae_bestfitRE_temp_AIC,
                                             non_cancer_crab_larvae_bestfitRE_dist_AIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_AIC,
                                             non_cancer_crab_larvae_bestfitRE_sdist_AIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_dist_year_AIC,
                                             non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_AIC,
                                             non_cancer_crab_larvae_bestfitRE_temp_sdist_AIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_dist_AIC,
                                             non_cancer_crab_larvae_bestfitRE_noFE_AIC),
                                     cAIC = c(model2_4_spatial_on_spatiotemporal_AR1_iso_cAIC,
                                              non_cancer_crab_larvae_bestfitRE_stemp_sdist_year_cAIC,
                                              non_cancer_crab_larvae_bestfitRE_temp_year_cAIC,
                                              non_cancer_crab_larvae_bestfitRE_dist_year_cAIC,
                                              non_cancer_crab_larvae_bestfitRE_stemp_year_cAIC,
                                              non_cancer_crab_larvae_bestfitRE_sdist_year_cAIC,
                                              non_cancer_crab_larvae_bestfitRE_year_cAIC,
                                              non_cancer_crab_larvae_bestfitRE_temp_dist_cAIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_sdist_cAIC,
                                             non_cancer_crab_larvae_bestfitRE_temp_cAIC,
                                             non_cancer_crab_larvae_bestfitRE_dist_cAIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_cAIC,
                                             non_cancer_crab_larvae_bestfitRE_sdist_cAIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_dist_year_cAIC,
                                             non_cancer_crab_larvae_bestfitRE_temp_sdist_year_1_cAIC,
                                             non_cancer_crab_larvae_bestfitRE_temp_sdist_cAIC,
                                             non_cancer_crab_larvae_bestfitRE_stemp_dist_cAIC,
                                              non_cancer_crab_larvae_bestfitRE_noFE_cAIC))

fixed_effects_model_comparison_table %>% 
  arrange(cAIC) %>% 
  mutate(AIC = round(AIC, 2)) %>% 
  mutate(cAIC = round(cAIC, 2)) %>% 
  kbl(caption = "AIC and cAIC scores for the fixed effect structures.") %>% 
  kable_styling(latex_options = "HOLD_position")
```




#### Export the tables of AIC and cAIC for the supplemental results
```{r}
write.csv(random_effects_model_comparison_table, here::here("R", "04_SDM_model_selection", "04.6.1_non_cancer_crab_larvae_biomass_SDM", "04.6.1_non_cancer_crab_larvae_biomass_random_effects_model_comparison_table.csv"))
write.csv(fixed_effects_model_comparison_table, here::here("R", "04_SDM_model_selection", "04.6.1_non_cancer_crab_larvae_biomass_SDM", "04.6.1_non_cancer_crab_larvae_biomass_fixed_effects_model_comparison_table.csv"))
```



